### 2021-07-26 dissertation/msc_dissertation_introduction.Rmd
Initialising (again)

---
title: "__How analysis strategy affects analysis results__"
subtitle: "Vibration of effects due to model specifications"
author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
linestretch: 1.5
output: pdf_document
bibliography: references.bib
csl: apa.csl
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

@silberzahn2018 investigated the effects of analytical choices on results. Specifically, they had 29 teams of researchers investigate the same research question with the same dataset. Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g. wanting/needing to publish). The variation between the analytic approaches was substantial. There were 29 different analyses with 21 different combinations of covariates. Twenty teams found a statistically significant effect, and the odds-ratios ranged from 0.89 to 2.93 (median = 1.31). The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results.

This study puts (psychological) research in a delicate position, because if researchers with honest intentions come to such drastically different conclusions despite starting off on the identical dataset what does this mean for the rest of studies?

The Covid-19 pandemic has reaffirmed the need for rigorous scientific research, but the process is not straightforward. In fact, Ioannidis 2005 argued that most study findings are false positives due to flawed study designs and insufficient sample sizes.

@silberzahn2018

@patel2015

@haessler2020

# Methods

\newpage
# References

### 2021-07-26 dissertation/msc_dissertation_script.Rmd
Initialising (again)

---
title             : "Understanding How Analysis Strategy Affects Analysis Results"
shorttitle        : "Vibration of Effects due to Model Specifications"

author: 
  - name          : "Sebastian Ploner"
    affiliation   : "*"

affiliation:
  - id            : "*"
    institution   : "The University of Sheffield"

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2021-07-26 dissertation/msc_dissertation_script.tex
Initialising (again)

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Understanding How Analysis Strategy Affects Analysis Results},
  pdfauthor={Sebastian Ploner*},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Vibration of Effects due to Model Specifications}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Understanding How Analysis Strategy Affects Analysis Results}
\author{Sebastian Ploner\textsuperscript{*}}
\date{}

\affiliation{\vspace{0.5cm}\textsuperscript{*} The University of Sheffield}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{material}{%
\subsection{Material}\label{material}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We used R {[}Version 4.0.4; R Core Team (2021){]} and the R-package \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2020){]} for all our analyses.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2020). \emph{{papaja}: {Create} {APA} manuscripts with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2021). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\end{CSLReferences}

\endgroup

\end{document}

### 2021-07-26 scripts/2_data_disaggregate.py

@@ -1,58 +1,0 @@
-print('importing libraries')
-import pandas as pd #for dealing with csv import
-import os # for joining paths and filenames sensibly
-# import matplotlib.pyplot as plt  # Matplotlib's pyplot: MATLAB-like syntax
-
-print('loading datafiles')
-# set wkdir
-file_path = '/Users/sebastian/Documents/Uni/Sheffield (MSc)/2. Semester/Research Project/msc_dissertation_project'
-os.chdir(file_path)
-
-# nb new datase
-filename = os.path.join('data', '1_crowdstorming_data.csv') 
-df = pd.read_csv(filename)
-
-#add new vars
-df['skintone'] = (df['rater1'] + df['rater2']) / 2
-df['allreds'] = df['yellowReds'] + df['redCards']
-df['allredsStrict'] = df['redCards']
-df['refCount'] = 0
-
-
-#add a column which tracks how many games each ref is involved in
-refs = pd.unique(df['refNum'].values.ravel()) #list all unique ref IDs
-
-#for each ref, count their dyads
-for r in refs:
-    df['refCount'][df['refNum'] == r] = len(df[df['refNum'] == r])    
-
-# for r in refs:
-#     if len(df[df['refNum'] == refs[r]]) > 1:
-#        print len(df[df['refNum'] == refs[r]])
-
-
-colnames = list(df.columns)
-
-
-j = 0
-out = [0 for _ in range(sum(df['games']))]
-
-for _, row in df.iterrows():
-        n = row['games']
-        c = row['allreds']
-        d = row['allredsStrict']
-        
-        #row['games'] = 1        
-        
-        for _ in range(n):
-                row['allreds'] = 1 if (c-_) > 0 else 0
-                row['allredsStrict'] = 1 if (d-_) > 0 else 0
-                rowlist = list(row)  #convert from pandas Series to prevent overwriting previous values of out[j]
-                out[j] = rowlist
-                j += 1
-                if j%10 == 0:    
-                    print("Number " + str(j) + " of " + str(df.shape[0]))
-
-
-#pd.DataFrame(out, columns = colnames).to_csv('2_disaggregated_data.csv', index = False) #have to make column headers explicit
-pd.DataFrame(out, columns = colnames).to_csv(file_path + '/data/2_disaggregated_data.csv', index = False) 

### 2021-07-26 scripts/3_data_preparation.R

@@ -1,110 +1,117 @@
-library("lme4")
 library("here")
+library("data.table")
+library("lme4")
 
 # from raw data generate disaggregated data for logistic regression via 2_data_disaggregate.py
-dat = read.csv(here("data", "2_disaggregated_data.csv"))
+dat = data.frame(fread(here("data", "2_disaggregated_data.csv")))
 
-# Select only the columns we will use in the analysis
-dat <- dat[, c("playerShort", "position", "redCards", "refNum", "meanIAT", "meanExp", "skintone", "allreds", "leagueCountry","refCount","allredsStrict")]
 
-#if we're on windows or mac you need to force the bias variables into numeric
+# select only the columns we will use in the analysis
+# adding all used covariates mentioned in Silberzahn et al. (2018) incl. main variables
+dat <- dat[, c("playerShort", "height", "weight", "birthday", "position",
+               "club", "leagueCountry",
+               "games",  "goals", "victories", "ties", 
+               "yellowCards", "yellowReds", "redCards", "allreds", "allredsStrict",
+               "refNum", "refCountry", "refCount",
+               "meanIAT", "meanExp", "skintone")]
+
+# if we're on windows or mac you need to force the bias variables into numeric
 dat$meanIAT <- as.numeric(as.character(dat$meanIAT))
 dat$meanExp <- as.numeric(as.character(dat$meanExp))
 
-# Remove NA skintone data
+# remove NA skintone data
 dat <- dat[!is.na(dat$skintone), ]
 
-#remove refs present in <22 dyads
-#reduces 373067 obs. by 16958
+# remove refs present in <22 dyads
+# reduces 373067 obs. by 16958
 dat <- dat[dat$refCount > 21, ]
 
-# Remove those individuals on whom we do not have position information
+# remove those individuals on whom we do not have position information
 NoPos <- which(dat$position == "")
 dat <- dat[-NoPos, ]
 
-# Remove those where referee has unknown bias
+# remove those where referee has unknown bias
 dat <- dat[!is.na(dat$meanIAT), ]
 dat <- dat[!is.na(dat$meanExp), ]
 
 # n = 335537 games
 
-# See what the current structure of the data is
+# see what the current structure of the data is
 str(dat)
 
-# Find out how many observations we have
+# find out how many observations we have
 ndat <- nrow(dat)
 
-# Now create another variable which splits individuals into GK, Defender, Midfielder, Attacker
+# now create another variable which splits individuals into GK, Defender, Midfielder, Attacker
[…]

### 2021-07-26 scripts/Functions.R

@@ -1,816 +1,0 @@
-#..........................................................................................................#
-#..........................................................................................................#
-#................................ Specification Curve Analysis ............................................#
-#..........................................................................................................#
-#..........................................................................................................#
-
-#...# Prepare variables needed for specification curve analysis
-specVars <- function(whichData) {
-  varList <- list()
-  
-  # Full variable names / descriptions (variables)
-  varList[["variables"]] <- c("Working in solidarity",           
-                              "Support empowering policies",
-                              "Raising IG awareness",
-                              "High cost collective action",
-                              "Low cost collective action",
-                              "Quantity of contact",
-                              "Positive contact",
-                              "Absence of negative contact",
-                              "Number of OG friends",
-                              "Frequency of meeting OG friends",
-                              "Quantity of indirect OG friends",
-                              "Positive indirect contact",
-                              "Absence of negative indirect contact",
-                              "Exclusion of attention check failures",
-                              "Exclusion of statistical outliers")
-  
-  # Support for social change variables (dv)
-  varList[["dv"]] <- c("WRKSO","SEMP","IDISC", "WILAC_D","WILAC_E")
-  
-  # Intergroup contact variables (iv)
-  varList[["iv"]] <- c("CQUA", "CQUAL", "NEG_R", "NFRI","FFRI", "EXTC_CQUA", "EXTC_CQUAL", "EXTC_NEG_R") 
-  
-  # Other variables: Attention check (var.other)
-  varList[["var.other"]] <- c("ATTCHK")  
-  
-  # Adapt "variables" and "iv" for datasets without quantity of contact and quantity of indirect OG friends 
-  # (SM, HE_short, AG_short, DG_short)
[…]

### 2021-07-26 scripts/Master_Script.R

@@ -1,464 +1,0 @@
-# WARNING: The following line of code will remove all objects from the workspace. 
-#          Please consider saving your workspace.
-rm(list=ls()) 
-
-
-#................................# Set parameters #..................................#
-
-# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
-# Please read the comments in this section to understand what each parameter does.
-# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
-# sourcing the entire script or running the "Prepare R Session" section.
-
-
-#... Set number of analyses with shuffled data
-R <- 1
-# This determines how many times the original data will be shuffled in the permutation test
-# To reproduce our results, set R to 1000
-# For quick, approximate calculations, please set R to, e.g., 5 or 10
-
-
-
-#... Define which group of datasets to load
-whichData <- "default"
-# Please choose one of the following options:
-# default: sub-sample mean differences removed (sample mean residualization)
-# keyDem: controlling for key demographic variables (SES ladder, age, gender). In the chosen datasets, these variables were partialled out from all contact and social change items.
-
-
-
-#... Define population for which specification curve analysis will be conducted
-dataset <- "AG"
-# In the line above, please indicate for which population the specification curve analysis should be conducted
-# Possible options:
-# AG       (ethnic majorities)
-# HE       (cis-heterosexuals)
-# DG       (ethnic minorities)
-# SM       (LGBTIQ+ individuals)
-
-# AG_short (ethnic majorities w/o quantity of direct contact and quantity of indirect outgroup friends)
-# HE_short (cis-heterosexuals w/o quantity of direct contact and quantity of indirect outgroup friends)
-# DG_short (ethnic minorities w/o quantity of direct contact and quantity of indirect outgroup friends)
-
-
-
-#... Save specification curve as PDF
-pdf <- F    # If FALSE, the plot is displayed in R. If TRUE, the plot is saved as a PDF in your working directory.
-
-
-#... Set seed for random number generation
-set.seed(999999) # set to 999999 to reproduce the results reported in the paper
-
-
-
[…]

### 2021-07-26 scripts/my_functions.R

@@ -1,0 +1,815 @@
+#..........................................................................................................#
+#..........................................................................................................#
+#................................ Specification Curve Analysis ............................................#
+#..........................................................................................................#
+#..........................................................................................................#
+
+#...# Prepare variables needed for specification curve analysis
+specVars <- function(whichData) {
+  varList <- list()
+  
+  # Full variable names / descriptions (variables)
+  varList[["variables"]] <- c("Working in solidarity",           
+                              "Support empowering policies",
+                              "Raising IG awareness",
+                              "High cost collective action",
+                              "Low cost collective action",
+                              "Quantity of contact",
+                              "Positive contact",
+                              "Absence of negative contact",
+                              "Number of OG friends",
+                              "Frequency of meeting OG friends",
+                              "Quantity of indirect OG friends",
+                              "Positive indirect contact",
+                              "Absence of negative indirect contact",
+                              "Exclusion of attention check failures",
+                              "Exclusion of statistical outliers")
+  
+  # Support for social change variables (dv)
+  varList[["dv"]] <- c("WRKSO","SEMP","IDISC", "WILAC_D","WILAC_E")
+  
+  # Intergroup contact variables (iv)
+  varList[["iv"]] <- c("CQUA", "CQUAL", "NEG_R", "NFRI","FFRI", "EXTC_CQUA", "EXTC_CQUAL", "EXTC_NEG_R") 
+  
+  # Other variables: Attention check (var.other)
+  varList[["var.other"]] <- c("ATTCHK")  
+  
+  # Adapt "variables" and "iv" for datasets without quantity of contact and quantity of indirect OG friends 
+  # (SM, HE_short, AG_short, DG_short)
[…]

### 2021-07-26 scripts/my_master_script.R

@@ -1,0 +1,493 @@
+# WARNING: The following line of code will remove all objects from the workspace. 
+#          Please consider saving your workspace.
+rm(list=ls()) 
+
+
+#................................# Set parameters #..................................#
+
+# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
+# Please read the comments in this section to understand what each parameter does.
+# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
+# sourcing the entire script or running the "Prepare R Session" section.
+
+
+#... Set number of analyses with shuffled data
+R <- 1
+# This determines how many times the original data will be shuffled in the permutation test
+# To reproduce our results, set R to 1000
+# For quick, approximate calculations, please set R to, e.g., 5 or 10
+
+
+
+#... Define which group of datasets to load
+# whichData <- "default"
+# Please choose one of the following options:
+# default: sub-sample mean differences removed (sample mean residualization)
+# keyDem: controlling for key demographic variables (SES ladder, age, gender). In the chosen datasets, these variables were partialled out from all contact and social change items.
+
+
+
+#... Define population for which specification curve analysis will be conducted
+# dataset <- "AG"
+# In the line above, please indicate for which population the specification curve analysis should be conducted
+# Possible options:
+# AG       (ethnic majorities)
+# HE       (cis-heterosexuals)
+# DG       (ethnic minorities)
+# SM       (LGBTIQ+ individuals)
+
+# AG_short (ethnic majorities w/o quantity of direct contact and quantity of indirect outgroup friends)
+# HE_short (cis-heterosexuals w/o quantity of direct contact and quantity of indirect outgroup friends)
+# DG_short (ethnic minorities w/o quantity of direct contact and quantity of indirect outgroup friends)
+
+
+
+#... Save specification curve as PDF
+pdf <- F    # If FALSE, the plot is displayed in R. If TRUE, the plot is saved as a PDF in your working directory.
+
+
+#... Set seed for random number generation
+set.seed(999999) # set to 999999 to reproduce the results reported in the paper
+
+
+
+#...........................................# Prepare R Session #...............................................#
+
[…]

### 2021-07-26 scripts/sca_script.R

@@ -1,0 +1,205 @@
+# WARNING: The following line of code will remove all objects from the workspace. 
+#          Please consider saving your workspace.
+rm(list=ls()) 
+
+
+#................................# Set parameters #..................................#
+
+# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
+# Please read the comments in this section to understand what each parameter does.
+# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
+# sourcing the entire script or running the "Prepare R Session" section.
+
+#... Set seed for random number generation
+set.seed(999999) # set to 999999 to reproduce the results reported in the paper
+
+
+#...........................................# Prepare R Session #...............................................#
+
+# WARNING: All currently loaded packages will be detached and packages needed 
+#          to run this script will be loaded. 
+#          Missing packages will be downloaded and installed.
+#          Also, the working directory will be automatically set to the location of this script.
+#          If you did not alter the structure of the folder containing this script and the data,
+#          in most cases you should be able to run this script without any manual adjustments 
+#          in a current version of RStudio (tested in RStudio Version 1.1.453 and newer releases running 
+#          R Version 3.5.1 and newer releases).
+
+
+# Detach packages
+if(!is.null(sessionInfo()$otherPkgs)) {
+  invisible(lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), 
+                   detach, character.only = TRUE, unload = TRUE))
+}
+
+# Download and install missing packages
+requiredPackages = c("psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
+                     "here", "data.table", "tidyverse", "lme4", "parallel", "pbmcapply", "PupillometryR")
+
+missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
+
+if(length(missingPackages)) {
+  install.packages(missingPackages)
+}
+
+# Load required packages
+invisible(lapply(requiredPackages, require, character.only = TRUE))
+
+# Source script containing functions required to run the master script
+# source(here("scripts", "my_functions.R"))
+
+# Load data
+dat = data.frame(fread(here("data", "3_prepared_data.csv")))
+
+
[…]

### 2021-07-26 scripts/sca_script.R
Adjusting viz

@@ -195,11 +195,9 @@ effects_df %>%
              linetype = "dashed", 
              color = "black") +
   
-  coord_flip() + theme_classic() + 
+  coord_flip() + theme_classic() + labs(y = "Skin tone odds ratio") +
   
   theme(legend.position = "none",
         axis.ticks.y = element_blank(),
         axis.title.y = element_blank(),
-        axis.text.y = element_blank()) +
-  
-  labs(y = "Skin tone odds ratio")
+        axis.text.y = element_blank())

### 2021-07-26 scripts/sca_script.R
Adjust cores used

@@ -106,7 +106,7 @@ ranef_sample =  sample(formula_ranef, n_sample, replace = FALSE)
 ef_sample = sample(formula_ef, n_sample, replace = FALSE)
 
 # preparing parallel processing
-num_cores = detectCores() - 2 # keep one core of other tasks
+num_cores = detectCores() - 2 # Not using all cores due to working memory limitations
 
 # Run models corresponding to their specification as glmer or glm. Parallel processing to decrease running time.
 all_models_glmer_results = pbmclapply(ranef_sample, mc.cores = num_cores, function(x) glmer(x, data = dat, family = "binomial", nAGQ = 0))

### 2021-07-26 scripts/sca_script.R
Refine functions

@@ -68,6 +68,9 @@ variables = c("Name", "Height (in cm)", "Weight (in kg)", "Age (in years)", "Spe
 # Dependent variables
 dv = c("all_reds", "red_cards", "yellow_red_cards", "yellow_cards")
 
+# "Base variable"
+base_var = c("skin_tone_num")
+
 # Independent variables
 iv = c("(1|player)", "height_cm", "weight_kg", "age_yrs", "specific_pos", "club", "league_country", "games", "goals", "victories", "(1|ref)", "ref_country", "ref_games", "imp_bias", "exp_bias")
 
@@ -78,19 +81,31 @@ var.all = c(dv, iv)
 varList = list()
 varList[["variables"]] = variables
 varList[["dv"]] = dv
+varList[["base_var"]] = base_var
 varList[["iv"]] = iv
 varList[["var.all"]] = var.all
 
 
 #.................. Prepare specifications: Create data frame from all combinations of covariates .........#
 # Create TRUE/FALSE matrix for every combination of covariates
-specifications = expand.grid(rep(list(0:1), length(iv))) == TRUE
+specifications = data.frame(expand.grid(rep(list(0:1), length(iv))) == TRUE)
 
 # Set column names
 colnames(specifications) = iv
 
-# Create a list included formula for every (standardised) covariate combination
-all_models = apply(specifications, 1, function(x) paste(c("all_reds ~ skin_tone_num", iv[x]), collapse = " + "))
+# Add dependent variable and "base variable"
+all_reds = c(rep(TRUE, nrow(specifications)))
+skin_tone_num = c(rep(TRUE, nrow(specifications)))
+
+specifications = add_column(specifications, all_reds, .before = iv[1])
+specifications = add_column(specifications, skin_tone_num, .before = iv[1])
+
+# Create a list included formula for every covariate combination
+base_model = paste(dv[1], "~", base_var)
+
+all_models = apply(specifications[3:ncol(specifications)], # starting from [, 3] as the first two columns being dv and base_var
+                   1, # loop through rows
+                   function(x) paste(c(base_model, iv[x]), collapse = " + ")) # create formulas
 
 # Identify random effects
 all_models_index = grepl("(1|", all_models, fixed = TRUE)
@@ -116,49 +131,50 @@ all_models_glm_results = pbmclapply(ef_sample, mc.cores = num_cores, function(x)
[…]

### 2021-07-26 scripts/my_functions.R
Ordering files

@@ -1,815 +1,0 @@
-#..........................................................................................................#
-#..........................................................................................................#
-#................................ Specification Curve Analysis ............................................#
-#..........................................................................................................#
-#..........................................................................................................#
-
-#...# Prepare variables needed for specification curve analysis
-specVars <- function(whichData) {
-  varList <- list()
-  
-  # Full variable names / descriptions (variables)
-  varList[["variables"]] <- c("Working in solidarity",           
-                              "Support empowering policies",
-                              "Raising IG awareness",
-                              "High cost collective action",
-                              "Low cost collective action",
-                              "Quantity of contact",
-                              "Positive contact",
-                              "Absence of negative contact",
-                              "Number of OG friends",
-                              "Frequency of meeting OG friends",
-                              "Quantity of indirect OG friends",
-                              "Positive indirect contact",
-                              "Absence of negative indirect contact",
-                              "Exclusion of attention check failures",
-                              "Exclusion of statistical outliers")
-  
-  # Support for social change variables (dv)
-  varList[["dv"]] <- c("WRKSO","SEMP","IDISC", "WILAC_D","WILAC_E")
-  
-  # Intergroup contact variables (iv)
-  varList[["iv"]] <- c("CQUA", "CQUAL", "NEG_R", "NFRI","FFRI", "EXTC_CQUA", "EXTC_CQUAL", "EXTC_NEG_R") 
-  
-  # Other variables: Attention check (var.other)
-  varList[["var.other"]] <- c("ATTCHK")  
-  
-  # Adapt "variables" and "iv" for datasets without quantity of contact and quantity of indirect OG friends 
-  # (SM, HE_short, AG_short, DG_short)
[…]

### 2021-07-26 scripts/my_master_script.R

@@ -1,493 +1,0 @@
-# WARNING: The following line of code will remove all objects from the workspace. 
-#          Please consider saving your workspace.
-rm(list=ls()) 
-
-
-#................................# Set parameters #..................................#
-
-# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
-# Please read the comments in this section to understand what each parameter does.
-# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
-# sourcing the entire script or running the "Prepare R Session" section.
-
-
-#... Set number of analyses with shuffled data
-R <- 1
-# This determines how many times the original data will be shuffled in the permutation test
-# To reproduce our results, set R to 1000
-# For quick, approximate calculations, please set R to, e.g., 5 or 10
-
-
-
-#... Define which group of datasets to load
-# whichData <- "default"
-# Please choose one of the following options:
-# default: sub-sample mean differences removed (sample mean residualization)
-# keyDem: controlling for key demographic variables (SES ladder, age, gender). In the chosen datasets, these variables were partialled out from all contact and social change items.
-
-
-
-#... Define population for which specification curve analysis will be conducted
-# dataset <- "AG"
-# In the line above, please indicate for which population the specification curve analysis should be conducted
-# Possible options:
-# AG       (ethnic majorities)
-# HE       (cis-heterosexuals)
-# DG       (ethnic minorities)
-# SM       (LGBTIQ+ individuals)
-
-# AG_short (ethnic majorities w/o quantity of direct contact and quantity of indirect outgroup friends)
-# HE_short (cis-heterosexuals w/o quantity of direct contact and quantity of indirect outgroup friends)
-# DG_short (ethnic minorities w/o quantity of direct contact and quantity of indirect outgroup friends)
-
-
-
-#... Save specification curve as PDF
-pdf <- F    # If FALSE, the plot is displayed in R. If TRUE, the plot is saved as a PDF in your working directory.
-
-
-#... Set seed for random number generation
-set.seed(999999) # set to 999999 to reproduce the results reported in the paper
-
-
-
-#...........................................# Prepare R Session #...............................................#
-
[…]

### 2021-07-26 scripts/4_sca_script.R
Update 4_sca_script.R

@@ -98,12 +98,13 @@ all_reds = c(rep(TRUE, nrow(specifications)))
 skin_tone_num = c(rep(TRUE, nrow(specifications)))
 
 specifications = add_column(specifications, all_reds, .before = iv[1])
-specifications = add_column(specifications, skin_tone_num, .before = iv[1])
+specifications = add_column(specifications, skin_tone_num, .after = "all_reds") # ".after" to make sure the previous column addition worked
 
-# Create a list included formula for every covariate combination
+# Create base model formula
 base_model = paste(dv[1], "~", base_var)
 
-all_models = apply(specifications[3:ncol(specifications)], # starting from [, 3] as the first two columns being dv and base_var
+# Create a list including a formula for every covariate combination
+all_models = apply(specifications[3:ncol(specifications)], # starting from [, 3] as the first two columns are dv and base_var
                    1, # loop through rows
                    function(x) paste(c(base_model, iv[x]), collapse = " + ")) # create formulas
 
@@ -114,16 +115,16 @@ all_models_index = grepl("(1|", all_models, fixed = TRUE)
 formula_ranef = all_models[all_models_index]
 formula_ef = all_models[!all_models_index]
 
-# Sample from all model specifications
+# Sample from all model specifications for a "minimum viable product" to work with
 n_sample = 10
 
 ranef_sample =  sample(formula_ranef, n_sample, replace = FALSE)
 ef_sample = sample(formula_ef, n_sample, replace = FALSE)
 
-# preparing parallel processing
+# Preparing parallel processing
 num_cores = detectCores() - 2 # Not using all cores due to working memory limitations
 
-# Run models corresponding to their specification as glmer or glm. Parallel processing to decrease running time.
+# Run models corresponding to their specification as glmer or glm. Parallel processing to decrease running time
 all_models_glmer_results = pbmclapply(ranef_sample, mc.cores = num_cores, function(x) glmer(x, data = dat, family = "binomial", nAGQ = 0))
 all_models_glm_results = pbmclapply(ef_sample, mc.cores = num_cores, function(x) glm(x, data = dat, family = "binomial"))
 

### 2021-07-28 scripts/3_data_preparation.R
Adjusting the scripts structure to use extractor function as wrapper

@@ -1,18 +1,24 @@
 library("here")
 library("data.table")
+library("tidyverse")
+library("plyr")
 library("lme4")
 
 # from raw data generate disaggregated data for logistic regression via 2_data_disaggregate.py
-dat = data.frame(fread(here("data", "2_disaggregated_data.csv")))
+dat = data.frame(fread(here::here("data", "2_disaggregated_data.csv")))
 
+# Silberzahn et al. 2018 list all covariates among them are two that need to be calculated: the cumulated number of cards received by a player and assigned by a referee.
+# The calculations of these covariates follows Morey and Wagenmarkers who were the team to use both covariates.
+dat = dat %>% group_by(player) %>% mutate(player_all_cards_received = sum(yellowCards + yellowReds + redCards))
+dat = dat %>% group_by(refNum) %>% mutate(ref_all_cards_assigned = sum(yellowCards + yellowReds + redCards))
 
 # select only the columns we will use in the analysis
 # adding all used covariates mentioned in Silberzahn et al. (2018) incl. main variables
 dat <- dat[, c("playerShort", "height", "weight", "birthday", "position",
                "club", "leagueCountry",
                "games",  "goals", "victories", "ties", 
-               "yellowCards", "yellowReds", "redCards", "allreds", "allredsStrict",
-               "refNum", "refCountry", "refCount",
+               "yellowCards", "yellowReds", "redCards", "allreds", "allredsStrict", "player_all_cards_received",
+               "refNum", "refCountry", "refCount", "ref_all_cards_assigned",
                "meanIAT", "meanExp", "skintone")]
 
 # if we're on windows or mac you need to force the bias variables into numeric
@@ -84,7 +90,7 @@ Finaldat <- data.frame(player = as.factor(dat$playerShort),
                        height_cm = as.numeric(dat$height),
                        weight_kg = as.numeric(dat$weight),
                        age_yrs = as.numeric(calc_age_years(dat$birthday, "%d.%m.%Y")),
-                       general_pos = as.factor(GeneralPos),
+                       # general_pos = as.factor(GeneralPos),
                        specific_pos = as.factor(dat$position),
                        club = as.factor(dat$club),
                        league_country = as.factor(dat$leagueCountry),
@@ -92,22 +98,24 @@ Finaldat <- data.frame(player = as.factor(dat$playerShort),
                        goals = as.numeric(dat$goals),
                        victories = as.numeric(dat$victories),
                        ties = as.numeric(dat$ties),
[…]

### 2021-07-28 scripts/4_sca_script.R

@@ -1,6 +1,6 @@
 # WARNING: The following line of code will remove all objects from the workspace. 
 #          Please consider saving your workspace.
-rm(list=ls()) 
+rm(list = ls()) 
 
 
 #................................# Set parameters #..................................#
@@ -32,9 +32,10 @@ if(!is.null(sessionInfo()$otherPkgs)) {
                    detach, character.only = TRUE, unload = TRUE))
 }
 
+
 # Download and install missing packages
 requiredPackages = c("psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
-                     "here", "data.table", "tidyverse", "lme4", "parallel", "pbmcapply", "PupillometryR")
+                     "here", "data.table", "tidyverse", "lme4", "parallel", "PupillometryR")
 
 missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
 
@@ -42,14 +43,16 @@ if(length(missingPackages)) {
   install.packages(missingPackages)
 }
 
+
 # Load required packages
 invisible(lapply(requiredPackages, require, character.only = TRUE))
+
 
 # Source script containing functions required to run the master script
 # source(here("scripts", "my_functions.R"))
 
 # Load data
-dat = data.frame(fread(here("data", "3_prepared_data.csv")))
+dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))
 
 
 #..........................................................................................................#
@@ -63,97 +66,95 @@ dat = data.frame(fread(here("data", "3_prepared_data.csv")))
 
 #...# Prepare variables needed for specification curve analysis
 # Full variable names
-variables = c("Name", "Height (in cm)", "Weight (in kg)", "Age (in years)", "Specific position", "Club", "League country", "Number of games", "Number of goals", "Number of victories", "Yellow cards", "Yellow/red cards", "Red cards", "all_reds",  "Referee ID", "Referee country", "Referee number of games", "Implicit bias", "Explicit bias")
+variables = c("All red cards",
+              "Skin tone (num)", "Implicit bias", "Explicit bias",
[…]

### 2021-07-28 scripts/4_sca_script.R
Adding data frame saving functionality

@@ -1,6 +1,6 @@
 # WARNING: The following line of code will remove all objects from the workspace. 
 #          Please consider saving your workspace.
-rm(list = ls()) 
+rm(list = ls())
 
 
 #................................# Set parameters #..................................#
@@ -96,6 +96,7 @@ varList[["co_var"]] = co_var
 varList[["var_all"]] = var_all
 
 
+
 #.................. Prepare specifications: Create data frame from all combinations of covariates .........#
 # Create TRUE/FALSE matrix for every combination of covariates
 specifications = data.frame(expand.grid(rep(list(0:1), length(co_var))) == TRUE)
@@ -165,7 +166,7 @@ extract_statistics = function(model, var_estimate) {
 
 
 # Sample from all model specifications for a "minimum viable product" to work with
-n_sample = 2
+n_sample = 100
 
 ranef_sample =  sample(formula_ranef, n_sample, replace = FALSE)
 ef_sample = sample(formula_ef, n_sample, replace = FALSE)
@@ -191,11 +192,16 @@ names(all_models_glmer_results) = ranef_sample
 names(all_models_glm_results) = ef_sample
 
 
-# Converting list to data frame while retaining item names
+# Converting list to data frame while retaining item names and combining to one data frame
 statistics_glmer = as.data.frame(do.call(rbind, all_models_glmer_results))
 statistics_glm = as.data.frame(do.call(rbind, all_models_glm_results))
 
 df = rbind(statistics_glmer, statistics_glm)
+
+
+# Converting row names to formula column and 
+df$formula = rownames(df)
+rownames(df) = NULL
 
 
 # Set significant level (alpha). Calculate confidence intervals (CI)
@@ -213,10 +219,12 @@ df$estimate_oddsratio = with(df, exp(estimate))
 df$ci_lower_oddsratio = with(df, exp(ci_lower))
 df$ci_upper_oddsratio = with(df, exp(ci_upper))
 
+write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
+
 
 #............................................# Rain cloud plot #...........................................#
 
-df %>%
+rain_cloud_plot = df %>%
   
   ggplot(aes(x = "", y = estimate_oddsratio)) +
   
@@ -247,3 +255,5 @@ df %>%
         axis.ticks.y = element_blank(),
         axis.title.y = element_blank(),
         axis.text.y = element_blank())
+
+# ggsave(here::here("figures", "rain_cloud_plot.png"), rain_cloud_plot, width = 10, height = 6)

### 2021-07-31 scripts/4_sca_script.R
Adding viz

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 
-#................................# Set parameters #..................................#
+#...............................................# Set parameters #..............................................#
 
 # This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
 # Please read the comments in this section to understand what each parameter does.
@@ -128,7 +128,7 @@ all_models = apply(specifications[ , (length(dv) + length(base_var) + 1):ncol(sp
                    function(x) paste(c(base_model, co_var[x]), collapse = " + ")) # create formulas
 
 
-# adding the formulas are a dedicated columns to the specification data frame
+# Adding the formulas are a dedicated columns to the specification data frame
 specifications$formula = all_models 
 
 
@@ -177,7 +177,7 @@ num_cores = detectCores() - 2 # Not using all cores due to working memory limita
 
 
 # Run models corresponding to their specification as glmer or glm. Parallel processing to decrease running time.
-# Calling summary to reduce the object size and save storage. Wrap model call within the above-define function.
+# Calling summary to generate a homogeneous outcome structure which needed for subsetting. Wrap model call within the above-define function.
 all_models_glmer_results = mclapply(ranef_sample,
                                     mc.cores = num_cores,
                                     function(x) extract_statistics(summary(glmer(x, data = dat, family = "binomial", nAGQ = 0)), "skin_tone_num"))
@@ -208,7 +208,7 @@ rownames(df) = NULL
 alpha_level = 0.05
 z = 1.96 # 95% CI
 
-df$below_alpha = with(df, ifelse(p_value < alpha_level, TRUE, FALSE))
+df$below_alpha = with(df, ifelse(p_value < alpha_level, "Significant", "Non-significant"))
 
 df$ci_lower = with(df, estimate - z*se)
 df$ci_upper = with(df, estimate + z*se)
@@ -224,36 +224,124 @@ write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
 
 #............................................# Rain cloud plot #...........................................#
 
-rain_cloud_plot = df %>%
+vibration_of_effect = df %>%
+  
[…]

### 2021-07-31 scripts/4_sca_script.R
Updating naming

@@ -293,7 +293,7 @@ impact_coef = data.frame(impact_summary$coefficients)[c(-1), 1]
 impact_se = data.frame(impact_summary$coefficients)[c(-1), 2]
 impact_pvalue = data.frame(impact_summary$coefficients)[c(-1), 4]
 
-impact_names = c("Position", "Height", "Weight", "League country", "Age", "Goals", "Club", "Ref country", "Ref", "Victories", "Cards received", "Player", "Cards assigned", "Ties", "Games")
+impact_names = c("Position", "Height", "Weight", "Lg. country", "Age", "Goals", "Club", "Ref country", "Ref", "Victories", "Cards rec.", "Player", "Cards assig.", "Ties", "Games")
 impact_df = data.frame(cbind(impact_names,
                              impact_coef,
                              impact_se,
@@ -337,8 +337,8 @@ covariate_effects = impact_df %>%
         legend.text = element_text(size = 11),
         legend.direction = "vertical",
         legend.title = element_blank(),
-        legend.margin = margin(l = 0.1, b = -0.75, unit = "cm"),
-        legend.box.margin = margin(l = 0.1, b = -0.75, unit = "cm"),
+        legend.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
+        legend.box.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
         plot.title = element_text(face = "bold"),
         plot.subtitle = element_text(face = "italic"))
 

### 2021-08-01 dissertation/msc_dissertation_introduction.Rmd
Adjusting the scripts structure to use extractor function as wrapper
Adding comments to multiverse script

@@
 ---
-title: "__How analysis strategy affects analysis results__"
-subtitle: "Vibration of effects due to model specifications"
+title: |
+  | __Assessing the median effect robustness in__
+  | __@silberzahn2018 through model specification__
+subtitle: "How analysis strategy affects analysis results"
 author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
@@
 # Introduction
+     The Covid-19 pandemic has reaffirmed the need for rigours scientific research to inform decision-making on a scientific and societal level.
+However, particularly the social sciences have a history of inflated positive findings (Ionanidis) leading to the reproducibility crisis [@pashler2012].
+@simmons2011 demonstrated how the so-called *researcher degree of freedom* affect study results.
+This concept refers to the analytical choices researchers make and their induced variation in results.
+For instance, sample size can be highly deterministic of finding statistically significant effects.
+@john2012 found 70% of researchers have at least once stopped their data collection based on a preliminary analysis.
+P-value simulations as a function of sample size showed an increased likelihood of obtaining false-positives for small sample sizes.
+Specifically, a sample size of ten had a 22.1% chance of being a false positive.
+Adding 20 observations per condition decreased the probability to 12.7% [@simmons2011].
+The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish.
+To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers, respectively.
 
@@
 
-The Covid-19 pandemic has reaffirmed the need for rigorous scientific research, but the process is not straightforward.
-In fact, Ioannidis 2005 argued that most study findings are false positives due to flawed study designs and insufficient sample sizes.
-
-@silberzahn2018
-
 @patel2015
-
-@haessler2020
 
@@
 
+# Results
+
+# Discussion
+
 \newpage

### 2021-08-01 scripts/4_multiverse_analysis.R

@@ -34,7 +34,7 @@ if(!is.null(sessionInfo()$otherPkgs)) {
 
 
 # Download and install missing packages
-requiredPackages = c("psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
+requiredPackages = c(# "psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
                      "here", "data.table", "tidyverse", "lme4", "parallel", "PupillometryR")
 
 missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
@@ -66,25 +66,25 @@ dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))
 
 #...# Prepare variables needed for specification curve analysis
 # Full variable names
-variables = c("All red cards",
-              "Skin tone (num)", "Implicit bias", "Explicit bias",
-              "Player position", "Player’s height (in cm)", "Player’s weight (in kg)", "Player’s league country", "Player’s age (in years)", "Goals scored by player", "Player’s club", "Referee’s country", "Referee", "Player’s number of victories", "Number of cards received by player", "Player", "Number of cards awarded by referee", "Number of draws", "Number games")
+variables = c("All red cards", # dv
+              "Skin tone (num)", "Implicit bias", "Explicit bias", # base var
+              "Player position", "Player’s height (in cm)", "Player’s weight (in kg)", "Player’s league country", "Player’s age (in years)", "Goals scored by player", "Player’s club", "Referee’s country", "Referee", "Player’s number of victories", "Number of cards received by player", "Player", "Number of cards awarded by referee", "Number of draws", "Number games") # covar
 
 
-# Dependent variables
+# Dependent variable
 dv = c("all_reds")
 
 
-# "Base variable"
+# Base variables
 base_var = c("skin_tone_num", "imp_bias", "exp_bias")
 
 
-# Independent variables
-co_var = c("as.factor(specific_pos)", "height_cm", "weight_kg", "as.factor(league_country)", "age_yrs", "goals", "as.factor(club)", "as.factor(ref_country)", "(1|ref)", "victories", "player_cards_received", "(1|player)", "ref_cards_assigned", "ties", "games")
+# Covariates
[…]

### 2021-08-01 scripts/4_multiverse_analysis.R
Adding comments and creating dedicated viz script

@@ -48,9 +48,6 @@ if(length(missingPackages)) {
 invisible(lapply(requiredPackages, require, character.only = TRUE))
 
 
-# Source script containing functions required to run the master script
-# source(here("scripts", "my_functions.R"))
-
 # Load data
 dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))
 
@@ -281,10 +278,12 @@ analysed_specifications = analysed_specifications[ , c(2:19, 28)] # select varia
 
 colnames(analysed_specifications) = c("skin_tone_num", "imp_bias", "exp_bias", "specific_pos", "height_cm", "weight_kg", "league_country", "age_yrs", "goals", "club", "ref_country", "ref", "victories", "player_cards_received", "player", "ref_cards_assigned", "ties", "games", "estimate_oddsratio")
 
+
 # Run linear model to get the specified effects of each covariate (included yes/no)
 impact_mod = lm(estimate_oddsratio ~ factor(specific_pos) + factor(height_cm) + factor(weight_kg) + factor(league_country) + factor(age_yrs) + factor(goals) + factor(club) + factor(ref_country) + factor(ref) + factor(victories) + factor(player_cards_received) + factor(player) + factor(ref_cards_assigned) + factor(ties) + factor(games),
                 data = analysed_specifications)
 impact_summary = summary(impact_mod)
+
 
 # Extract data from model and save as data frame
 impact_coef = data.frame(impact_summary$coefficients)[c(-1), 1]
@@ -305,6 +304,9 @@ impact_df$ci_lower_oddsratio = with(impact_df, exp(ci_lower))
 impact_df$ci_upper_oddsratio = with(impact_df, exp(ci_upper))
 
 impact_df$below_alpha = with(impact_df, ifelse(as.numeric(impact_pvalue) < 0.05, "Significant", "Non-significant"))
+
+write.csv(impact_df, here::here("data", "4_covariate_effect_data.csv"), row.names = FALSE)
+
 
 # Visualise the specified effect of covariates
 covariate_effects = impact_df %>%

### 2021-08-01 scripts/5_visualisation.R

@@ -1,0 +1,142 @@
+# WARNING: The following line of code will remove all objects from the workspace. 
+#          Please consider saving your workspace.
+rm(list = ls())
+
+
+#...............................................# Set parameters #..............................................#
+
+# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
+# Please read the comments in this section to understand what each parameter does.
+# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
+# sourcing the entire script or running the "Prepare R Session" section.
+
+#... Set seed for random number generation
+set.seed(999999) # set to 999999 to reproduce the results reported in the paper
+
+
+#...........................................# Prepare R Session #...............................................#
+
+# WARNING: All currently loaded packages will be detached and packages needed 
+#          to run this script will be loaded. 
+#          Missing packages will be downloaded and installed.
+#          Also, the working directory will be automatically set to the location of this script.
+#          If you did not alter the structure of the folder containing this script and the data,
+#          in most cases you should be able to run this script without any manual adjustments 
+#          in a current version of RStudio (tested in RStudio Version 1.1.453 and newer releases running 
+#          R Version 3.5.1 and newer releases).
+
+
+# Detach packages
+if(!is.null(sessionInfo()$otherPkgs)) {
+  invisible(lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), 
+                   detach, character.only = TRUE, unload = TRUE))
+}
+
+
+# Download and install missing packages
+requiredPackages = c(# "psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
+                     "here", "data.table", "tidyverse", "PupillometryR")
+
+missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
+
+if(length(missingPackages)) {
+  install.packages(missingPackages)
+}
+
+
+# Load required packages
+invisible(lapply(requiredPackages, require, character.only = TRUE))
+
+
+# Load data
[…]

### 2021-08-01 scripts/4_multiverse_analysis.R
Cleaning multiverse script

@@ -219,56 +219,6 @@ df$ci_upper_oddsratio = with(df, exp(ci_upper))
 write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
 
 
-#............................................# Rain cloud plot #...........................................#
-
-vibration_of_effect = df %>%
-  
-  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 1.5) %>% 
-  
-  ggplot(aes(x = "", y = estimate_oddsratio)) +
-  
-  geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
-  
-  geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
-             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.6, stroke = 0) +
-  
-  geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
-  
-  geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
-  
-  annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color="black") +
-  
-  scale_color_manual(values = c("red", "black")) +
-  
-  scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(1, 1.5), expand = c(0, 0)) +
-  
-  labs(title = "Vibration of effect due to covariate specification", subtitle = "Odds ratio (OR), n = 200") +
-  
-  coord_flip() + theme_classic() +
-  
-  theme(axis.line.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
-        legend.position = "top",
-        legend.justification = "left",
-        legend.title.align = 1,
-        legend.text = element_text(size = 11),
-        legend.direction = "vertical",
-        legend.title = element_blank(),
-        legend.margin = margin(b = -0.735, unit = "cm"),
-        legend.box.margin = margin(b = -0.735, unit = "cm"),
-        plot.title = element_text(face = "bold"),
-        plot.subtitle = element_text(face = "italic")) +
-  
[…]

### 2021-08-01 scripts/5_visualisation.R

@@ -53,9 +53,16 @@ dat_models = data.frame(fread(here::here("data", "4_model_outcome_data.csv")))
 dat_covariates = data.frame(fread(here::here("data", "4_covariate_effect_data.csv")))
 
 
+# Adjusting class for proper visualisation
+dat_models$below_alpha = factor(dat_models$below_alpha)
+
+dat_covariates$impact_names = factor(dat_covariates$impact_names)
+dat_covariates$below_alpha = factor(dat_covariates$below_alpha)
+
+
 #............................................# Rain cloud plot #...........................................#
 
-vibration_of_effect = df %>%
+vibration_of_effect = dat_models %>%
   
   filter(estimate_oddsratio > 0 & estimate_oddsratio <= 1.5) %>% 
   
@@ -105,7 +112,7 @@ ggsave(here::here("figures", "vibration_of_effect.png"), vibration_of_effect, wi
 
 #.........................................# Covariate Effects plot #.......................................#
 
-covariate_effects = impact_df %>%
+covariate_effects = dat_covariates %>%
   
   ggplot(aes(x = reorder(x = impact_names, X = estimate_oddsratio), estimate_oddsratio)) +
   

### 2021-08-01 scripts/4_multiverse_analysis.R
Adding the SCA plot

@@ -35,7 +35,7 @@ if(!is.null(sessionInfo()$otherPkgs)) {
 
 # Download and install missing packages
 requiredPackages = c(# "psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
-                     "here", "data.table", "tidyverse", "lme4", "parallel", "PupillometryR")
+                     "here", "data.table", "tidyverse", "lme4", "parallel", "PupillometryR", "cowplot")
 
 missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
 
@@ -224,9 +224,10 @@ write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
 # Build data frame that includes the specifications and the analysis results
 analysed_specifications = inner_join(specifications, df, by = "formula")
 analysed_specifications = analysed_specifications %>% filter(estimate_oddsratio < 2 & estimate_oddsratio <= 1.5) # Excluding the zero cases makes a big difference: excluded R^2 = 0.9734, included R^2 = 0.3852
-analysed_specifications = analysed_specifications[ , c(2:19, 28)] # select variable of interest i.e., covariates and dv
+analysed_specifications = analysed_specifications[ , c(2:19, 24:25, 28:ncol(analysed_specifications))] # select variable of interest i.e., covariates and dv
+analysed_specifications$below_alpha = with(analysed_specifications, ifelse(p_value < alpha_level, "Significant", "Non-significant"))
 
-colnames(analysed_specifications) = c("skin_tone_num", "imp_bias", "exp_bias", "specific_pos", "height_cm", "weight_kg", "league_country", "age_yrs", "goals", "club", "ref_country", "ref", "victories", "player_cards_received", "player", "ref_cards_assigned", "ties", "games", "estimate_oddsratio")
+colnames(analysed_specifications) = c("skin_tone_num", "imp_bias", "exp_bias", "specific_pos", "height_cm", "weight_kg", "league_country", "age_yrs", "goals", "club", "ref_country", "ref", "victories", "player_cards_received", "player", "ref_cards_assigned", "ties", "games", "p_value", "below_alpha", "estimate_oddsratio", "ci_lower_oddsratio", "ci_upper_oddsratio")
 
 
 # Run linear model to get the specified effects of each covariate (included yes/no)

### 2021-08-01 scripts/5_visualisation.R

@@ -77,7 +77,7 @@ vibration_of_effect = dat_models %>%
   
   geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
   
-  annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color="black") +
+  annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
   
   scale_color_manual(values = c("red", "black")) +
   
@@ -147,3 +147,67 @@ covariate_effects = dat_covariates %>%
 covariate_effects
 
 ggsave(here::here("figures", "covariate_effects.png"), covariate_effects, width = 11, height = 5)
+
+
+#.........................................# SCA plot #.......................................#
+
+analysed_specifications = analysed_specifications[order(analysed_specifications$estimate_oddsratio), ] # order by odds ratio, low to high
+analysed_specifications$id = factor(seq(1:nrow(analysed_specifications))) # assign identifier
+
+# Preparing specification matrix i.e., the bottom part of the sca plot
+prepare_long_df = analysed_specifications[ , c(4:18, 21, 24)] # Subset covariates and statistics
+prepare_long_df[prepare_long_df == FALSE] = NA # FALSE ro NA to better subset
+long_df = melt(setDT(prepare_long_df), id.vars = c("id"), variable.name = "estimate_oddsratio") # Wide to long format
+long_df = long_df[complete.cases(long_df), ] # Exclude NA cases
+long_df = long_df[!grepl("estimate_oddsratio", long_df$estimate_oddsratio), ] # Exclude outcome variable
+colnames(long_df) = c("id", "covariate", "x")
+long_df = long_df[ , c("id", "covariate")] # Exclude value column
+
+impact_df = impact_df[order(impact_df$estimate_oddsratio), ] # order covariate effects low to high
+
+long_df$covariate = factor(long_df$covariate,
+                           levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
[…]

### 2021-08-01 dissertation/msc_dissertation.Rmd
Cleaning the dissertation scripts

--- 2021-08-01 dissertation/msc_dissertation_introduction.Rmd
@@
 The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish.
-To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers, respectively.
+To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers.
 
- @silberzahn2018 investigated the effects of analytical choices on results.
+ Among these requirements and guidelines are a minimum sample size, reporting all collected variables and presenting results with and without covariates [@simmons2011].
+While these are certainly worthwhile suggestions they still struggle to cope with the full spectrum of the analytical possibilities.
+@silberzahn2018 investigated precisely the effects of analytical choices on results.
 Specifically, they had 29 teams of researchers investigate the same research question with the same dataset.
@@
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect, and the odds-ratios ranged from 0.89 to 2.93 (median = 1.31).
+Twenty teams found a statistically significant effect, and the odds ratios ranged from 0.89 to 2.93 (median = 1.31).
 The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results.
+Similar studies support these *researcher degrees of freedom* effects [@botvinik-nezer2020].
 
-This study puts (psychological) research in a delicate position, because if researchers with honest intentions come to such drastically different conclusions despite starting off on the identical dataset what does this mean for the rest of studies?
+ @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis strategies (also known as specification-curve analysis) to get a more accurate representation of the results space.
+This analytical strategy refers to conducting analyses for all "reasonable specifications" (or at least a large sample thereof).
+"Reasonably" is defined as statistically valid tests, scrutinising the research question and non-redundant[@simonsohn2020].
 

### 2021-08-01 dissertation/msc_dissertation.Rmd
Updating scripts

@@
 This analytical strategy refers to conducting analyses for all "reasonable specifications" (or at least a large sample thereof).
-"Reasonably" is defined as statistically valid tests, scrutinising the research question and non-redundant[@simonsohn2020].
+"Reasonably" is defined as statistically valid tests, scrutinising the research question and non-redundant [@simonsohn2020].
+The idea is to simulate a
+Vibration of effect @patel2015
 
-@patel2015
+The present study will build upon @silberzahn2018, @botvinik-nezer2020 and @patel2015. It has been established that results vary and
 

### 2021-08-01 scripts/5_visualisation.R

@@ -63,30 +63,19 @@ dat_covariates$below_alpha = factor(dat_covariates$below_alpha)
 #............................................# Rain cloud plot #...........................................#
 
 vibration_of_effect = dat_models %>%
-  
   filter(estimate_oddsratio > 0 & estimate_oddsratio <= 1.5) %>% 
-  
   ggplot(aes(x = "", y = estimate_oddsratio)) +
-  
   geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
-  
   geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
              position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.6, stroke = 0) +
-  
   geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
-  
   geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
-  
   annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
-  
   scale_color_manual(values = c("red", "black")) +
-  
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(1, 1.5), expand = c(0, 0)) +
-  
-  labs(title = "Vibration of effect due to covariate specification", subtitle = "Odds ratio (OR), n = 200") +
-  
-  coord_flip() + theme_classic() +
-  
+  labs(title = "Vibration of effect due to covariate specification", subtitle = "Odds ratio (OR), N = 200") +
+  coord_flip() +
+  theme_classic() +
   theme(axis.line.y = element_blank(),
         axis.ticks.y = element_blank(),
         axis.title.y = element_blank(),
@@ -98,11 +87,10 @@ vibration_of_effect = dat_models %>%
         legend.text = element_text(size = 11),
         legend.direction = "vertical",
         legend.title = element_blank(),
-        legend.margin = margin(b = -0.735, unit = "cm"),
-        legend.box.margin = margin(b = -0.735, unit = "cm"),
+        legend.margin = margin(b = -0.73, unit = "cm"),
+        legend.box.margin = margin(b = -0.73, unit = "cm"),
         plot.title = element_text(face = "bold"),
         plot.subtitle = element_text(face = "italic")) +
-  
[…]

### 2021-08-02 dissertation/msc_dissertation.Rmd
Minor viz adjustments and further thesis writing

--- 2021-08-01 dissertation/msc_dissertation_introduction.Rmd
@@
 However, particularly the social sciences have a history of inflated positive findings (Ionanidis) leading to the reproducibility crisis [@pashler2012].
-@simmons2011 demonstrated how the so-called *researcher degree of freedom* affect study results.
+@simmons2011 demonstrated how the so-called *researcher degree of freedom* (RDF) affect study results.
 This concept refers to the analytical choices researchers make and their induced variation in results.
@@
 The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish.
-To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers, respectively.
+This example shows how easy it is to engineer statistically significant effects.
+To prevent such bad practices @simmons2011 suggested requirements and guidelines for authors and reviewers.
 
- @silberzahn2018 investigated the effects of analytical choices on results.
-Specifically, they had 29 teams of researchers investigate the same research question with the same dataset.
+ Among these requirements and guidelines are a minimum sample size, reporting all collected variables and presenting results with and without covariates [@simmons2011].
+These are important suggestions, but they struggle to cope with the full spectrum of the analytical possibilities.
+@silberzahn2018 further scrutinised the effects of RDF.
+Specifically, they had 29 teams of researchers investigate the same research question based on the same dataset.
 Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g.
@@
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect, and the odds-ratios ranged from 0.89 to 2.93 (median = 1.31).
+Twenty teams found a statistically significant effect; the odds ratios ranged from 0.89 to 2.93 (median = 1.31).
 The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results.
+Similar studies support these RDF effects [@botvinik-nezer2020].
+These studies show that even extended efforts do not necessarily yield conclusive results.
+Moreover, this crowdsourcing approach is not practical.
 
-This study puts (psychological) research in a delicate position, because if researchers with honest intentions come to such drastically different conclusions despite starting off on the identical dataset what does this mean for the rest of studies?
+ @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis, or specification-curve analysis, as an alternative to crowdsourcing approaches.
+This approach requires a researcher, ideally a more than one, to conduct with every plausible analysis, aggregate and report it, and conduct an inferential test across all possibilities to come to a conclusion.
+"Plausible" refers to statistically valid tests appropriate to the research question and non-redundancy [@simonsohn2020].
+This approach is statistically more complex and computationally more intense than conventional analyses, but makes inherent strategy bias and noise transparent, and forces researchers to look at the entire "garden of forking paths" [@gelman2013].
+A good example of quantifying the extend of variation due to analytical choices was conducted by @patel2015.
+They tested the association of clinical, environmental and physiological variables with all-cause mortality.
+Specifically, they specified three model approaches (gaussian, cox and binomial) and 13 different covariates.
+Their algorithm ran every possible covariate combination (on top of the base variables) for each model which allowed the authors to assess the "vibration of effect" (VoE).
+They concluded that a larger VoE calls for more caution when interpreting observational associations.
 
-@patel2015
-
-# Methods
-
-# Results
-
-# Discussion
+ Taken together, statistically effects are easy to engineer through adjusting analysis parameters like sample size.
+Crowdsourcing analysis is a good strategy to understand the RDF induced variation but is impractical due to its substantial personnel, time and effort requirement.
+Alternatives like multiverse analysis present a promising, low-personnel, but computationally intense alternative to identify RDF induced variation.
+The present study will build upon @silberzahn2018, @botvinik-nezer2020 and @patel2015.
+The objective is to conduct a multiverse analysis based on the data used in @silberzahn2018 and identify the VoE i.e., the results space of the outcome variable.
+This will allow for a better understanding of how robust crowdsourced identified effects are.
+Due to time constraints this study will function as a prove-of-concept.
+Specifically, it will take the analytical approach which produced the median outcome and run it with every possible covariate combination.
+Similarly to @patel2015 this produces a distribution of the outcome effects and its variation provides insight into the @silberzahn2018 median effect robustness.
+The smaller the variation, the greater is the effect's robustness.
 

### 2021-08-02 scripts/5_visualisation.R

@@ -105,8 +105,8 @@ covariate_effects = dat_covariates %>%
   geom_point(aes(color = below_alpha)) +
   geom_hline(yintercept = 1) +
   geom_errorbar(aes(ymin = ci_lower_oddsratio, ymax = ci_upper_oddsratio, color = below_alpha), width = 0.1) +
-  scale_color_manual(values = c("black", "red")) +
-  scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.85, 1.1, 0.05)), limits = c(0.85, 1.1)) +
+  scale_color_manual(values = c("red", "black")) +
+  scale_y_continuous(name = "Odds ratio\n", breaks = c(seq(0.85, 1.1, 0.05)), limits = c(0.85, 1.1)) +
   labs(title = "Specified effect of covatiates", subtitle = "N = 200, 95% CI", x = "Covariate") +
   theme_classic() +
   theme(panel.grid.major.y = element_line(colour = "grey"),
@@ -121,7 +121,8 @@ covariate_effects = dat_covariates %>%
         legend.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
         legend.box.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
         plot.title = element_text(face = "bold"),
-        plot.subtitle = element_text(face = "italic"))
+        plot.subtitle = element_text(face = "italic")) +
+  guides(colour = guide_legend(reverse = TRUE))
 
 covariate_effects
 
@@ -133,7 +134,7 @@ ggsave(here::here("figures", "covariate_effects.png"), covariate_effects, width 
 analysed_specifications = analysed_specifications[order(analysed_specifications$estimate_oddsratio), ] # order by odds ratio, low to high
 analysed_specifications$id = factor(seq(1:nrow(analysed_specifications))) # assign identifier
 
-# Preparing specification matrix i.e., the bottom part of the sca plot
+# Preparing specification matrix i.e., the bottom part of the SCA plot
 prepare_long_df = analysed_specifications[ , c(4:18, 21, 24)] # Subset covariates and statistics
 prepare_long_df[prepare_long_df == FALSE] = NA # FALSE ro NA to better subset
 long_df = melt(setDT(prepare_long_df), id.vars = c("id"), variable.name = "estimate_oddsratio") # Wide to long format
@@ -165,6 +166,7 @@ top = analysed_specifications %>%
         axis.ticks.x = element_blank(),
[…]

### 2021-08-10 dissertation/msc_dissertation.Rmd
Update

--- 2021-08-01 dissertation/msc_dissertation.Rmd
@@
 title: |
-  | __Assessing the median effect robustness in__
-  | __@silberzahn2018 through model specification__
-subtitle: "How analysis strategy affects analysis results"
+  | __How analysis strategy affects analysis results__
+subtitle: |
+  | *Assessing the median effect robustness in @silberzahn2018 through model specification*
+  |
+  | Master's thesis - DRAFT
 author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
 linestretch: 1.5
-output: pdf_document
+output:
+    pdf_document:
+        includes:
+            in_header: header.tex
+        extra_dependencies: ["flafter"]
 bibliography: references.bib
 csl: apa.csl
+header-includes:
+  - \usepackage[ruled,vlined]{algorithm2e}
 ---
@@
 
+#...............................................# Set parameters #..............................................#
+
+# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
+# Please read the comments in this section to understand what each parameter does.
+# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
+# sourcing the entire script or running the "Prepare R Session" section.
+
+#... Set seed for random number generation
+set.seed(999999) # set to 999999 to reproduce the results reported in the paper
+
+#...........................................# Prepare R Session #...............................................#
+
+# WARNING: All currently loaded packages will be detached and packages needed 
+#          to run this script will be loaded. 
+#          Missing packages will be downloaded and installed.
+#          Also, the working directory will be automatically set to the location of this script.
+#          If you did not alter the structure of the folder containing this script and the data,
+#          in most cases you should be able to run this script without any manual adjustments 
+#          in a current version of RStudio (tested in RStudio Version 1.1.453 and newer releases running 
+#          R Version 3.5.1 and newer releases).
+
+# Detach packages
+if(!is.null(sessionInfo()$otherPkgs)) {
+  invisible(lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), 
+                   detach, character.only = TRUE, unload = TRUE))
+}
+
+# Download and install missing packages
+requiredPackages = c("here", "data.table", "tidyverse", "PupillometryR", "cowplot", "knitr", "kableExtra", "papeR", "DiagrammeR")
+
+missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
+
+if(length(missingPackages)) {
+  install.packages(missingPackages)
+}
+
+# Load required packages
+invisible(lapply(requiredPackages, require, character.only = TRUE))
+
+# Load data
+dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))
+
 ```
+
+```{=latex}
+
+% Trigger ToC creation in LaTeX
+\renewcommand{\baselinestretch}{1.5}\normalsize
+\tableofcontents
+\renewcommand{\baselinestretch}{1.5}\normalsize
+
+```
+
+... \newline
+*Abstract--in progress.* \newline
+... \newline
+
+\newpage
 
@@
      The Covid-19 pandemic has reaffirmed the need for rigours scientific research to inform decision-making on a scientific and societal level.
-However, particularly the social sciences have a history of inflated positive findings (Ionanidis) leading to the reproducibility crisis [@pashler2012].
-@simmons2011 demonstrated how the so-called *researcher degree of freedom* affect study results.
+However, particularly the social sciences have a history of inflated positive findings leading to the reproducibility crisis [@pashler2012].
+@simmons2011 demonstrated how the so-called *researcher degree of freedom* (RDF) affect study results.
 This concept refers to the analytical choices researchers make and their induced variation in results.
@@
 Adding 20 observations per condition decreased the probability to 12.7% [@simmons2011].
-The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant results to publish.
-To prevent such bad practices the authors suggest requirements and guidelines for authors and reviewers.
+The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant effects to publish.
+This example shows how easy it is to engineer these statistically significant effects.
+To prevent such bad practices @simmons2011 suggested requirements and guidelines for authors and reviewers.
 
  Among these requirements and guidelines are a minimum sample size, reporting all collected variables and presenting results with and without covariates [@simmons2011].
-While these are certainly worthwhile suggestions they still struggle to cope with the full spectrum of the analytical possibilities.
-@silberzahn2018 investigated precisely the effects of analytical choices on results.
-Specifically, they had 29 teams of researchers investigate the same research question with the same dataset.
+These are important suggestions but they struggle to cope with the full spectrum of the analytical possibilities.
+@silberzahn2018 further scrutinised the effects of RDF.
+Specifically, they had 29 teams of researchers investigate the effects of skin colour on the odds of being sent off the football field based on the same dataset.
 Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g.
@@
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect, and the odds ratios ranged from 0.89 to 2.93 (median = 1.31).
-The authors also controlled for prior believes, experience and peer rated analysis quality, none of them accounted for the variation of results.
-Similar studies support these *researcher degrees of freedom* effects [@botvinik-nezer2020].
+Twenty teams found a statistically significant effect; the odds ratios ranged from 0.89 to 2.93 (median = 1.31).
+The authors also controlled for researchers' prior believes and experience as well as peer rated analysis quality, none of them accounted for the variation of results.
+Other crowdsourced analysis projects have made similar observations [@botvinik-nezer2020].
+These studies are interesting for two reasons (a) from a statistical perspective they highlight the RDF effects, and (b) from a content perspective they provide the most robust insights into the scrutinised concepts.
+Nonetheless, crowdsourcing approaches are impractical as they require extensive personnel (N number of researchers) and time resources (years) and are therefore often not feasible analytical strategies [@silberzahn2018].
 
- @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis strategies (also known as specification-curve analysis) to get a more accurate representation of the results space.
-This analytical strategy refers to conducting analyses for all "reasonable specifications" (or at least a large sample thereof).
-"Reasonably" is defined as statistically valid tests, scrutinising the research question and non-redundant[@simonsohn2020].
+ @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis, or specification-curve analysis, as an alternative to crowdsourcing approaches.
+This approach requires a researcher, ideally a more than one, to conduct with all plausible analyses, aggregate the results and conduct an inferential test across all possibilities to come to a conclusion.
+"Plausible" refers to statistically valid, non-redundant, tests appropriate to the research question [@simonsohn2020].
+This approach is statistically more complex and computationally more intense than conventional analyses, but makes inherent strategy bias and noise transparent, and forces researchers to look at the entire "garden of forking paths" [@gelman2013].
+A good example of quantifying the extend of variation due to model specification was conducted by @patel2015.
+They specified three modelling approaches (gaussian, cox and binomial) and 13 covariates.
+Their algorithm ran every possible covariate combination for each modelling approach which allowed the authors to assess the "vibration of effect" (VoE).
+The VoE is another term for quantifying the extend of variation in the results.
+The measure allows to assess the effect's robustness.
+The smaller the VoE, the less the effect depends on the analytical choices and vice versa.
+The authors therefore called for more caution when interpreting observational associations when finding large VoE's.
 
-@patel2015
+ Taken together, statistically significant effects are easy to engineer through adjusting analysis parameters like sample size.
+Crowdsourcing analysis presents a suitable strategy to understand the RDF induced variation but is impractical due to its substantial personnel and time requirements.
+Alternatives like multiverse analysis present a promising, low-personnel, but computationally intense alternative to identify RDF induced variation.
+The present research project will build upon the crowdsourcing findings of @silberzahn2018.
+Specifically, I will take the median analysis and rerun it with all possible covariate combinations.
+The objective is to define and understand the results space of the analysis i.e., assess the robustness of the median effect (VoE).
+Based on these insights I aim at contributing to further understanding the value, limitations and mechanisms of the multiverse approach.
 
-# Methods
+# Analysis
+## Analysis plan
+     The project's objective is to define the results space of the @silberzahn2018 in order to draw inferences about the relationship between crowdsourcing and multiverse strategies.
+The results space refers to the numerical interval between the lowest and highest possible outcome.
+It is created by running every possible analytical strategy.
+@patel2015 refer to a related concept as the "Vibration of Effect" (VoE) and have developed a standardised approach to identifying this space.
+Essentially there are two factors: the analytical approach (i.e., the statistical model) and the covariates (or control variables).
+For @silberzahn2018 this means running every combination of 29 different analytical approaches and 15 different covariates used across all teams.
+These are $29 * 2^{15} = 950,272$ possibilities, which exceeds the scope of this research project.
+To get a more manageable number of possibilities of I will, therefore, focus on one analytical approach.
+In particular, on the approach that produced the median outcome of all analyses.
+The median outcome, being the middle number of any set of values, is a reasonable starting point in order to assess the VoE.
+Focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities.
+Due to computational limitations of my computer I will focus on a random sample of 1,000 thereof.
 
-# Results
+ In @silberzahn2018 the median outcome was produced by @team23 which will hereinafter be referred to as *team 23* due to it's designated team number in the original study.
+*Team 23* first transformed the data and then conducted a mixed-model logistic regression.
+To recreate the identical starting point for the analyses I will, first, apply the same transformations using the scripts provided by *team 23*.
+(The transformation process will be described in more detain the "data" section.) Second, I will create a matrix of all possible covariate combinations whereof a random sample of 1,000 is drawn.
+Third, because the original study investigated the relationship between football players' skin colour and the odds of being sent off the field, there will be a set of core variables included in all models to be able answer the original research question.
+Following on *team 23*, I will define two interaction terms as the model's core: "skin tone *X* implicit bias" and "skin tone *X* explicit bias".
+(The data set's variables will be described in more detail in the "data" section.) Forth, I will combine the core variables and the randomly drawn covariates as formulas to run the models.
+Fifth, after running the models I will visualise the outcomes as the conventional specification curve plot as well as a rain-cloud plot [@allen2021] and scatter plot.
+The aim here is develop easy to understand graphs that succinctly give insight into the results.
+Finally, I will discuss the implications for multiverse analysis.
 
+## Code
+     The code is based on @haessler2020 and @patel2015.
+The former provided the operational backbone of the code.
+The latter the idea for the algorithm.
+R's random number generator is first fixed to be able to reproduce the outcomes.
+Then, all the packages are detached and only the relevant ones are loaded to prevent other functions from interfering.
+If the required packages are not installed, the code will do so.
+Next, the prepared data is loaded using the "Here" package [@here] for operating systems independence and the "data.table" package for faster importing of the data [@datatable].
+(The data has been prepared by running the "disaggregate_v3.py" script of *team 23*; for more detail see the "data" section.) After having loaded the data the variables are defined as dependent, base (or core) variables and covariates.
+This includes defining the variables' types: numeric, factor or random effect.
+Next, a matrix is created established all possible combination of covariates (including base variables).
+Each row represents one combination, the columns represent the variables.
+The cells are set to TRUE or FALSE to indicate a variables presence or absence, respectively.
+Thus, the matrix structure looks as the following:
+
+$$\begin{bmatrix}
+\text{dv} & \text{base var}_{1} & \text{base var}_{2} & \text{base var}_{3} & \text{covar}_{1} & \text{covar}_{2} & \text{covar}_{...} & \text{covar}_{15} \\
+\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{FALSE} & \text{...} & \text{FALSE} \\
+\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{...} & \text{FALSE} \\
+\text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} \\
+\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{...} & \text{TRUE} \\
+\end{bmatrix}$$ \newline
+
+ Using the *apply* function, looping over each row, the variables are pasted together thereby creating the corresponding formulas which are then saved as a vector and are appended as column to the aforementioned matrix.
+
+ *Team 23* ran a mixed-model logistic regression analysis using the "lme4" package more specifically, its "glmer" function [@lme4].
+The function only works if a random structure is included in the formula.
+As all possible covariate combinations are explored, a random structure is not included in every formula.
+Therefore, before running the models, the formulas need to separated based on including a random structure or not.
+The random structures were defined at the beginning of the code using the following syntax: "(1|*covariate*)".
+Hence, this pattern is also used to identify random structures through the "grepl" function.
+For the non-random structure formulas the regular "glm" function is run.
+Both models, mixed-effect and "regular", are estimated using Maximum Likelihood estimation to ensure the outcomes are as comparable as possible.
+*Team 23* had also set the "nAGQ" parameter to zero (for the mixed-model) this sacrificed parameter estimation accuracy for speed.
+The same was done here to stay as close to the original analysis as possible.
+Finally, 1,000 formula samples were drawn.
+75% of all formulas include at least one random structure this proportion is reflected in the sample.
+Furthermore, the samples are drawn without replacement i.e., every formula is unique.
+
+ Running the models yielded another challenge.
+Not only were the running times per model at times above 30 minutes, storing the model summaries also required significant memory.
+For instance, testing the code with a minimal sample of 10 (random structure) formulas yielded an analysis output larger than 400MB.
+This does not scale well, hence, a wrapper function was build.
+It's purpose was to extract the estimates, standard errors, test-statistics and p-values for the variable of interest (in this case: skin tone).
+This effect was substantial.
+The final outcome object for 1,000 models was reduced to 0.3MB(!).
+
+ The models were run using the "lapply" function more specifically, the paralellised version thereof including progress bars ("pbmcapply", @pbmcapply).
+Parallisation was necessary in order to reduce running time.
+The standard R is set to use one core, only.
+Parallelising the processes to six cores decreased the running time by six-fold.
+The results of each model iteration was appended to a list.
+Hence, the outcomes were two lists, one from the mixed-models, and one from the "regular" models.
+These lists were transformed to data frames and merged.
+Moreover, the p-values was coded into significant and non-significant ($\alpha = 0.05$).
+The odds ratios as well as their confidence intervals were calculated (the same way *team 23* did it).
+The resulting data frame was save as a comma-separated values (CSV) file.
+
+\begin{algorithm}
+\caption{Pseudocode overview--in progress}
+\DontPrintSemicolon
+\SetAlgoLined
+\KwData{Prepared data based on team 23 script}
+\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} 
+$variables$ $\leftarrow$ Define dependent, base variables and covariates \\
+$specifications$ $\leftarrow$ Use $variables$ to create matrix containing all possible covariate combinations \\
+$formula$ $\leftarrow$ Paste $specifications$ by row and append as column to $specifications$ \\
+$formula.ranef; formula.ef$ $\leftarrow$ Separate formulas based on including a random-structure \\
+$samlple.ranef; sample.ef$ $\leftarrow$ Random, without replacement, sample from $formula.ranef; formula.ef$
+\BlankLine
+ initialization\;
+  \While{not at end of this document}{
+    read current\;
+    \eIf{understand}{
+      go to next section\;
+      current section becomes this one\;
+      }{
+      go back to the beginning of current section\;
+      }
+    }
+\KwResult{List containing model statistics}
+\end{algorithm}
+
+... \newline
+*Explaining why not adjusting for multiple comparison--in progress* \newline
+... \newline
+
+To better understand the specified effects of the covariates I ran an ANOVA where each covariate was coded into included yes/no.
+The above retrieved effects where used outcome measure.
+The results were stored and saved as a separate data frame.
+As a final step the graphs were created using the "tidyverse" (specifically, "ggplot", @tidyverse), "PupillometryR" [@PupillometryR] and "cowplot" [@cowplot] packages.
+\newline
+
+... \newline
+*Explaining how the graphs were created including necessary data transformation needed--in progress.* \newline
+... \newline
+
+The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [*GitHub repository*](https://github.com/sebastianplnr/msc_dissertation_project).
+
+## Data origin and preparation
+Outline:
+
+* The same dataset as in @silberzahn2018 is used.
+
+* What is the data about
+
+* Data origin i.e., how did silberzahn et al. (2018) come up with the data set
+
+* how was the data processed by team 23
+
+* summary statistics of the data set
+
+```{r, echo = FALSE}
+
+# Setting variable class
+dat$player = as.factor(dat$player)
+dat$specific_pos = as.factor(dat$specific_pos)
+dat$club = as.factor(dat$club)
+dat$league_country = as.factor(dat$league_country)
+dat$ref = as.factor(dat$ref)
+dat$ref_country = as.factor(dat$ref_country)
+
+# Building a summary table
+kable(papeR::summarise(dat,
+                       type = "numeric",
+                       variables = c("height_cm", "weight_kg", "age_yrs", "games", "goals", "victories", "ties", "all_reds", "player_cards_received",
+                                     "ref_cards_assigned", "skin_tone_num", "imp_bias", "exp_bias")),
+      format = "latex", booktabs = TRUE) %>%
+  kable_styling(latex_options = "scale_down")
+
+```
+
+\newpage
+## Outcome I: vibration of effect
+## Outcome II: specified covariate effects
+## Outcome III: specification curve
+
+\newpage
 # Discussion

### 2021-08-10 scripts/4_multiverse_analysis.R

@@ -1,8 +1,3 @@
-# WARNING: The following line of code will remove all objects from the workspace. 
-#          Please consider saving your workspace.
-rm(list = ls())
-
-
 #...............................................# Set parameters #..............................................#
 
 # This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
@@ -34,8 +29,7 @@ if(!is.null(sessionInfo()$otherPkgs)) {
 
 
 # Download and install missing packages
-requiredPackages = c(# "psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
-                     "here", "data.table", "tidyverse", "lme4", "parallel", "PupillometryR", "cowplot")
+requiredPackages = c("here", "data.table", "tidyverse", "lme4", "parallel", "pbmcapply", "PupillometryR", "cowplot")
 
 missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
 
@@ -138,6 +132,16 @@ formula_ranef = all_models[all_models_index]
 formula_ef = all_models[!all_models_index]
 
 
+# Sample from all model specifications accounting for the proportion of random effects to non-random effects in the formulas
+n_sample = 1000
+prop_ranef = length(all_models[all_models_index]) / (length(all_models[!all_models_index]) + length(all_models[all_models_index]))
+prop_ef = length(all_models[!all_models_index]) / (length(all_models[!all_models_index]) + length(all_models[all_models_index]))
+
+
+ranef_sample =  sample(formula_ranef, n_sample*prop_ranef, replace = FALSE)
+ef_sample = sample(formula_ef, n_sample*prop_ef, replace = FALSE)
+
+
 # Function extracting estimates, se, z-values and p-value from models.
 # Use as wrapper function around the models to extract relevant statistics only and reducing the output size.
 extract_statistics = function(model, var_estimate) {
@@ -162,26 +166,19 @@ extract_statistics = function(model, var_estimate) {
 }
 
 
-# Sample from all model specifications for a "minimum viable product" to work with
-n_sample = 100
-
-ranef_sample =  sample(formula_ranef, n_sample, replace = FALSE)
-ef_sample = sample(formula_ef, n_sample, replace = FALSE)
-
-
 # Preparing parallel processing
 num_cores = detectCores() - 2 # Not using all cores due to working memory limitations
 
 
 # Run models corresponding to their specification as glmer or glm. Parallel processing to decrease running time.
[…]

### 2021-08-10 scripts/5_visualisation.R

@@ -1,8 +1,3 @@
-# WARNING: The following line of code will remove all objects from the workspace. 
-#          Please consider saving your workspace.
-rm(list = ls())
-
-
 #...............................................# Set parameters #..............................................#
 
 # This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
@@ -63,17 +58,17 @@ dat_covariates$below_alpha = factor(dat_covariates$below_alpha)
 #............................................# Rain cloud plot #...........................................#
 
 vibration_of_effect = dat_models %>%
-  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 1.5) %>% 
+  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 4) %>% 
   ggplot(aes(x = "", y = estimate_oddsratio)) +
   geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
   geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
-             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.6, stroke = 0) +
+             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.5, stroke = 0) +
   geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
   geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
   annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
   scale_color_manual(values = c("red", "black")) +
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(1, 1.5), expand = c(0, 0)) +
-  labs(title = "Vibration of effect due to covariate specification", subtitle = "Odds ratio (OR), N = 200") +
+  labs(title = "Vibration of effect due to covariate specification", subtitle = "N = 1,000; Odds ratio (OR)") +
   coord_flip() +
   theme_classic() +
   theme(axis.line.y = element_blank(),
[…]

### 2021-08-12 dissertation/msc_dissertation.Rmd
Further writing

@@
 # Introduction
-     The Covid-19 pandemic has reaffirmed the need for rigours scientific research to inform decision-making on a scientific and societal level.
-However, particularly the social sciences have a history of inflated positive findings leading to the reproducibility crisis [@pashler2012].
-@simmons2011 demonstrated how the so-called *researcher degree of freedom* (RDF) affect study results.
-This concept refers to the analytical choices researchers make and their induced variation in results.
-For instance, sample size can be highly deterministic of finding statistically significant effects.
-@john2012 found 70% of researchers have at least once stopped their data collection based on a preliminary analysis.
-P-value simulations as a function of sample size showed an increased likelihood of obtaining false-positives for small sample sizes.
-Specifically, a sample size of ten had a 22.1% chance of being a false positive.
-Adding 20 observations per condition decreased the probability to 12.7% [@simmons2011].
-The authors did not insinuate any malicious intent per-se, but rather attributed the issue to an ambiguity of a "right way" to analyse data, and needing statistically significant effects to publish.
-This example shows how easy it is to engineer these statistically significant effects.
-To prevent such bad practices @simmons2011 suggested requirements and guidelines for authors and reviewers.
+     The Covid-19 pandemic has reaffirmed the need for sound scientific research to inform decision-making on a scientific and societal level [@collins2021].
+Rigorous research builds upon a systematic and well-reasoned approach to solving a research problem.
+Based on available literature, a falsifiable research question is defined and a corresponding hypothesis is developed.
+An appropriate study is then designed and conducted.
+To draw inferences from the gathered data, statistical models are developed and the effect of each variable is assessed.
+Every one of these steps is influenced by the decisions researchers make, which are known as *researcher degrees of freedom* [RDF, @simmons2011; @wicherts2016].
+Usually there are many feasible analytical strategies to answering a research question and none of them are inherently right or wrong  [@carp2012].
+This often creates uncertainty, for example, about what covariates to include and how to model them, which in turn leads to inconsistent findings [@patel2015; @ioannidis2008].
+Recently, efforts have been made to better understand the scope of variation induced by different analytical strategies.
 
- Among these requirements and guidelines are a minimum sample size, reporting all collected variables and presenting results with and without covariates [@simmons2011].
-These are important suggestions but they struggle to cope with the full spectrum of the analytical possibilities.
-@silberzahn2018 further scrutinised the effects of RDF.
-Specifically, they had 29 teams of researchers investigate the effects of skin colour on the odds of being sent off the football field based on the same dataset.
-Other than being part of the project there was no incentive for researchers to participate i.e., there was no ulterior motive for researcher to manipulate outcomes (e.g.
-wanting/needing to publish).
-The variation between the analytic approaches was substantial.
+ These efforts include a crowdsourcing approach to data analysis [@silberzahn2018; @botvinik-nezer2020].
+Its premise is to have a large number of researchers team up into to smaller, independent groups to investigate the same research question based on the same dataset.
+For instance, @silberzahn2018 had 29 teams investigate the effects of a football player's skin colour on the odds of being sent off the field.
+The variation between the analytical strategies was substantial.
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect; the odds ratios ranged from 0.89 to 2.93 (median = 1.31).
-The authors also controlled for researchers' prior believes and experience as well as peer rated analysis quality, none of them accounted for the variation of results.
-Other crowdsourced analysis projects have made similar observations [@botvinik-nezer2020].
-These studies are interesting for two reasons (a) from a statistical perspective they highlight the RDF effects, and (b) from a content perspective they provide the most robust insights into the scrutinised concepts.
-Nonetheless, crowdsourcing approaches are impractical as they require extensive personnel (N number of researchers) and time resources (years) and are therefore often not feasible analytical strategies [@silberzahn2018].
+Twenty teams found a statistically significant effect.
+The authors also controlled for researchers' prior believes and experience as well as peer-rated analysis quality, none of which accounted for the variation of results.
+@botvinik-nezer2020 made similar observations.
+Crowdsourcing data analysis excels at emphasising the substantial impact of different analytical strategies, but has a significant drawback.
+It is extremely time and personnel intensive, the two mentioned studies lasted between 2 to 3+ years, and included 61 and 180 analysts, respectively.
+Not to mention the organisational effort.
 
- @silberzahn2018 and @botvinik-nezer2020 suggest using multiverse analysis, or specification-curve analysis, as an alternative to crowdsourcing approaches.
-This approach requires a researcher, ideally a more than one, to conduct with all plausible analyses, aggregate the results and conduct an inferential test across all possibilities to come to a conclusion.
-"Plausible" refers to statistically valid, non-redundant, tests appropriate to the research question [@simonsohn2020].
-This approach is statistically more complex and computationally more intense than conventional analyses, but makes inherent strategy bias and noise transparent, and forces researchers to look at the entire "garden of forking paths" [@gelman2013].
-A good example of quantifying the extend of variation due to model specification was conducted by @patel2015.
-They specified three modelling approaches (gaussian, cox and binomial) and 13 covariates.
-Their algorithm ran every possible covariate combination for each modelling approach which allowed the authors to assess the "vibration of effect" (VoE).
-The VoE is another term for quantifying the extend of variation in the results.
-The measure allows to assess the effect's robustness.
-The smaller the VoE, the less the effect depends on the analytical choices and vice versa.
-The authors therefore called for more caution when interpreting observational associations when finding large VoE's.
+ @silberzahn2018 and @botvinik-nezer2020, therefore, suggest using multiverse analysis (also known as specification-curve analysis) as an alternative to crowdsourcing.
+This approach requires identifying and running all plausible analytical strategies.
+Plausible strategies are defined as statistically valid, non-redundant, tests appropriate to the research question.
+Their aggregated results are then used to make inferences about the research question.
+Despite being statistically more complex and computationally more intense, this approach has the advantages of only needing one, or ideally a couple, researchers.
+Moreover, a researcher's inherent strategy bias is neutralised and noise is made transparent [@simonsohn2020].
+Multiverse approaches are therefore well suited for assessing the scope of variation induced by different analytical strategies.
+For instance, @patel2015 assessed an effect's robustness by running every combination of covariates and modelling options.
+Their conclusion was the larger the variation of outcomes, the less robust is the effect, and the less it should therefore to be trusted.
+The authors termed this measure the *vibration of effect* (VoE).
 
- Taken together, statistically significant effects are easy to engineer through adjusting analysis parameters like sample size.
-Crowdsourcing analysis presents a suitable strategy to understand the RDF induced variation but is impractical due to its substantial personnel and time requirements.
-Alternatives like multiverse analysis present a promising, low-personnel, but computationally intense alternative to identify RDF induced variation.
-The present research project will build upon the crowdsourcing findings of @silberzahn2018.
-Specifically, I will take the median analysis and rerun it with all possible covariate combinations.
-The objective is to define and understand the results space of the analysis i.e., assess the robustness of the median effect (VoE).
+ Taken together, researchers make many analytical decisions throughout conduction a study which are know as *researcher degrees of freedom*.
+Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its substantial time and personnel requirements.
+Multiverse analysis present a promising, low-personnel, but computationally intense alternative to assessing the scope this variation.
+Despite observing substantial variation in crowdsourced data analysis projects little is known about the extend to which these approaches cover the full range of possible results.
+In this project, using methods of multiverse analysis I will define the results space of @silberzahn2018 to estimate study's representativeness.
 Based on these insights I aim at contributing to further understanding the value, limitations and mechanisms of the multiverse approach.
@@
 ## Analysis plan
-     The project's objective is to define the results space of the @silberzahn2018 in order to draw inferences about the relationship between crowdsourcing and multiverse strategies.
+     The project's objective is to define the results space of the @silberzahn2018 to the estimate study's representativeness.
 The results space refers to the numerical interval between the lowest and highest possible outcome.
 It is created by running every possible analytical strategy.
-@patel2015 refer to a related concept as the "Vibration of Effect" (VoE) and have developed a standardised approach to identifying this space.
-Essentially there are two factors: the analytical approach (i.e., the statistical model) and the covariates (or control variables).
-For @silberzahn2018 this means running every combination of 29 different analytical approaches and 15 different covariates used across all teams.
-These are $29 * 2^{15} = 950,272$ possibilities, which exceeds the scope of this research project.
-To get a more manageable number of possibilities of I will, therefore, focus on one analytical approach.
+Hence, it is closely related to assessing an effect's robustness.
+@patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*.
+I will therefore use this standardised approach to define the results space.
+The approach essentially comprises two parameters: the statistical models and the covariates (or control variables).
+In @silberzahn2018 the analysts used numerous different statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression.
+In total there were 29 different modelling approaches.
+Additionally, each team used a different set of covariates.
+Across all teams there 15 covariates used.
+This gives $2^n$ i.e., $2^{15} = 32,768$ possible combinations of covariates.
+Adding all modelling possibilities to the equation gives a total of $29 * 2^{15} = 950,272$ combinations to run.
+This number exceeds the project's scope, hence, it needs to be reduce to a more manageable count.
+I am, therefore, focusing on one modelling approach.
 In particular, on the approach that produced the median outcome of all analyses.
-The median outcome, being the middle number of any set of values, is a reasonable starting point in order to assess the VoE.
-Focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities.
-Due to computational limitations of my computer I will focus on a random sample of 1,000 thereof.
+The median, being the middle number of any set of values, is a reasonable starting point in order to estimate the results space.
+Nevertheless, even focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities.
+Due to computational limitations I will first run a sample of 200, followed by a sample of 1,000 combinations.
+If there's a substantial difference between the median and the spread I will run more models, if not, I will make the assumption that 1,000 reasonably estimates all ~33k combinations.
 
@@
 *Team 23* first transformed the data and then conducted a mixed-model logistic regression.
-To recreate the identical starting point for the analyses I will, first, apply the same transformations using the scripts provided by *team 23*.
-(The transformation process will be described in more detain the "data" section.) Second, I will create a matrix of all possible covariate combinations whereof a random sample of 1,000 is drawn.
-Third, because the original study investigated the relationship between football players' skin colour and the odds of being sent off the field, there will be a set of core variables included in all models to be able answer the original research question.
-Following on *team 23*, I will define two interaction terms as the model's core: "skin tone *X* implicit bias" and "skin tone *X* explicit bias".
-(The data set's variables will be described in more detail in the "data" section.) Forth, I will combine the core variables and the randomly drawn covariates as formulas to run the models.
-Fifth, after running the models I will visualise the outcomes as the conventional specification curve plot as well as a rain-cloud plot [@allen2021] and scatter plot.
-The aim here is develop easy to understand graphs that succinctly give insight into the results.
-Finally, I will discuss the implications for multiverse analysis.
+The first step of this project will be replicating the transformation and the analysis.
+Replicating other researchers analyses has the benefit of checking their work and, if results are indeed replicated, increasing confidence in them.
+It also ensures that this project has the same starting point as *team 23* did.
+Given the team made all their scripts publicly available I expect this step to be straightforward.
+Next is drawing a random sample of covariates, without replacement.
+"Without replacement" ensures all covariate combinations are unique in the sample.
+The covariates will be appended to the base (or core) variables.
+Base variables are those variables that are primarily assessed to answer the research question.
+In this case whether a football player's skin colour effects the odds of being sent off the field.
+*Team 23* defined two interaction terms as the base: "skin tone *X* implicit bias" and "skin tone *X* explicit bias".
+(The variables are described in the "data" section.) Hence, all models will have the following structure:
+$$
+\begin{aligned}
+RedCard = SkinTone*ImplicitBias + SkinTone*ExplicitBias + CovariateCombination_{i}
+\end{aligned}
+$$ 
+While running each model, the relevant statistics will be extracted.
+Those are the coefficient (or effect) of skin tone, its standard error, test-statistic and p-value.
+Just as *team 23* did I will then calculate the 95% confidence intervals (CI).
+The estimates and their CIs are then transformed to odds ratios (OR) through exponentiating them to the power of two.
+Odds ratios quantify the strength of association between two variables.
+If greater than one the dependent variable is more likely to occur given the independent variable, if lower than one it less likely.
+The odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot.
+The following describes them in more detail:
+
+(i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done.
+Its objective is describe the results space while allowing the reader to identify the specified covariate for each outcome [@simonsohn2020].
+It therefore has two vertically stacked components.
+The top part shows a sorted scatter plot.
+On the horizontal axis are the models specifications and on the vertical axis is the outcome measure.
+The points are sorted from lowest to highest outcome measure.
+This way the lowest outcome measure is on the bottom left corner and the highest in the top right one.
+The bottom part of the plot represents are table.
+Just like the top the horizontal axis hold the model specifications, the vertical axis lists the covariates.
+This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model.
+For the top and bottom plot to work together it is imperative that the horizontal axes are identically arranged.
+If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot.
+For an example check the results section.
+The issue, however, is that if too many specifications are on the horizontal axis it becomes very hard to identify specific covariates.
+Hence, I also included the rain-cloud and scatter plot.
+
+(ii) The rain-cloud plot was developed by @allen2021.
+Its objective is to give an unbiased, transparent view on the raw data.
+It therefore combines a density, box and scatter plot.
+If stacked vertically from top to bottom, respectively, they look like a cloud with rain drops, hence, its name.
+The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot.
+Given that this project seeks to define and describe a results space the rain-cloud plot the ideal graph to do this.
+However, the plot neither allows to identify the covariates giving rise to the outcome nor their specified effects.
+(For an example check the results section.) Therefore a third plot is developed which is to be assessed in tandem with this plot.
+
+(iii) The third and final plot is a sorted scatter plot.
+While there is nothing special about the plot itself its content is important.
+Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation.
+Instead the effects for each covariates neeed to be aggregated.
+Hence, the goal of this plot is to viusalise the specified effect for each covariate.
+Prior to doing so, the coviates first need to be aggregated and their specified effects need to be calucualted.
+To this end, each covariate is recoded into a binary factor: included in the model yes/no.
+This newly defined factors are than used as an independent variables in an ANOVA.
+The previously calcualted odds rations are used as dependent variables.
+The estimates of the fitted model are the specified effect for each covariates.
+These are visualised similarly to the top part of the specification curve plot.
+The covariates are on the horizental axis, will the odds ratios are on the vertical axis.
+Here too, the lowest oucome is in the bottom left corner and the highest outcome in the top right one.
+Additionally the point are coloured indicating a statistically significant effect (or non-signifcant).
 

### 2021-08-15 dissertation/msc_dissertation.Rmd
Cleaning up the code a bit

@@
 subtitle: |
-  | *Assessing the median effect robustness in @silberzahn2018 through model specification*
+  | *Defining the results space in @silberzahn2018 through model specification*
   |
@@
             in_header: header.tex
-        extra_dependencies: ["flafter"]
+        fig_caption: yes
+urlcolor: blue
 bibliography: references.bib
@@
 # Download and install missing packages
-requiredPackages = c("here", "data.table", "tidyverse", "PupillometryR", "cowplot", "knitr", "kableExtra", "papeR", "DiagrammeR")
+requiredPackages = c("here", "data.table", "tidyverse", "PupillometryR", "cowplot", "knitr", "DiagrammeR")
 
@@
 # Load data
+## prepared data
 dat = data.frame(fread(here::here("data", "3_prepared_data.csv")))
+
+## Outcomes of running the 1,000 models
+dat_models = data.frame(fread(here::here("data", "4_model_outcome_data.csv")))
+
+# 
+dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcome_data_specifications.csv")))
+
+## Outcome of running the "meta-ANOVA" to get the covariates' specified effects
+dat_covariates = data.frame(fread(here::here("data", "4_covariate_effect_data.csv")))
+
+# Adjusting class for proper visualisation
+dat_models$below_alpha = factor(dat_models$below_alpha)
+
+dat_covariates$impact_names = factor(dat_covariates$impact_names)
+dat_covariates$below_alpha = factor(dat_covariates$below_alpha)
 
@@
 Despite observing substantial variation in crowdsourced data analysis projects little is known about the extend to which these approaches cover the full range of possible results.
-In this project, using methods of multiverse analysis I will define the results space of @silberzahn2018 to estimate study's representativeness.
-Based on these insights I aim at contributing to further understanding the value, limitations and mechanisms of the multiverse approach.
+In this project, using principles of multiverse analysis I will define the results space of @silberzahn2018.
+By doing so I aim at further understanding how a specific analysis strategy affects its result.
+To my knowldge this is the first project runing not only all *plausible* but all *possible* analytical strategies to a more granular view of the results space.
 
@@
 ## Analysis plan
-     The project's objective is to define the results space of the @silberzahn2018 to the estimate study's representativeness.
+     The project's objective is to define the results space of the @silberzahn2018 to draw conclusions about the mechanisms of multiverse analysis.
 The results space refers to the numerical interval between the lowest and highest possible outcome.
@@
 Hence, it is closely related to assessing an effect's robustness.
+Robustness refers to an effect's consistency under different model specifications.
 @patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*.
@@
 Base variables are those variables that are primarily assessed to answer the research question.
-In this case whether a football player's skin colour effects the odds of being sent off the field.
+In this case whether a football player's skin colour affects the odds of being sent off the field.
 *Team 23* defined two interaction terms as the base: "skin tone *X* implicit bias" and "skin tone *X* explicit bias".
@@
 If greater than one the dependent variable is more likely to occur given the independent variable, if lower than one it less likely.
-The odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot.
-The following describes them in more detail:
+
+ These odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot. The following describes the three graphs in more detail:
 
 (i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done.
-Its objective is describe the results space while allowing the reader to identify the specified covariate for each outcome [@simonsohn2020].
+Its objective is to describe the results space while allowing the reader to identify the specified covariates for each outcome [@simonsohn2020].
 It therefore has two vertically stacked components.
 The top part shows a sorted scatter plot.
-On the horizontal axis are the models specifications and on the vertical axis is the outcome measure.
-The points are sorted from lowest to highest outcome measure.
+Its horizontal axis shows the models specifications and its vertical axis displays the outcome measure.
+The scatter points are sorted from lowest to highest outcome measure.
 This way the lowest outcome measure is on the bottom left corner and the highest in the top right one.
-The bottom part of the plot represents are table.
-Just like the top the horizontal axis hold the model specifications, the vertical axis lists the covariates.
+The bottom part of the plot represents a table.
+As in the top plot the horizontal axis holds the model specifications, the vertical axis lists the covariates.
 This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model.
@@
 If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot.
-For an example check the results section.
-The issue, however, is that if too many specifications are on the horizontal axis it becomes very hard to identify specific covariates.
-Hence, I also included the rain-cloud and scatter plot.
+For an example check Figure \ref{fig:sca_plot}.
+As the current project has 1,000+ specifications it becomes very hard to identify specific covariates given the space limitations.
+Hence, I to had come up with different more intuitive ways of visualisation.
 
@@
 The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot.
-Given that this project seeks to define and describe a results space the rain-cloud plot the ideal graph to do this.
-However, the plot neither allows to identify the covariates giving rise to the outcome nor their specified effects.
-(For an example check the results section.) Therefore a third plot is developed which is to be assessed in tandem with this plot.
+The current project seeks to define and describe a results space the rain-cloud plot is, therefore, the ideal choice of graphs.
+(For an example check the results section.) Unfortunately, the plot does not allow to identify the covariates giving rise to the outcomes or their specified effects.
+As the project also seeks to understand the mechanisms of multiverse analysis it important to get insights about the specified covariates effects.
+Therefore a third plot is developed which is to be assessed in tandem with this plot.
 
@@
 Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation.
-Instead the effects for each covariates neeed to be aggregated.
-Hence, the goal of this plot is to viusalise the specified effect for each covariate.
-Prior to doing so, the coviates first need to be aggregated and their specified effects need to be calucualted.
+Instead the effects for each covariates need to be aggregated.
+Hence, the goal of this plot is to visualise the specified effect for each covariate.
+Prior to doing so, the covariates first need to be aggregated and their specified effects need to be calculated.
 To this end, each covariate is recoded into a binary factor: included in the model yes/no.
 This newly defined factors are than used as an independent variables in an ANOVA.
-The previously calcualted odds rations are used as dependent variables.
+The previously calculated odds rations are used as dependent variables.
 The estimates of the fitted model are the specified effect for each covariates.
 These are visualised similarly to the top part of the specification curve plot.
-The covariates are on the horizental axis, will the odds ratios are on the vertical axis.
-Here too, the lowest oucome is in the bottom left corner and the highest outcome in the top right one.
-Additionally the point are coloured indicating a statistically significant effect (or non-signifcant).
+The covariates are on the horizontal axis, will the odds ratios are on the vertical axis.
+Here too, the lowest outcome is in the bottom left corner and the highest outcome in the top right one.
+Additionally the point are coloured indicating a statistically significant effect (or non-significant).
+
+ Taken together, the current project's objective is to define the results space of @silberzahn2018 in order to make inferences about the underlying mechanisms of multiverse analysis.
+The results space is created by running all possible model and covariate combinations.
+Applied to @silberzahn2018 this means running roughly 1M models.
+Due to time and computational limitations, I am focusing on the model that produced the median outcome of all analytical strategies.
+Hence, I will run a mixed-model logistic regression with 1,000 randomly sampled covariate combinations.
+For each model relevant statistics will extracted and visualised.
+The conventional specification curve plot does not deal well with a large number of model specifications.
+Therefore, I will produce a rain-cloud and scatter plot, which visualise the results space (and its distribution) as well as the specified covariate effects, respectively.
+For the latter, I will first have to run an additional statistical model to calculate the specified effects.
+These graphs will give insight into the results space, its building blocks and ultimately into the mechanisms of multiverse analysis.
 
 ## Code
-     The code is based on @haessler2020 and @patel2015.
-The former provided the operational backbone of the code.
-The latter the idea for the algorithm.
-R's random number generator is first fixed to be able to reproduce the outcomes.
-Then, all the packages are detached and only the relevant ones are loaded to prevent other functions from interfering.
-If the required packages are not installed, the code will do so.
-Next, the prepared data is loaded using the "Here" package [@here] for operating systems independence and the "data.table" package for faster importing of the data [@datatable].
-(The data has been prepared by running the "disaggregate_v3.py" script of *team 23*; for more detail see the "data" section.) After having loaded the data the variables are defined as dependent, base (or core) variables and covariates.
-This includes defining the variables' types: numeric, factor or random effect.
-Next, a matrix is created established all possible combination of covariates (including base variables).
-Each row represents one combination, the columns represent the variables.
-The cells are set to TRUE or FALSE to indicate a variables presence or absence, respectively.
-Thus, the matrix structure looks as the following:
+     This section provides a brief overview of the code and explains the reasoning behind it.
+As mentioned in the analysis section the first step was replicating the data transformation and analysis of *team 23* [@team23].
+The team made their project folder publicly available.
+It contains three scripts relevant to this project: the data exploration, transformation and analysis.
+After duplicating their project folder on my local hard drive all scripts run without issues (only the working directories needed adjustment).
+The data exploration included the reasoning behind transforming the data and some cleaning.
+The data transformation restructured the data into a more intuitive format (more details on the topics of exploration and transformation are in the data section).
+The analysis script prepared the data by assigning variables classes (factors, numerical or boolean) and standardised a few variables i.e., they were centred around the mean.
+Finally, the models were specified.
+The team ran both frequentist and Bayesian models, however, this project covers the frequentist approach, only.
 
-$$\begin{bmatrix}
-\text{dv} & \text{base var}_{1} & \text{base var}_{2} & \text{base var}_{3} & \text{covar}_{1} & \text{covar}_{2} & \text{covar}_{...} & \text{covar}_{15} \\
-\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{FALSE} & \text{...} & \text{FALSE} \\
-\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{FALSE} & \text{...} & \text{FALSE} \\
-\text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} & \text{...} \\
-\text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{TRUE} & \text{...} & \text{TRUE} \\
-\end{bmatrix}$$ \newline
+ My code is based on @haessler2020 and @patel2015.
+The rough structure of the code is shown in the pseudocode table (Figure X).
+Here, I am briefly motivating two analytical choices I made.
+For a closer look please check the commented code itself:
 
- Using the *apply* function, looping over each row, the variables are pasted together thereby creating the corresponding formulas which are then saved as a vector and are appended as column to the aforementioned matrix.
+(i) Choosing the function to run the statistical model.
+*Team 23* ran a mixed-model logistic regression.
+This model has two relevant components to this explanation: fixed effects and random effects.
+Fixed effects are those that are consistently observed in different situations because the construct is of direct interest to the research question.
+In this case, it is, for example, skin tone.
+Independent of the player, the game or the league skin tone is always observed.
+Their counter parts are random effects which change between situations.
+A player, for example, is just one "unit" a measurement, and is himself not directly relevant to answering the research question.
+The function used by *team 23* requires to specify a random effect.
+However, given the project seeks to sample from all possible covariate combinations, a random effect is not always included.
+Hence, the "non-random" counter part to this function needs to be utilised.
+For this project, it was important to check that both functions use the same estimation method (maximum likelihood or restricted maximum likelihood estimation) to calculate the statistics.
+It was important to check to ensure the outcomes are comparable.
 
- *Team 23* ran a mixed-model logistic regression analysis using the "lme4" package more specifically, its "glmer" function [@lme4].
-The function only works if a random structure is included in the formula.
-As all possible covariate combinations are explored, a random structure is not included in every formula.
-Therefore, before running the models, the formulas need to separated based on including a random structure or not.
-The random structures were defined at the beginning of the code using the following syntax: "(1|*covariate*)".
-Hence, this pattern is also used to identify random structures through the "grepl" function.
-For the non-random structure formulas the regular "glm" function is run.
-Both models, mixed-effect and "regular", are estimated using Maximum Likelihood estimation to ensure the outcomes are as comparable as possible.
-*Team 23* had also set the "nAGQ" parameter to zero (for the mixed-model) this sacrificed parameter estimation accuracy for speed.
-The same was done here to stay as close to the original analysis as possible.
-Finally, 1,000 formula samples were drawn.
-75% of all formulas include at least one random structure this proportion is reflected in the sample.
-Furthermore, the samples are drawn without replacement i.e., every formula is unique.
-
- Running the models yielded another challenge.
-Not only were the running times per model at times above 30 minutes, storing the model summaries also required significant memory.
-For instance, testing the code with a minimal sample of 10 (random structure) formulas yielded an analysis output larger than 400MB.
-This does not scale well, hence, a wrapper function was build.
-It's purpose was to extract the estimates, standard errors, test-statistics and p-values for the variable of interest (in this case: skin tone).
-This effect was substantial.
-The final outcome object for 1,000 models was reduced to 0.3MB(!).
-
- The models were run using the "lapply" function more specifically, the paralellised version thereof including progress bars ("pbmcapply", @pbmcapply).
-Parallisation was necessary in order to reduce running time.
-The standard R is set to use one core, only.
-Parallelising the processes to six cores decreased the running time by six-fold.
-The results of each model iteration was appended to a list.
-Hence, the outcomes were two lists, one from the mixed-models, and one from the "regular" models.
-These lists were transformed to data frames and merged.
-Moreover, the p-values was coded into significant and non-significant ($\alpha = 0.05$).
-The odds ratios as well as their confidence intervals were calculated (the same way *team 23* did it).
-The resulting data frame was save as a comma-separated values (CSV) file.
+(ii) No multiple comparison.
+This is based on the fact that in statistics null hypotheses are one rejected if the probability of finding a false positive is below 5% (known as the significance level, $\alpha$).
+The more statistical tests a run, the more likely it is that one of those tests is indeed a false positive.
+Hence, the are techniques that account of this issue.
+It is usually good practice to account for those "multiple comparisons".
+Here, I refrained from doing so though.
+This project seeks to simulate multiple researchers running different models.
+Therefore, for the the purpose of this project all models are run independently by individual researchers.
+Those would not have knowledge of the other analyses and, hence, would not account for them.
+To main the highest possible ecological validity, I therefore did not account of multiple comparisons.
 
@@
 
-... \newline
-*Explaining why not adjusting for multiple comparison--in progress* \newline
-... \newline
+ The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [GitHub repository](https://github.com/sebastianplnr/msc_dissertation_project).
+The code from *team 23* [@team23] can be retrieved from their [OSF repository](https://osf.io/akqt4/).
+Following R packages were used *here 1.0.1* [@here], *data.table 1.14.0* [@datatable], *tidyverse 1.3.1* [@tidyverse], *lme4 1.1-27.1* [@lme4], *pbmcapply 1.5.0* [@pbmcapply], *PupillometryR 0.0.3* [@PupillometryR] and *cowplot 1.1.1* [@cowplot].
 
-To better understand the specified effects of the covariates I ran an ANOVA where each covariate was coded into included yes/no.
-The above retrieved effects where used outcome measure.
-The results were stored and saved as a separate data frame.
-As a final step the graphs were created using the "tidyverse" (specifically, "ggplot", @tidyverse), "PupillometryR" [@PupillometryR] and "cowplot" [@cowplot] packages.
+## Data description and preparation
+     The data was retrieved from @silberzahn2018.
+It contains information about football players, their encounters with referees and the received cards (yellow, yellow/red and red).
+Moreover, a player's position, age, club, league country, victories, ties, defeats and goals.
+In addition, a skin tone rating based on two independent raters was included.
+The referees are numerically coded to protect their identify.
+The referees origin countries are also included as well as an implicit and explicit racism bias scores for their respective countries.
+The exploratory data analysis (EDA) of *team 23* showed that each row of the dataset represents a unique player-referee combination listing all their encounters as well as the couple's total number of received/assigned cards.
+The team states that it preferred a different data format.
+More specifically, a format where each player-referee encounter is reflected by one row.
+This way each encounter had a maximum of one red card.
+To achieve this format the data had to be "disaggregated" i.e., transformed.
+Further EDA showed that receiving a red card is highly unlikely, 0.008%, i.e., data is highly skewed.
+This property informed the teams decision to use a type of logistic regression (a statistical modelling technique equipped to deal with skewed/binary distributions).
+Finally, the team excluded all referee that did have at least 22 encounters.
+Their reasoning was a game includes (at least) 22 players, 11 players per team, if this were not the case referee most likely did not officiate in one for the four leagues of interest (England, Germany, Italy and Spain).
+This step excluded roughly 66% of the referees but retained 97.4% of the cases.
+The final dataset used for the current project contained 335,537 observations and 19 variables.
+Figure \ref{fig:var_plots} shows the properties of the variables relevant to current project's analysis \newline
+
+```{r var_plots, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "A) Showing the strong skew of receiving no red to receiving a red card.
+B) Showing most players had a lower skin tone rating i.e., most players were on the whiter side.
+C) Showing there is not much variation in implict racial scores.
+D) Showing more variation in explitic reacial bias scores.
+The latter two were centred around the mean, hence, zero does not mean there is no bias."}
+
+red_card_plot = dat %>% ggplot(aes(factor(all_reds))) +
+  geom_bar() +
+  xlab("Red card") +
+  scale_y_continuous(name = "Count", labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))
+
+skin_tone_num_plot = dat %>% ggplot(aes(skin_tone_num)) +
+  geom_density() +
+  scale_x_continuous(name = "Skin tone rating") +
+  scale_y_continuous(name = "Density", breaks = c(seq(0, 8, 2)), limits = c(0, 8)) +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))
+
+imp_bias_plot = dat %>% ggplot(aes(imp_bias)) + 
+  geom_density(adjust = 2) +
+  scale_x_continuous(name = "Implicit racial bias", breaks = c(seq(-2, 2, 1)), limits = c(-2, 2)) +
+  scale_y_continuous(name = "Density", breaks = c(seq(0, 40, 10)), limits = c(0, 40)) +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))
+
+exp_bias_plot = dat %>% ggplot(aes(exp_bias)) +
+  geom_density(adjust = 2) + 
+  scale_x_continuous(name = "Explicit racial bias", breaks = c(seq(-2, 2, 1)), limits = c(-2, 2)) +
+  scale_y_continuous(name = "Density", breaks = c(seq(0, 40, 10)), limits = c(0, 40)) +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"), text = element_text(size = 10))
+
+plot_grid(red_card_plot, skin_tone_num_plot, imp_bias_plot, exp_bias_plot, labels = "AUTO")
+
+```
+
+# Results
+## Replicating *Team 23* analysis
+     As mentioned in the above section the first steps were replicating the *team 23* analysis using their provided scripts.
+The results are reproducible without any adjustment.
+Skin tone has a significant effect on the odds of being send off the field ($OR = 1.311$, 95%CI [1.099, 1.563], $p = 0.003$).
+This means while keeping all other variables constant for every unit increase in skin tone rating (darker skin tone), the odds of being send off the field increase by roughly 131%.
+The interaction terms of skin tone and implicit racial bias ($OR = 0.004$, 95%CI [0.000, 23.259], $p = 0.211$) as well as skin tone and explicit bias ($OR = 1.837$, 95%CI [0.493, 6.848], $p = 0.365$) are both non-significant.
+In accordance with the original analysis, there are also significant differences between leagues and positions as well as implicit racial bias scores.
+Explicit bias scores are on verge of being non-significant.
+This analysis only aimed at testing reproducibility, which proved to be true, further results details are forfeited.
+
+## Specification curve
+     Shown in Figure \ref{fig:sca_plot}.
+The horizontal axes of the top and bottom chart show the specifications sorted based on their odds rations from lowest to highest.
+The odds ratios themselves are shown on the vertical axis of the top plot.
+The points in the top plot, which look like a line, each represent the outcome of one statistical model.
+Black points refers to statistical significant, red to non-significant outcomes.
+Overall are 89.7% of the outcomes significant.
+Except for the lowest outcome ($OR_{lowest} = 1.080$) are all points close to the their respective neighbours.
+This indicates that none of the specified models are responsible for a sudden increase in the outcome measure.
+The relatively smooth line suggest that sampling 1,000 specifications is, in fact, representative for all 33k models.
+The law of large numbers  The points can be divided into roughly three sections all non-significant ($\leq 57^{th}$), mixture of (non-)significance ($58^{th} - 247^{th}$) and all significance ($\geq 248^{th}$).
+The bottom table lists all covariates on its vertical axis, sorted from most positive to most negative impact.
+For example, player has the strongest positive impact, while club has the strongest negative impact.
+Each column of the table represents one mode.
+A coloured cell indicates the presence of the variable, an uncoloured cell its absence.
+Just like a above, black indicates the outcome measure was significant, red non-significance.
+
+ The top and bottom plot work in tandem.
+Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table.
+Regular specification curve analyses focus on all *reasonable* specifications, hence, their number of specifications is much lower and one is, in fact, able to tell the different model specifications apart.
+However, the current project's goal was to define the results space, which requires all *possible* specifications.
+The specifications are therefore huddled extremely closely which does not allow the plot to functions in its supposed way.
+Thus, I create two alternative graphs.
 \newline
 
-... \newline
-*Explaining how the graphs were created including necessary data transformation needed--in progress.* \newline
-... \newline
+```{r sca_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Convential specification curve.
+Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table.
+Due to the large number of specifications, this plot is not as informative as it is supposed to be.
+It is, therefore, merely making an argument for needing and alternative.", warning = FALSE}
 
-The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [*GitHub repository*](https://github.com/sebastianplnr/msc_dissertation_project).
+# Order by odds ratio, low to high
+analysed_specifications = dat_models_specifications[order(dat_models_specifications$estimate_oddsratio), ]
 
-## Data origin and preparation
-Outline:
+# Assign identifier
+analysed_specifications$id = factor(seq(1:nrow(analysed_specifications)))
 
-* The same dataset as in @silberzahn2018 is used.
+# Preparing specification matrix i.e., the bottom part of the SCA plot
+## Subset covariates and statistics
+prepare_long_df = analysed_specifications[ , c(4:18, 20, 21, 24)]
 
-* What is the data about
+## FALSE ro NA to better subset
+prepare_long_df[prepare_long_df == FALSE] = NA
 
-* Data origin i.e., how did silberzahn et al. (2018) come up with the data set
+## Wide to long format
+long_df = melt(setDT(prepare_long_df), id.vars = c("id", "below_alpha"), variable.name = c("estimate_oddsratio"))
 
-* how was the data processed by team 23
+## Exclude NA cases
+long_df = long_df[complete.cases(long_df), ]
 
-* summary statistics of the data set
+## Exclude outcome variable
+long_df = long_df[!grepl("estimate_oddsratio", long_df$estimate_oddsratio), ]
+colnames(long_df) = c("id", "below_alpha", "covariate", "x")
 
-```{r, echo = FALSE}
+## Exclude value column
+long_df = long_df[ , c("id", "below_alpha", "covariate")]
 
-# Setting variable class
-dat$player = as.factor(dat$player)
-dat$specific_pos = as.factor(dat$specific_pos)
-dat$club = as.factor(dat$club)
-dat$league_country = as.factor(dat$league_country)
-dat$ref = as.factor(dat$ref)
-dat$ref_country = as.factor(dat$ref_country)
+## order covariate effects low to high, based on the 4_covariate_effect_data.csv dataset
+impact_df = dat_covariates[order(dat_covariates$estimate_oddsratio), ]
 
-# Building a summary table
-kable(papeR::summarise(dat,
-                       type = "numeric",
-                       variables = c("height_cm", "weight_kg", "age_yrs", "games", "goals", "victories", "ties", "all_reds", "player_cards_received",
-                                     "ref_cards_assigned", "skin_tone_num", "imp_bias", "exp_bias")),
-      format = "latex", booktabs = TRUE) %>%
-  kable_styling(latex_options = "scale_down")
+long_df$covariate = factor(long_df$covariate,
+                           levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
+                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Lg.
+country", "Position", "Age", "Ref", "Cards assig.", "Goals", "Height", "Games", "Ref country", "Player"))
+
+# build plot
+top = analysed_specifications %>% 
+  ggplot(aes(as.numeric(id), estimate_oddsratio)) +
+  geom_point(aes(colour = below_alpha), size = 0.3, alpha = 0.5) +
+  geom_ribbon(aes(ymin = ci_lower_oddsratio, ymax = ci_upper_oddsratio), alpha = 0.3) +
+  geom_hline(yintercept = 1, color = "black") +
+  scale_color_manual(values = c("red", "black")) +
+  scale_x_discrete(name = "", expand = c(0.01, 0)) +
+  scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
+  labs(title = "Specification curve", subtitle = "N = 1,000; 95% CI") +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"),
+        axis.ticks.x = element_blank(),
+        axis.text.x = element_blank(),
+        axis.line.x = element_blank(),
+        plot.margin = unit(c(0.5, 1, 0, 1), "cm"),
+        legend.position = "top",
+        legend.title.align = 1,
+        legend.text = element_text(size = 11),
+        legend.direction = "horizontal",
+        legend.title = element_blank(),
+        legend.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
+        plot.title = element_text(face = "bold"),
+        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
+        text = element_text(size = 10)) +
+  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))
+
+bottom = long_df %>% 
+  filter(id != "1") %>%
+  ggplot(aes(x = id, y = covariate)) +
+  geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
+  scale_fill_manual(values = c("red", "black")) +
+  scale_x_discrete(name = "Specifications", expand = c(0.01, 0)) +
+  scale_y_discrete(name = "Covariates", expand = c(0.03, 0)) +
+  theme_classic() +
+  theme(legend.position = "none", 
+        axis.ticks.y = element_blank(),
+        axis.line.y = element_blank(),
+        axis.ticks.x = element_blank(),
+        axis.text.x = element_blank(),
+        plot.margin = unit(c(0, 1, 0.5, 1), "cm"),
+        text = element_text(size = 10))
+
+sca_plot = plot_grid(top, bottom, ncol = 1, align = "v")
+sca_plot
+
+```
+
+## Results space
+     Shown in Figure \ref{fig:rain_cloud_plot}.
+The horizontal axis shows the odds rations.
+From top to bottom the three component show the probability density distribution, the box plot and the raw data, respectively.
+The dashed line indicates the original results of *team 23*, which has also been the median outcome of all analytical strategies in @silberzahn2018.
+Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383.
+The 1rd quartile   \newline
+
+```{r rain_cloud_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "The rain cloud plot is comprised of three parts.
+The probability density distribution, the box plot and the raw data points.
+This allows for an transparent and bias assessment of the results space and its distribution."}
+
+vibration_of_effect = dat_models %>%
+  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 4) %>% 
+  ggplot(aes(x = "", y = estimate_oddsratio)) +
+  geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
+  geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
+             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.3, stroke = 0) +
+  geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
+  geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
+  annotate("text", x = 1.516, y = 1.375, label = "Silberzahn et al. (2018) \nMedian OR", color = "black") +
+  scale_color_manual(values = c("red", "black")) +
+  scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
+  labs(title = "Results space defined through covariate specification", subtitle = "N = 1,000; Odds ratio (OR)") +
+  coord_flip() +
+  theme_classic() +
+  theme(axis.line.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
+        legend.position = "top",
+        legend.justification = "left",
+        legend.title.align = 1,
+        legend.text = element_text(size = 11),
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.margin = margin(b = -0.8, unit = "cm"),
+        legend.box.margin = margin(b = -0.8, unit = "cm"),
+        plot.title = element_text(face = "bold"),
+        plot.subtitle = element_text(face = "italic"),
+        text = element_text(size = 10)) +
+  guides(colour = guide_legend(override.aes = list(alpha = 1), reverse = TRUE))
+
+vibration_of_effect
+
+```
+
+## Specified covariate effects
+     \newline
+
+```{r covar_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "The specified covariate effects indicate the extend to which a covariate influences the outcome measures."}
+
+dat_covariates = dat_covariates[order(dat_covariates$estimate_oddsratio), ]
+
+covariate_effects = dat_covariates %>%
+  ggplot(aes(x = reorder(x = impact_names, X = impact_coef), impact_coef)) +
+  geom_point(aes(color = below_alpha)) +
+  geom_hline(yintercept = 0) +
+  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper, color = below_alpha), width = 0.1) +
+  scale_color_manual(values = c("red", "black")) +
+  scale_y_continuous(name = "Estimates", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
+  labs(title = "Specified covariate effects", subtitle = "N = 1,000; 95% CI", x = "Covariates") +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey"),
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
+        legend.position = "top",
+        legend.justification = "left",
+        legend.title.align = 1,
+        legend.text = element_text(size = 11),
+        legend.direction = "vertical",
+        legend.title = element_blank(),
+        legend.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
+        legend.box.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
+        plot.title = element_text(face = "bold"),
+        plot.subtitle = element_text(face = "italic"),
+        text = element_text(size = 10)) +
+  guides(colour = guide_legend(reverse = TRUE))
+
+covariate_effects
 
@@
 \newpage
-## Outcome I: vibration of effect
-## Outcome II: specified covariate effects
-## Outcome III: specification curve
 
-\newpage
 # Discussion

### 2021-08-15 scripts/4_multiverse_analysis.R

@@ -190,6 +190,7 @@ names(all_models_glm_results) = ef_sample
 statistics_glmer = data.frame(do.call(rbind, all_models_glmer_results))
 statistics_glm = data.frame(do.call(rbind, all_models_glm_results))
 
+
 # Combining to one data frame
 df = rbind(statistics_glmer, statistics_glm)
 
@@ -214,7 +215,50 @@ df$estimate_oddsratio = with(df, exp(estimate))
 df$ci_lower_oddsratio = with(df, exp(ci_lower))
 df$ci_upper_oddsratio = with(df, exp(ci_upper))
 
+
+# Calculate proportion of significant effects
+sum(df$below_alpha == "Significant") / nrow(df) # = 89.7%
+
+
 write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
+
+
+#............................................# Rain cloud plot #...........................................#
+
+vibration_of_effect = df %>%
+  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 4) %>% 
+  ggplot(aes(x = "", y = estimate_oddsratio)) +
+  geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
+  geom_point(aes(x = 0.6, y = estimate_oddsratio, colour = below_alpha),
+             position = position_jitter(width = .03, seed = 123), size = 3, shape = 20, alpha = 0.5, stroke = 0) +
+  geom_boxplot(aes(x = 0.81, y = estimate_oddsratio), alpha = 0.5, width = 0.1, colour = "black") +
+  geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
+  annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
+  scale_color_manual(values = c("red", "black")) +
+  scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
+  labs(title = "Vibration of effect due to covariate specification", subtitle = "N = 1,000; Odds ratio (OR)") +
+  coord_flip() +
+  theme_classic() +
+  theme(axis.line.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
[…]

### 2021-08-15 scripts/5_visualisation.R

@@ -1,224 +1,0 @@
-#...............................................# Set parameters #..............................................#
-
-# This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
-# Please read the comments in this section to understand what each parameter does.
-# After setting the parameters for the first time, we recommend reading the WARNINGS on lines 1 and 56 before
-# sourcing the entire script or running the "Prepare R Session" section.
-
-#... Set seed for random number generation
-set.seed(999999) # set to 999999 to reproduce the results reported in the paper
-
-
-#...........................................# Prepare R Session #...............................................#
-
-# WARNING: All currently loaded packages will be detached and packages needed 
-#          to run this script will be loaded. 
-#          Missing packages will be downloaded and installed.
-#          Also, the working directory will be automatically set to the location of this script.
-#          If you did not alter the structure of the folder containing this script and the data,
-#          in most cases you should be able to run this script without any manual adjustments 
-#          in a current version of RStudio (tested in RStudio Version 1.1.453 and newer releases running 
-#          R Version 3.5.1 and newer releases).
-
-
-# Detach packages
-if(!is.null(sessionInfo()$otherPkgs)) {
-  invisible(lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), 
-                   detach, character.only = TRUE, unload = TRUE))
-}
-
-
-# Download and install missing packages
-requiredPackages = c(# "psych", "plotrix", "rstudioapi", "DescTools", # packages needed?
-                     "here", "data.table", "tidyverse", "PupillometryR")
-
-missingPackages = requiredPackages[!requiredPackages %in% installed.packages()[ , "Package"]]
-
-if(length(missingPackages)) {
-  install.packages(missingPackages)
-}
-
-
-# Load required packages
-invisible(lapply(requiredPackages, require, character.only = TRUE))
-
-
-# Load data
[…]

### 2021-08-15 dissertation/msc_dissertation.Rmd
Cleaning up the code

@@
 ## Outcomes of running the 1,000 models
-dat_models = data.frame(fread(here::here("data", "4_model_outcome_data.csv")))
-
-# 
-dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcome_data_specifications.csv")))
+dat_models_specifications = data.frame(fread(here::here("data", "4_model_outcomes_specifications_merged.csv")))
 
@@
 # Adjusting class for proper visualisation
-dat_models$below_alpha = factor(dat_models$below_alpha)
+dat_models_specifications$below_alpha = factor(dat_models_specifications$below_alpha)
 
@@
 
-# Order by odds ratio, low to high
+# Order by odds ratio, low to high and assign identifier
 analysed_specifications = dat_models_specifications[order(dat_models_specifications$estimate_oddsratio), ]
-
-# Assign identifier
 analysed_specifications$id = factor(seq(1:nrow(analysed_specifications)))
 
-# Preparing specification matrix i.e., the bottom part of the SCA plot
-## Subset covariates and statistics
-prepare_long_df = analysed_specifications[ , c(4:18, 20, 21, 24)]
+# Preparing the bottom part of the SCA plot by transforming wide to long
+## Subset covariates and relevant statistics
+prepare_long_df = analysed_specifications %>% select(specific_pos:games, id)
 
-## FALSE ro NA to better subset
-prepare_long_df[prepare_long_df == FALSE] = NA
+## transform wide to long
+long_df = melt(data.table(prepare_long_df), "id", variable.names = "covariate")
+long_df = long_df[long_df$value == TRUE, c("id", "variable")]
 
-## Wide to long format
-long_df = melt(setDT(prepare_long_df), id.vars = c("id", "below_alpha"), variable.name = c("estimate_oddsratio"))
+## statistics are still missing (odds ratio and significance)
+id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha)
 
-## Exclude NA cases
-long_df = long_df[complete.cases(long_df), ]
+## Adding (or joining) the statistics to the long data frame
+long_df = data.frame(left_join(long_df, id_oddsratio_below_alpha, by = "id"))
 
-## Exclude outcome variable
-long_df = long_df[!grepl("estimate_oddsratio", long_df$estimate_oddsratio), ]
-colnames(long_df) = c("id", "below_alpha", "covariate", "x")
-
-## Exclude value column
-long_df = long_df[ , c("id", "below_alpha", "covariate")]
-
-## order covariate effects low to high, based on the 4_covariate_effect_data.csv dataset
-impact_df = dat_covariates[order(dat_covariates$estimate_oddsratio), ]
-
-long_df$covariate = factor(long_df$covariate,
+# Assign proper names for visualisation
+long_df$variable = factor(long_df$variable,
                            levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
-                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Lg.
-country", "Position", "Age", "Ref", "Cards assig.", "Goals", "Height", "Games", "Ref country", "Player"))
+                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Country", "Position", "Age", "Referee", "Cards assig.", "Goals", "Height", "Games", "Ref.
+country", "Player"))
 
@@
 top = analysed_specifications %>% 
+  filter(estimate_oddsratio < 4) %>% 
   ggplot(aes(as.numeric(id), estimate_oddsratio)) +
@@
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
-  labs(title = "Specification curve", subtitle = "N = 1,000; 95% CI") +
+  labs(title = "Specification curve", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = "")) +
   theme_classic() +
@@
 bottom = long_df %>%
-  filter(id != "1") %>%
-  ggplot(aes(x = id, y = covariate)) +
+  ggplot(aes(x = id, y = reorder(x = variable, X = as.numeric(estimate_oddsratio)))) +
   geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
@@
 Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383.
-The 1rd quartile   \newline
+The 1st quatile equals to 1.206, the median equals to 1.248 and the 3rd quartile equals to 1.293.
+Therefore, as also evidenced by the probability density distribution, 50% of the lay withing a very narrow interval (between 1st and 3rd quartile).
+The outcome of *team 23* (1.31) is still included in the interval, but it is outside the middle 50% of the data.
+
+ The probability density distribution also has an interesting shape, having two peaks.
+This suggests a particular covariate combination structure causing two focal points.
+Based on this graph alone though it is not possible to decipher the underlying structure though.
+What also seem to be more apparent in this graph than in the specification curve is that the non-significant values also seem to revolve round to focal points.
+One being around 1.13 and the other around 1.19.
+This also suggest some influential covariate or covariate structure that is cause non-significance.
+
+ Finally, for transparency purpose, it also needs to be mentioned that eight points were excluded from the visualisation.
+Four of them were "infinite", one effectively approaching infinity, one being in the lowest million range, one being 150, and a final being 4.6.
+I specifically chose not to include the latter one in the graph as it would despite is relatively close numerical proximity skewed the distribution being at least a visual outlier.
+\newline
 
@@
 
-vibration_of_effect = dat_models %>%
-  filter(estimate_oddsratio > 0 & estimate_oddsratio <= 4) %>% 
+vibration_of_effect = dat_models_specifications %>%
+  filter(estimate_oddsratio < 4) %>% 
   ggplot(aes(x = "", y = estimate_oddsratio)) +
@@
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
-  labs(title = "Results space defined through covariate specification", subtitle = "N = 1,000; Odds ratio (OR)") +
+  labs(title = "Results space defined through covariate specification", subtitle = paste0("N = ", nrow(dat_models_specifications), ", Odds ratio (OR)", collapse = "")) +
   coord_flip() +
@@
 ## Specified covariate effects
+     The final graph shows the specified covariate effects.
+On the horizontal axis are the covariates, on the vertical axis are the estimates.
+The covariates are sorted from negative to positive impact.
+Black indicates a significant effect, red non-significant.
+The error bars represent the 95% CI.
+While four covariates are significant, their effects seem relatively weak, ranging from -0.1 to 0.05, only.
+Interesting is also which covariates are significant.
+The *team 23* analysis included league country, position, referee and player as covariates.
 \newline
@@
   scale_y_continuous(name = "Estimates", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
-  labs(title = "Specified covariate effects", subtitle = "N = 1,000; 95% CI", x = "Covariates") +
+  labs(title = "Specified covariate effects", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = ""), x = "Covariates") +
   theme_classic() +

### 2021-08-15 scripts/4_multiverse_analysis.R

@@ -220,13 +220,47 @@ df$ci_upper_oddsratio = with(df, exp(ci_upper))
 sum(df$below_alpha == "Significant") / nrow(df) # = 89.7%
 
 
-write.csv(df, here::here("data", "4_model_outcome_data.csv"), row.names = FALSE)
+# Save a potential interim outcome
+# write.csv(df, here::here("data", "4_model_outcomes_data.csv"), row.names = FALSE)
+
+
+# Build data frame that includes the specifications and the analysis results
+analysed_specifications = inner_join(specifications, df, by = "formula")
+
+
+# Joining results in 5 missing cases, which are identified. Apparently due to a "1" at the formulas' end.
+# No apparent systematic error. N = 5, hence, deemed negligible. 
+indexing_missing_cases = df$formula %in% analysed_specifications$formula
+df[!indexing_missing_cases, ]
+
+
+# Summary statistics show OR higher than 150; those are excluded as outliers
+unsually_high_odds_ratio = df %>% filter(estimate_oddsratio > 100)
+
+
+# Excluding the seven outliers
+analysed_specifications = analysed_specifications %>% filter(estimate_oddsratio < 100)
+
+
+# Assigning significance labels
+analysed_specifications$below_alpha = with(analysed_specifications, ifelse(p_value < alpha_level, "Significant", "Non-significant"))
+
+# Selecting relevant columns for a better overview (covariates and "estimate_oddsratio")
+analysed_specifications = analysed_specifications[ , c(base_var, covar, "p_value", "below_alpha", "estimate_oddsratio", "ci_lower_oddsratio", "ci_upper_oddsratio")]
+
+
+# The original variable names contain variable class e.g., "factor(specific_pos)". Those are removed by assigning "clean" names.
+colnames(analysed_specifications) = c("skin_tone_num", "imp_bias", "exp_bias", "specific_pos", "height_cm", "weight_kg", "league_country", "age_yrs", "goals", "club", "ref_country", "ref", "victories", "player_cards_received", "player", "ref_cards_assigned", "ties", "games", "p_value", "below_alpha", "estimate_oddsratio", "ci_lower_oddsratio", "ci_upper_oddsratio")
+
+
+# Save as csv
+write.csv(analysed_specifications, here::here("data", "4_model_outcomes_specifications_merged.csv"), row.names = FALSE)
 
 
[…]

### 2021-08-15 dissertation/msc_dissertation.Rmd
Further cleaning

@@
   - \usepackage[ruled,vlined]{algorithm2e}
+  - \usepackage{float}
 ---
@@
 
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')
 
@@
                            levels = c("club", "victories", "weight_kg", "player_cards_received", "ties", "league_country", "specific_pos", "age_yrs", "ref", "ref_cards_assigned", "goals", "height_cm", "games", "ref_country", "player"),
-                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "Country", "Position", "Age", "Referee", "Cards assig.", "Goals", "Height", "Games", "Ref.
+                           labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "League", "Position", "Age", "Referee", "Cards assig.", "Goals", "Height", "Games", "Ref.
 country", "Player"))
+
+# Covariates should be sorted based in their impact.
+dat_covariates = dat_covariates[order(dat_covariates$estimate_oddsratio), ]
+ordered_covar = dat_covariates$impact_names
 
@@
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(0.9, 1.7, 0.1)), limits = c(0.89, 1.71), expand = c(0, 0)) +
-  labs(title = "Specification curve", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = "")) +
+  labs(title = "Specification curve",
+       subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = "")) +
   theme_classic() +
@@
         axis.line.x = element_blank(),
-        plot.margin = unit(c(0.5, 1, 0, 1), "cm"),
+        plot.margin = unit(c(0, 1, 0, 1), "cm"),
         legend.position = "top",
@@
 bottom = long_df %>%
-  ggplot(aes(x = id, y = reorder(x = variable, X = as.numeric(estimate_oddsratio)))) +
+  ggplot(aes(x = id, y = factor(variable, levels = ordered_covar))) +
   geom_tile(aes(fill = below_alpha), width = 0.5, height = 0.5, color = "white") +
@@
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
-  labs(title = "Results space defined through covariate specification", subtitle = paste0("N = ", nrow(dat_models_specifications), ", Odds ratio (OR)", collapse = "")) +
+  labs(title = "Results space defined through covariate specification",
+       subtitle = paste0("N = ", nrow(dat_models_specifications), ", Odds ratio (OR)", collapse = "")) +
   coord_flip() +
@@
   scale_color_manual(values = c("red", "black")) +
-  scale_y_continuous(name = "Estimates", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
-  labs(title = "Specified covariate effects", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = ""), x = "Covariates") +
+  scale_y_continuous(name = "Estimates",
+                     labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
+  labs(title = "Specified covariate effects",
+       subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = ""), x = "Covariates") +
   theme_classic() +

### 2021-08-15 scripts/4_multiverse_analysis.R

@@ -270,7 +270,8 @@ vibration_of_effect = analysed_specifications %>%
   annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
   scale_color_manual(values = c("red", "black")) +
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
-  labs(title = "Vibration of effect due to covariate specification", subtitle = paste0("N = ", nrow(analysed_specifications), ", Odds ratio (OR)", collapse = "")) +
+  labs(title = "Vibration of effect due to covariate specification",
+       subtitle = paste0("N = ", nrow(analysed_specifications), ", Odds ratio (OR)", collapse = "")) +
   coord_flip() +
   theme_classic() +
   theme(axis.line.y = element_blank(),
@@ -307,7 +308,7 @@ impact_summary = summary(impact_mod)
 impact_coef = data.frame(impact_summary$coefficients)[c(-1), "Estimate"]
 impact_se = data.frame(impact_summary$coefficients)[c(-1), "Std..Error"]
 impact_pvalue = data.frame(impact_summary$coefficients)[c(-1), "Pr...t.."]
-impact_names = c("Position", "Height", "Weight", "Leauge", "Age", "Goals", "Club", "Ref. country", "Referee", "Victories", "Cards rec.", "Player", "Cards assig.", "Ties", "Games")
+impact_names = c("Position", "Height", "Weight", "League", "Age", "Goals", "Club", "Ref. country", "Referee", "Victories", "Cards rec.", "Player", "Cards assig.", "Ties", "Games")
 
 impact_df = data.frame(cbind(impact_names,
                              impact_coef,
@@ -337,8 +338,10 @@ covariate_effects = impact_df %>%
   geom_hline(yintercept = 0) +
   geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper, color = below_alpha), width = 0.1) +
   scale_color_manual(values = c("red", "black")) +
-  scale_y_continuous(name = "Estimates\n", labels = scales::comma_format(accuracy = 0.05), breaks = c(seq(-0.15, 0.15, 0.05)), limits = c(-0.15, 0.15)) +
-  labs(title = "Specified effect of covatiates", subtitle = paste0("N = ", nrow(analysed_specifications), ", 95% CI", collapse = ""), x = "Covariates") +
+  scale_y_continuous(name = "Estimates\n",
[…]

### 2021-08-16 dissertation/msc_dissertation.Rmd
adding volcano plot

@@
   - \usepackage{float}
+  - \usepackage{caption}
+  - \captionsetup[figure]{font={stretch=1}}
 ---
@@
 # Introduction
+## Theoretical framework
      The Covid-19 pandemic has reaffirmed the need for sound scientific research to inform decision-making on a scientific and societal level [@collins2021].
@@
 
- These efforts include a crowdsourcing approach to data analysis [@silberzahn2018; @botvinik-nezer2020].
+ These efforts include crowdsourcing approaches to data analysis [@silberzahn2018; @botvinik-nezer2020].
 Its premise is to have a large number of researchers team up into to smaller, independent groups to investigate the same research question based on the same dataset.
@@
 @botvinik-nezer2020 made similar observations.
-Crowdsourcing data analysis excels at emphasising the substantial impact of different analytical strategies, but has a significant drawback.
-It is extremely time and personnel intensive, the two mentioned studies lasted between 2 to 3+ years, and included 61 and 180 analysts, respectively.
-Not to mention the organisational effort.
+Crowdsourcing projects excels at emphasising the substantial impact of different analytical strategies, but they are also extremely time and personnel intensive.
+The two mentioned studies lasted between 2 to 3+ years, and included 61 and 180 analysts, respectively.
+Additionally, despite covering a more diverse set of analytical strategies than a single conventional analysis, crowdsourcing is limited by the number of participating teams.
+It, therefore, still only covers a few selected strategies, not all possible ones.
 
@@
 Multiverse approaches are therefore well suited for assessing the scope of variation induced by different analytical strategies.
-For instance, @patel2015 assessed an effect's robustness by running every combination of covariates and modelling options.
-Their conclusion was the larger the variation of outcomes, the less robust is the effect, and the less it should therefore to be trusted.
-The authors termed this measure the *vibration of effect* (VoE).
+However, one aspect has been, to my knowledge, neglected thus far.
+To properly assess the scope of variation it needs to be contextualised.
+That is, the space in which it can vary needs to be defined in order to understand to what degree the scope covers the space.
+The concept of statistical association provides a suitable analogy.
+Covariation and correlation both indicate the direction of relationship (positive, null, negative), however, only correlation allows for assessing the association's strength.
+Correlation is the standardised version of covariation, and only knowing the its limits gives the correlation coefficient meaning.
+The same principle applies to analyses on a broader level.
 
- Taken together, researchers make many analytical decisions throughout conduction a study which are know as *researcher degrees of freedom*.
-Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its substantial time and personnel requirements.
+## Current project
+     Taken together, researchers make many analytical decisions throughout conducting a study introducing so-called *researcher degrees of freedom*.
+Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its time and personnel requirements, and is still limited by the number of participating teams.
 Multiverse analysis present a promising, low-personnel, but computationally intense alternative to assessing the scope this variation.
-Despite observing substantial variation in crowdsourced data analysis projects little is known about the extend to which these approaches cover the full range of possible results.
-In this project, using principles of multiverse analysis I will define the results space of @silberzahn2018.
-By doing so I aim at further understanding how a specific analysis strategy affects its result.
-To my knowldge this is the first project runing not only all *plausible* but all *possible* analytical strategies to a more granular view of the results space.
+Despite observing substantial variation in crowdsourced data analysis projects, little is known about the extend to which these approaches cover the full range of possible results.
+Determining the full results space and its underlying mechanisms is an important exercise to understand the impact of analytical strategies.
+To my knowledge no previous study has attempted to define and understand a study's entire results space.
+In this project, running all possible analytical strategies, I, therefore, aim to define and understand the results space of @silberzahn2018.
+By doing so, I hope further the understanding of how an analytic strategy affects its result.
 
@@
 ## statistics are still missing (odds ratio and significance)
-id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha)
+id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha, p_value)
 
@@
 
-# Covariates should be sorted based in their impact.
+# Covariates should be sorted based in their impact
 dat_covariates = dat_covariates[order(dat_covariates$estimate_oddsratio), ]
@@
         legend.position = "top",
+        legend.justification = "right",
         legend.title.align = 1,
-        legend.text = element_text(size = 11),
+        legend.text = element_text(size = 10),
         legend.direction = "horizontal",
         legend.title = element_blank(),
-        legend.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
-        legend.box.margin = margin(t = -0.5, b = -0.2, unit = "cm"),
+        legend.margin = margin(t = -0.61, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.61, b = -0.2, unit = "cm"),
         plot.title = element_text(face = "bold"),
@@
   geom_hline(yintercept = 1.31, linetype = "dashed", color = "black") +
-  annotate("text", x = 1.516, y = 1.375, label = "Silberzahn et al. (2018) \nMedian OR", color = "black") +
+  annotate("text", x = 1.559, y = 1.395, label = "Silberzahn et al. (2018) median OR", color = "black", size = 3.5) +
   scale_color_manual(values = c("red", "black")) +
@@
         legend.title.align = 1,
-        legend.text = element_text(size = 11),
+        legend.text = element_text(size = 10),
         legend.direction = "vertical",
@@
      The final graph shows the specified covariate effects.
-On the horizontal axis are the covariates, on the vertical axis are the estimates.
+On the horizontal axis are the covariates, on the vertical axis the estimates.
 The covariates are sorted from negative to positive impact.
@@
 The error bars represent the 95% CI.
-While four covariates are significant, their effects seem relatively weak, ranging from -0.1 to 0.05, only.
+Four out of 15 covariates are significant, roughly 27%.
+Thought the significant covariates' effects seem relatively weak, ranging from -0.1 to 0.05, only.
 Interesting is also which covariates are significant.
-The *team 23* analysis included league country, position, referee and player as covariates.
+The *team 23* analysis included league, position, referee and player as covariates.
+Of those only player is a significant covariate.
+Using crude mathematics, one might make the case that the added effect of players is the reason for moving the overall effect from the median, roughly 1.25 to 1.30.
+Though that's more speculation than causal reasoning.
+As mentioned above, the roughly 90% of all outcome were statistically significant, given that only 27% of the covariates have an overall significant effects suggest that the effect is of skin tone itself is fairly robust.
 \newline
@@
         legend.position = "top",
-        legend.justification = "left",
+        legend.justification = "right",
         legend.title.align = 1,
-        legend.text = element_text(size = 11),
-        legend.direction = "vertical",
+        legend.text = element_text(size = 10),
+        legend.direction = "horizontal",
         legend.title = element_blank(),
-        legend.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
-        legend.box.margin = margin(l = 0.1, b = -0.729, unit = "cm"),
+        legend.margin = margin(t = -0.33, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.33, b = -0.2, unit = "cm"),
         plot.title = element_text(face = "bold"),
@@
 
+```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center"}
+
+volcano_plot = dat_models_specifications %>% 
+  filter(estimate_oddsratio < 4) %>% 
+  ggplot(aes(estimate_oddsratio, log10(p_value))) +
+  geom_point(aes(colour = below_alpha), alpha = 0.3) +
+  scale_x_continuous(name = "Odds ratios", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
+  scale_y_continuous(name = "Log10 p-value") +
+  scale_color_manual(values = c("red", "black")) +
+  labs(title = "Vibration of Effect",
+       subtitle = paste0("N = ", nrow(analysed_specifications), collapse = "")) +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "grey"),
+        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
+        legend.position = "top",
+        legend.justification = "right",
+        legend.title.align = 1,
+        legend.text = element_text(size = 10),
+        legend.direction = "horizontal",
+        legend.title = element_blank(),
+        legend.margin = margin(t = -0.6, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.6, b = -0.2, unit = "cm"),
+        plot.title = element_text(face = "bold"),
+        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
+        text = element_text(size = 10)) +
+  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))
+
+volcano_plot
+
+```
+
 \newpage
@@
 # Discussion
+The current project used multiverse analysis to define the results space of the @silberzahn2018.
+Among others multiverse analysis combines different statistical modelling approaches along with covariate combinations.
+The goals is assess the robustness of effects.
+Due to time and resources limitations, I chose to focus on statistical modelling approach, but explore every possible covariate combination.
+Usually, multiverse analysis would focus on all plausible (or reasonable) analytic strategies.
+However, this project sought to define the results space of all *possible* analytical outcomes to investigate in which space the outcomes of the *plausible* analytical strategies vary in the first place.
+Only if the results space is defined, is possible assess its content.
+The concept of statistical association, provided a neat analogy.
+There is covariation and correlation, both quantify the variability of two variables.
+Both indicate the direction of relationship (positive, none, negative).
+However, to assess the strength of the relationship only correlation is informative.
+Correlation is the standardised version of covariation, hence, it has a scale to it.
+And only knowing the limits to space of variation gives the correlation coefficient meaning.
+The same applies to analyses on a broader level.
+If researchers want to be able to tell if their analysis reflects a true effect or simply a luck guess, they need to know the entire results space.
+Only then they can identify the robustness of effects.
+
+ Due to time and resource limitation this project focused on pne
+
+In order to assess the strength of the association covariation is hardly useful as covariate it does not have a standarised scale to it.
 

### 2021-08-16 scripts/4_multiverse_analysis.R

@@ -386,7 +386,7 @@ long_df = melt(data.table(prepare_long_df), "id", variable.names = "covariate")
 long_df = long_df[long_df$value == TRUE, c("id", "variable")]
 
 ## statistics are still missing (odds ratio and significance)
-id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha)
+id_oddsratio_below_alpha = analysed_specifications %>% select(id, estimate_oddsratio, below_alpha, p_value)
 
 ## Adding (or joining) the statistics to the long data frame
 long_df = data.frame(left_join(long_df, id_oddsratio_below_alpha, by = "id"))
@@ -398,7 +398,7 @@ long_df$variable = factor(long_df$variable,
                            labels = c("Club", "Victories", "Weight", "Cards rec.", "Ties", "League", "Position", "Age", "Referee", "Cards assig.", "Goals", "Height", "Games", "Ref. country", "Player"))
 
 
-# Covariates should be sorted based in their impact.
+# Covariates should be sorted based in their impact
 impact_df = impact_df[order(impact_df$estimate_oddsratio), ]
 ordered_covar = impact_df$impact_names
 

### 2021-08-16 scripts/4_multiverse_analysis.R
Adding volcano plot

@@ -451,3 +451,36 @@ sca_plot = plot_grid(top, bottom, ncol = 1, align = "v")
 sca_plot
 
 ggsave(here::here("figures", "sca_plot.png"), sca_plot, width = 10, height = 6)
+
+
+#..............................................# Volcano plot #............................................#
+
+
+volcano_plot = dat_models_specifications %>% 
+  filter(estimate_oddsratio < 4) %>% 
+  ggplot(aes(estimate_oddsratio, log10(p_value))) +
+  geom_point(aes(colour = below_alpha), alpha = 0.3) +
+  scale_x_continuous(name = "Odds ratios", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
+  scale_y_continuous(name = "Log10 p-value") +
+  scale_color_manual(values = c("red", "black")) +
+  labs(title = "Vibration of Effect",
+       subtitle = paste0("N = ", nrow(analysed_specifications), collapse = "")) +
+  theme_classic() +
+  theme(panel.grid.major = element_line(colour = "grey"),
+        plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
+        legend.position = "top",
+        legend.justification = "right",
+        legend.title.align = 1,
+        legend.text = element_text(size = 10),
+        legend.direction = "horizontal",
+        legend.title = element_blank(),
+        legend.margin = margin(t = -0.6, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.6, b = -0.2, unit = "cm"),
+        plot.title = element_text(face = "bold"),
+        plot.subtitle = element_text(face = "italic", margin = margin(0, 0, 20, 0)),
+        text = element_text(size = 10)) +
+  guides(colour = guide_legend(override.aes = list(alpha = 1, size = 1), reverse = TRUE))
+
+volcano_plot
+
+ggsave(here::here("figures", "volcano_plot.png"), volcano_plot, width = 10, height = 6)

### 2021-08-21 dissertation/msc_dissertation.Rmd
Dissertation editing

@@
 subtitle: |
-  | *Defining the results space in @silberzahn2018 through model specification*
+  | *Assessing results space and structure of @silberzahn2018 through model specification*
   |
-  | Master's thesis - DRAFT
+  | MSc Dissertation - DRAFT
 author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
-linestretch: 1.5
 output:
@@
     fig_caption: yes
+  word_document: default
+linestretch: 1.5
 urlcolor: blue
@@
 
-... \newline
-*Abstract--in progress.* \newline
-... \newline
+ \newline
+
+__ABSTRACT__ Numerous studies have show that analytical choices researchers make throughout the process of conducting a study affect the outcome.
+Recent crowdsourcing approaches to data analysis show that even well-intentioned analytical choices can make the difference between find a significant and strong or a non-significant effect.
+IN PROGRESS
 
@@
 Multiverse approaches are therefore well suited for assessing the scope of variation induced by different analytical strategies.
-However, one aspect has been, to my knowledge, neglected thus far.
-To properly assess the scope of variation it needs to be contextualised.
-That is, the space in which it can vary needs to be defined in order to understand to what degree the scope covers the space.
+However, to my knowledge, there is one aspect that has been neglected thus far.
+That is to what extend does the identified scope of variation cover the range of possible outcome.
+If the results space and its structure is unknown, it is impossible to tell the observed results were coincidental or in fact point to a true latent construct.
 The concept of statistical association provides a suitable analogy.
-Covariation and correlation both indicate the direction of relationship (positive, null, negative), however, only correlation allows for assessing the association's strength.
+Covariation and correlation both indicate the direction of relationships (positive, null, negative), however, only correlation allows for assessing the association's strength.
 Correlation is the standardised version of covariation, and only knowing the its limits gives the correlation coefficient meaning.
-The same principle applies to analyses on a broader level.
+The same principle applies the scope of variation and the range of all possible results.
 
@@
 Despite observing substantial variation in crowdsourced data analysis projects, little is known about the extend to which these approaches cover the full range of possible results.
-Determining the full results space and its underlying mechanisms is an important exercise to understand the impact of analytical strategies.
-To my knowledge no previous study has attempted to define and understand a study's entire results space.
+Determining the full results space and its underlying structure is an important exercise to understand the impact of analytical strategies.
+To my knowledge no previous study has attempted to define and understand a study's full results space.
 In this project, running all possible analytical strategies, I, therefore, aim to define and understand the results space of @silberzahn2018.
@@
 ## Analysis plan
-     The project's objective is to define the results space of the @silberzahn2018 to draw conclusions about the mechanisms of multiverse analysis.
+     The project's objective is to define the results space and structure of @silberzahn2018 to draw conclusions about the mechanisms of multiverse analysis.
 The results space refers to the numerical interval between the lowest and highest possible outcome.
@@
 Robustness refers to an effect's consistency under different model specifications.
-@patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*.
+@patel2015 developed a standardised approach to assessing an effect's robustness.
 I will therefore use this standardised approach to define the results space.
@@
 In particular, on the approach that produced the median outcome of all analyses.
-The median, being the middle number of any set of values, is a reasonable starting point in order to estimate the results space.
-Nevertheless, even focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities.
-Due to computational limitations I will first run a sample of 200, followed by a sample of 1,000 combinations.
-If there's a substantial difference between the median and the spread I will run more models, if not, I will make the assumption that 1,000 reasonably estimates all ~33k combinations.
+The median, being the middle number of a given set of values, is a reasonable starting point in order to estimate the results space.
+Nevertheless, even focusing on one analytical approach leaves $1 * 2^{15} = 32,768$ possibilities.
+Due to computational limitations I will run a sample of 1,000 combinations.
 
@@
 
-(i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done.
-Its objective is to describe the results space while allowing the reader to identify the specified covariates for each outcome [@simonsohn2020].
+(i) The specification curve plot was developed to be a descriptive plot i.e., raw outcome data being plotted without any aggregation done.
+Its objective is to describe the results space while allowing the reader to identify the model specifications for each outcome [@simonsohn2020].
 It therefore has two vertically stacked components.
 The top part shows a sorted scatter plot.
-Its horizontal axis shows the models specifications and its vertical axis displays the outcome measure.
+Its horizontal axis lists the specifications and its vertical axis displays the outcome measure.
 The scatter points are sorted from lowest to highest outcome measure.
@@
 The bottom part of the plot represents a table.
-As in the top plot the horizontal axis holds the model specifications, the vertical axis lists the covariates.
+As in the top plot the horizontal axis lists the specifications, the vertical axis lists the covariates.
 This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model.
-For the top and bottom plot to work together it is imperative that the horizontal axes are identically arranged.
+For the top and bottom plot to work together, it is therefore imperative that the horizontal axes are identically arranged.
 If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot.
 For an example check Figure \ref{fig:sca_plot}.
-As the current project has 1,000+ specifications it becomes very hard to identify specific covariates given the space limitations.
+As the current project has 1,000+ specifications it gets very hard to identify specific covariates due to the space limitations.
 Hence, I to had come up with different more intuitive ways of visualisation.
@@
 Its objective is to give an unbiased, transparent view on the raw data.
-It therefore combines a density, box and scatter plot.
+It combines a density, box and scatter plot.
 If stacked vertically from top to bottom, respectively, they look like a cloud with rain drops, hence, its name.
 The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot.
-The current project seeks to define and describe a results space the rain-cloud plot is, therefore, the ideal choice of graphs.
-(For an example check the results section.) Unfortunately, the plot does not allow to identify the covariates giving rise to the outcomes or their specified effects.
-As the project also seeks to understand the mechanisms of multiverse analysis it important to get insights about the specified covariates effects.
+The current project seeks to define and describe a results space, the rain-cloud plot is, therefore, the ideal choice of graphs.
+(For an example check Figure \ref{fig:rain_cloud_plot}.) Unfortunately, the plot does not allow to identify the covariates giving rise to the outcomes or their specified effects.
+As the project also seeks to understand the structure of the results space it important to get insights about the specified covariates effects.
 Therefore a third plot is developed which is to be assessed in tandem with this plot.
 
-(iii) The third and final plot is a sorted scatter plot.
+(iii) The third plot is a sorted scatter plot.
 While there is nothing special about the plot itself its content is important.
-Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation.
-Instead the effects for each covariates need to be aggregated.
-Hence, the goal of this plot is to visualise the specified effect for each covariate.
+Given the large number of models, it does not make sense to assess each model and its effects in individually.
+Instead the effects for each covariate need to be aggregated.
+The goal of this plot is to visualise the specified effect for each covariate.
 Prior to doing so, the covariates first need to be aggregated and their specified effects need to be calculated.
 To this end, each covariate is recoded into a binary factor: included in the model yes/no.
-This newly defined factors are than used as an independent variables in an ANOVA.
+These newly defined factors are than used as an independent variables in an ANOVA.
 The previously calculated odds rations are used as dependent variables.
@@
 
- Taken together, the current project's objective is to define the results space of @silberzahn2018 in order to make inferences about the underlying mechanisms of multiverse analysis.
+ Taken together, the current project's objective is to define the results space and structure of @silberzahn2018 to gain a better understanding of how analysis strategy affects analysis results.
 The results space is created by running all possible model and covariate combinations.
@@
  My code is based on @haessler2020 and @patel2015.
-The rough structure of the code is shown in the pseudocode table (Figure X).
+The rough structure of the code is shown in the pseudocode table (see Algorithm 1).
 Here, I am briefly motivating two analytical choices I made.
@@
 For this project, it was important to check that both functions use the same estimation method (maximum likelihood or restricted maximum likelihood estimation) to calculate the statistics.
-It was important to check to ensure the outcomes are comparable.
+It is important to check estimation method to ensure the outcomes are comparable.
 
@@
 Those would not have knowledge of the other analyses and, hence, would not account for them.
-To main the highest possible ecological validity, I therefore did not account of multiple comparisons.
+To maintain the highest possible ecological validity, I therefore did not account of multiple comparisons.
 
@@
 The exploratory data analysis (EDA) of *team 23* showed that each row of the dataset represents a unique player-referee combination listing all their encounters as well as the couple's total number of received/assigned cards.
-The team states that it preferred a different data format.
+*team 23* states that it preferred a different data format.
 More specifically, a format where each player-referee encounter is reflected by one row.
 This way each encounter had a maximum of one red card.
-To achieve this format the data had to be "disaggregated" i.e., transformed.
+To achieve this format the data had to be transformed more specifically, disaggregated.
 Further EDA showed that receiving a red card is highly unlikely, 0.008%, i.e., data is highly skewed.
 This property informed the teams decision to use a type of logistic regression (a statistical modelling technique equipped to deal with skewed/binary distributions).
-Finally, the team excluded all referee that did have at least 22 encounters.
+Finally, the team excluded all referees that did have at least 22 encounters.
 Their reasoning was a game includes (at least) 22 players, 11 players per team, if this were not the case referee most likely did not officiate in one for the four leagues of interest (England, Germany, Italy and Spain).
@@
 The relatively smooth line suggest that sampling 1,000 specifications is, in fact, representative for all 33k models.
-The law of large numbers  The points can be divided into roughly three sections all non-significant ($\leq 57^{th}$), mixture of (non-)significance ($58^{th} - 247^{th}$) and all significance ($\geq 248^{th}$).
+The points can be divided into roughly three sections all non-significant ($\leq 57^{th}$), mixture of (non-)significance ($58^{th} - 247^{th}$) and all significance ($\geq 248^{th}$).
 The bottom table lists all covariates on its vertical axis, sorted from most positive to most negative impact.
@@
 Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table.
-Regular specification curve analyses focus on all *reasonable* specifications, hence, their number of specifications is much lower and one is, in fact, able to tell the different model specifications apart.
-However, the current project's goal was to define the results space, which requires all *possible* specifications.
-The specifications are therefore huddled extremely closely which does not allow the plot to functions in its supposed way.
-Thus, I create two alternative graphs.
+Regular specification curve analyses focus on all *reasonable* specifications, hence, their number of specifications is much lower.
+Keeping the intended purpose in mind one would, in fact, be able to tell the different model specifications apart.
+However, the current project's goal was to define the results space, which requires *all possible* specifications.
+The specifications are therefore huddled too closely which does not allow the plot to functions in its supposed way.
+What is interesting though are the top and bottom row of the table.
+The top row shows that player is particularly often included in models with a higher outcome.
+Conversely, club is particularly often included in models with a lower outcome.
+Full disclosure, I only noticed this pattern after having calculating the specified covariate effects.
 \newline
@@
 Based on this graph alone though it is not possible to decipher the underlying structure though.
-What also seem to be more apparent in this graph than in the specification curve is that the non-significant values also seem to revolve round to focal points.
+What also seems to be more apparent in this graph than in the specification curve is that the non-significant values also seem to revolve round to focal points.
 One being around 1.13 and the other around 1.19.
@@
 
- Finally, for transparency purpose, it also needs to be mentioned that eight points were excluded from the visualisation.
-Four of them were "infinite", one effectively approaching infinity, one being in the lowest million range, one being 150, and a final being 4.6.
-I specifically chose not to include the latter one in the graph as it would despite is relatively close numerical proximity skewed the distribution being at least a visual outlier.
+ Concering the results space, for transparency purposes, it also needs to be mentioned that eight points were excluded from all visualisation.
+Four of them were "infinite", one effectively approaching infinity, one being in the lower million range, one being 150, and a final being 4.6.
+I specifically chose not to include the latter one in the graph as it would despite is relatively close numerical proximity skewed the distribution, at least visually.
 \newline
@@
 ## Specified covariate effects
-     The final graph shows the specified covariate effects.
+     This graph shows the specified covariate effects.
 On the horizontal axis are the covariates, on the vertical axis the estimates.
@@
 Four out of 15 covariates are significant, roughly 27%.
-Thought the significant covariates' effects seem relatively weak, ranging from -0.1 to 0.05, only.
-Interesting is also which covariates are significant.
+The significant covariates' effects seem relatively weak, ranging from -0.1 to 0.05, only.
+Interesting is which covariates are significant.
 The *team 23* analysis included league, position, referee and player as covariates.
@@
 Though that's more speculation than causal reasoning.
+If anything, this graph shows that the individual covariate effect is likely not as influential as interactions of covariates.
 As mentioned above, the roughly 90% of all outcome were statistically significant, given that only 27% of the covariates have an overall significant effects suggest that the effect is of skin tone itself is fairly robust.
@@
 
-```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center"}
+## Systematic latent structure
+     This graph was not planned and yet it is too interesting not to include it in this report (though it will only serve to emphasise previous points and probe future research).
+I was very keen to investigating the underlying covariate structure assumed based on \ref{fig:rain_cloud_plot}.
+I, therefore, went back to the previous mentioned @patel2015 study.
+Despite having a different objective, their methodology is similar to the current project's (rather the other way around).
+They visualised the results as so-called volcano plot (see Figure \ref{fig:volcano_plot}) which show the relationship between the p-value and effects size.
+So, I was interested how the graph would look like in the current project's case.
+It turns the graph evidences two distinct underlying constructs--perhaps even three.
+Particular salient is how distinct these "stream" are.
+Given that the specified covariate effects are relative weak, it is highly unlikely that only one is responsible for cause such structure.
+In fact, it is highly likely that a combination of covariates in play here.
+Considering the apparent influence of this deterministic structure is for future research to decipher the underlying covariate combination.
+
+```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Volcano plot showing the the relationship between p-values and effect sizes.
+The degree of strictness in which the group are seperated suggest a strong deterministic latent covariate structure."}
 
@@
   scale_color_manual(values = c("red", "black")) +
-  labs(title = "Vibration of Effect",
+  labs(title = "Systemtic latent structure",
        subtitle = paste0("N = ", nrow(analysed_specifications), collapse = "")) +
@@
 
-\newpage
+# Discussion
+     The current project aimed at defining and understanding the full results space of @silberzahn2018.
+The rational was if the full results space is known, it possible to better estimate the extend to which analytical strategies influence their results.
+To this end, out all analytic strategies, every possible combination was supposed to be run in to determine the full results space.
+Due to time and computational limitations, the current project focused on one analytical strategy and sampled 1,000 covariate combinations out of all possible one.
+The chosen analytical strategy (i.e., the statistical model) was the one producing the median outcome of the original study.
 
-# Discussion
-The current project used multiverse analysis to define the results space of the @silberzahn2018.
-Among others multiverse analysis combines different statistical modelling approaches along with covariate combinations.
-The goals is assess the robustness of effects.
-Due to time and resources limitations, I chose to focus on statistical modelling approach, but explore every possible covariate combination.
-Usually, multiverse analysis would focus on all plausible (or reasonable) analytic strategies.
-However, this project sought to define the results space of all *possible* analytical outcomes to investigate in which space the outcomes of the *plausible* analytical strategies vary in the first place.
-Only if the results space is defined, is possible assess its content.
-The concept of statistical association, provided a neat analogy.
-There is covariation and correlation, both quantify the variability of two variables.
-Both indicate the direction of relationship (positive, none, negative).
-However, to assess the strength of the relationship only correlation is informative.
-Correlation is the standardised version of covariation, hence, it has a scale to it.
-And only knowing the limits to space of variation gives the correlation coefficient meaning.
-The same applies to analyses on a broader level.
-If researchers want to be able to tell if their analysis reflects a true effect or simply a luck guess, they need to know the entire results space.
-Only then they can identify the robustness of effects.
+ This very analysis was first attempted to be fully replicated, yielding the identical results as the original team, *team 23*, did.
+This ensured an identical starting from all further analyses.
+After drawing the 1,000 unique combinations of covariates and running the respective models, the results were visualised.
+Removing unrealistic values (anything above 100 OR), yielded a relatively narrow results space ranging from roughly 1.1 to 1.4.
+Though my analysis showed one additional value of roughly 4.5, its plausibility is it to be determined.
+The outcome of 1.31 found by *team 23* was not included in the interval between the 1st and 3rd quartile, though its was part of the 95% confidence interval.
+It is, therefore, concluded that its specified covariates have an above averages effect i.e., are more influential than a random combination of covariates would have had.
+The distribution of the results space showed two peaks.
+To understand the distribution's underlying structure the specified covariate effects be analysed.
+The results showed that 4 out of 15 covariates are statistically significant, though their effects are small.
+Given the outcome measures showed values that are beyond the apparent influence of a single covariates it is, therefore, assumed that specific covariate interactions have significant effect on the outcome.
+The final volcano plot supports this hypothesis by clearly outlining two (perhaps a weak third) structure.
+Unfortunately, is identifying these underlying structures beyond the scope of this project.
 
- Due to time and resource limitation this project focused on pne
+ Taken together, considering the roughly 90% of all outcomes are significant it is no surprise that the original findings by *team 23* significant.
+In fact, it would have been more surprising if they had not found a significant effect.
+That being said, the effects is still strong than a seemingly random combination of covariates.
+The pressing question, therefore, is how did the authors arrive at this specific effect.
+IN PROGRESS.
 
-In order to assess the strength of the association covariation is hardly useful as covariate it does not have a standarised scale to it.
+ All materials used for this analysis are publicly available and all results should be reproducible and expendable.
+The biggest constraint of this project were time and computational resources.
+However, reasonable decisions were made to achieve the best possible compromised between rigour and feasibility.
+The constraints and the findings of the current project both yield great potential for futures research.
+The first step would running all possible covariates combinations, not just a sample.
+It would highly interesting if the observed patter is facilitated or vanishes.
+Moreover, it would be highly interesting to decipher the underlying covariate structure that is responsible the observed "streams".
+Based on the present findings I would assumes there is not a single covariate but rather a combination of them responsible.
+Finally, running not mixed-model logistic regressions but all other models used in the original study by @silberzahn2018.
 

### 2021-08-22 scripts/3_data_preparation.R
Adding codebook

@@ -4,16 +4,15 @@ library("tidyverse")
 library("plyr")
 library("lme4")
 
-# from raw data generate disaggregated data for logistic regression via 2_data_disaggregate.py
+# From raw data generate disaggregated data for logistic regression via 2_data_disaggregate.py
 dat = data.frame(fread(here::here("data", "2_disaggregated_data.csv")))
 
-# Silberzahn et al. 2018 list all covariates among them are two that need to be calculated: the cumulated number of cards received by a player and assigned by a referee.
-# The calculations of these covariates follows Morey and Wagenmarkers who were the team to use both covariates.
+# Silberzahn et al. 2018 list all covariates among them are two that need to be calculated: the cumulated number of cards received by a player and assigned by a referee. (More information in the codebook)
 dat = dat %>% group_by(player) %>% mutate(player_all_cards_received = sum(yellowCards + yellowReds + redCards))
 dat = dat %>% group_by(refNum) %>% mutate(ref_all_cards_assigned = sum(yellowCards + yellowReds + redCards))
 
-# select only the columns we will use in the analysis
-# adding all used covariates mentioned in Silberzahn et al. (2018) incl. main variables
+# Select only the columns we will use in the analysis
+# Adding all used covariates mentioned in Silberzahn et al. (2018) incl. main variables
 dat <- dat[, c("playerShort", "height", "weight", "birthday", "position",
                "club", "leagueCountry",
                "games",  "goals", "victories", "ties", 
@@ -21,34 +20,34 @@ dat <- dat[, c("playerShort", "height", "weight", "birthday", "position",
                "refNum", "refCountry", "refCount", "ref_all_cards_assigned",
                "meanIAT", "meanExp", "skintone")]
 
-# if we're on windows or mac you need to force the bias variables into numeric
+# If we're on windows or mac you need to force the bias variables into numeric
 dat$meanIAT <- as.numeric(as.character(dat$meanIAT))
 dat$meanExp <- as.numeric(as.character(dat$meanExp))
 
-# remove NA skintone data
+# Remove NA skintone data
 dat <- dat[!is.na(dat$skintone), ]
 
-# remove refs present in <22 dyads
-# reduces 373067 obs. by 16958
+# Remove refs present in <22 dyads
+# Reduces 373067 obs. by 16958
 dat <- dat[dat$refCount > 21, ]
 
[…]

### 2021-08-22 scripts/4_multiverse_analysis.R
Editing code

@@ -1,3 +1,9 @@
+# DISCLAIMER: The running times of the models included in the is script are extremely long (14 to 18 hours).
+# If you do not want to wait this long, you can load the models output data in line 240 (file: "4_model_outcomes_data.csv").
+# In this case please only refrain from running the section "Wrapper function, running models and output wrangling",
+# All other code is required to be run
+
+
 #...............................................# Set parameters #..............................................#
 
 # This script can be run in its entirety to reproduce the specification curve analysis for one population at a time. 
@@ -88,7 +94,7 @@ varList[["var_all"]] = var_all
 
 
 
-#.................. Prepare specifications: Create data frame from all combinations of covariates .........#
+#......................... Prepare specifications: create formulas and draw random sample .................#
 # Create TRUE/FALSE matrix for every combination of covariates
 specifications = data.frame(expand.grid(rep(list(0:1), length(covar))) == TRUE)
 
@@ -141,6 +147,9 @@ prop_ef = length(all_models[!all_models_index]) / (length(all_models[!all_models
 ranef_sample =  sample(formula_ranef, n_sample*prop_ranef, replace = FALSE)
 ef_sample = sample(formula_ef, n_sample*prop_ef, replace = FALSE)
 
+
+
+#........................# Wrapper function, running models and output wrangling #.........................#
 
 # Function extracting estimates, se, z-values and p-value from models.
 # Use as wrapper function around the models to extract relevant statistics only and reducing the output size.
@@ -220,9 +229,15 @@ df$ci_upper_oddsratio = with(df, exp(ci_upper))
 sum(df$below_alpha == "Significant") / nrow(df) # = 89.7%
 
 
-# Save a potential interim outcome
-# write.csv(df, here::here("data", "4_model_outcomes_data.csv"), row.names = FALSE)
+# Save the models output data
+write.csv(df, here::here("data", "4_model_outcomes_data.csv"), row.names = FALSE)
 
+
+
+#...............................# Process model output of visualisations #.................................#
+
+# In case models, were not run, load their outcomes
+# df = read.csv(here::here("data", "4_model_outcomes_data.csv"))
 
 # Build data frame that includes the specifications and the analysis results
 analysed_specifications = inner_join(specifications, df, by = "formula")

### 2021-08-22 dissertation/msc_dissertation.Rmd
Cleaning up code

@@
   |
-  | MSc Dissertation - DRAFT
+  | MSc Dissertation
 author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
@@
 $formula.ranef; formula.ef$ $\leftarrow$ Separate formulas based on including a random-structure \\
-$samlple.ranef; sample.ef$ $\leftarrow$ Random, without replacement, sample from $formula.ranef; formula.ef$
+$samlple.ranef; sample.ef$ $\leftarrow$ Random, without replacement, sample from $formula.ranef; formula.ef$ \\
+$wrapper function$ $\leftarrow$ Extracting model statistics and stores them in a list\\
+\textbf{tapply}
 \BlankLine

### 2021-08-22 scripts/4_multiverse_analysis.R

@@ -258,6 +258,9 @@ analysed_specifications = analysed_specifications %>% filter(estimate_oddsratio 
 
 
 # Assigning significance labels
+## Alpha and z have been defined earlier, but in case that section has not been run, it's redefined here
+alpha_level = 0.05
+z = 1.96 # 95% CI
 analysed_specifications$below_alpha = with(analysed_specifications, ifelse(p_value < alpha_level, "Significant", "Non-significant"))
 
 # Selecting relevant columns for a better overview (covariates and "estimate_oddsratio")
@@ -471,7 +474,7 @@ ggsave(here::here("figures", "sca_plot.png"), sca_plot, width = 10, height = 6)
 #..............................................# Volcano plot #............................................#
 
 
-volcano_plot = dat_models_specifications %>% 
+volcano_plot = analysed_specifications %>% 
   filter(estimate_oddsratio < 4) %>% 
   ggplot(aes(estimate_oddsratio, log10(p_value))) +
   geom_point(aes(colour = below_alpha), alpha = 0.3) +

### 2021-08-22 scripts/4_multiverse_analysis.R
Update 4_multiverse_analysis.R

@@ -1,7 +1,7 @@
-# DISCLAIMER: The running times of the models included in the is script are extremely long (14 to 18 hours).
-# If you do not want to wait this long, you can load the models output data in line 240 (file: "4_model_outcomes_data.csv").
-# In this case please only refrain from running the section "Wrapper function, running models and output wrangling",
-# All other code is required to be run
+# DISCLAIMER: The running times of the models included in this script are extremely long (14 to 18 hours).
+# If you do not want to wait this long, you can load the models output, "4_model_outcomes_data.csv", in line 240,
+# please uncomment that line. Please only refrain from running the section "Wrapper function, running models and output wrangling",
+# line 152 to 233, all other code is required to be run.
 
 
 #...............................................# Set parameters #..............................................#

### 2021-08-22 dissertation/msc_dissertation.Rmd
Editing dissertation

--- 2021-08-16 dissertation/msc_dissertation.Rmd
@@
 subtitle: |
-  | *Defining the results space in @silberzahn2018 through model specification*
+  | *Assessing results space and structure of @silberzahn2018 through model specification*
   |
-  | Master's thesis - DRAFT
+  | MSc Psychological Research Methods with Data Science
+  | Dissertation
 author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
-linestretch: 1.5
 output:
@@
 header-includes:
-  - \usepackage[ruled,vlined]{algorithm2e}
+- \usepackage[ruled]{algorithm2e}
 - \usepackage{float}
@@
 
-knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')
+knitr::opts_chunk$set(echo = FALSE, fig.pos = 'H')
 
@@
 ## Theoretical framework
-     The Covid-19 pandemic has reaffirmed the need for sound scientific research to inform decision-making on a scientific and societal level [@collins2021].
+     The Covid-19 pandemic has reaffirmed the crucial relevance of sound scientific research for political and societal decision-making [@collins2021].
 Rigorous research builds upon a systematic and well-reasoned approach to solving a research problem.
-Based on available literature, a falsifiable research question is defined and a corresponding hypothesis is developed.
-An appropriate study is then designed and conducted.
-To draw inferences from the gathered data, statistical models are developed and the effect of each variable is assessed.
-Every one of these steps is influenced by the decisions researchers make, which are known as *researcher degrees of freedom* [RDF, @simmons2011; @wicherts2016].
-Usually there are many feasible analytical strategies to answering a research question and none of them are inherently right or wrong  [@carp2012].
-This often creates uncertainty, for example, about what covariates to include and how to model them, which in turn leads to inconsistent findings [@patel2015; @ioannidis2008].
-Recently, efforts have been made to better understand the scope of variation induced by different analytical strategies.
+Based on current knowledge, researchers define a research question and develop a hypothesis.
+To test the hypothesis, they design and conduct a research study that yields data.
+To draw conclusions from the data, researchers apply statistical models and assess how different variables have influenced the data.
+Each of these steps is influenced by the researchers’ decisions, which are known as *researcher degrees of freedom* [RDF, @simmons2011; @wicherts2016].
+In most cases, there is not only one but many feasible analysis strategies to answer a research question [@carp2012].
+This ambiguity often creates uncertainty and inconsistency.
+Specifically, researchers are often uncertain about which covariates to include and how to model them, which leads to inconsistent findings [@patel2015; @ioannidis2008].
+Recently, efforts have been made to better understand how different analysis strategies influence research results.
 
- These efforts include crowdsourcing approaches to data analysis [@silberzahn2018; @botvinik-nezer2020].
-Its premise is to have a large number of researchers team up into to smaller, independent groups to investigate the same research question based on the same dataset.
-For instance, @silberzahn2018 had 29 teams investigate the effects of a football player's skin colour on the odds of being sent off the field.
-The variation between the analytical strategies was substantial.
+ Crowdsourcing is one approach to better understand the influence of analysis strategies on research findings [@botvinik-nezer2020; @silberzahn2018].
+A large number of researchers team up into smaller, independent groups to investigate the same research question based on the same dataset.
+For instance, in @silberzahn2018, 29 teams investigated the effects of a football player’s skin colour on the odds of being sent off the field.
+The variation between the analysis strategies was substantial.
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect.
-The authors also controlled for researchers' prior believes and experience as well as peer-rated analysis quality, none of which accounted for the variation of results.
+Twenty teams found a statistically significant effect of skin colour on the risk of being sent off the field.
+The authors also controlled for researchers’ prior believes and experience as well as peer-rated analysis quality, but these factors did not account for the variation of results.
 @botvinik-nezer2020 made similar observations.
-Crowdsourcing projects excels at emphasising the substantial impact of different analytical strategies, but they are also extremely time and personnel intensive.
-The two mentioned studies lasted between 2 to 3+ years, and included 61 and 180 analysts, respectively.
-Additionally, despite covering a more diverse set of analytical strategies than a single conventional analysis, crowdsourcing is limited by the number of participating teams.
-It, therefore, still only covers a few selected strategies, not all possible ones.
+Such crowdsourcing approaches are well suited to show how different analysis strategies influence research findings, but they are extremely time and resource consuming.
+For instance, it took two and more than three years and 61 and 180 analysts, respectively, to perform the two studies mentioned above.
+Furthermore, although crowdsourcing approaches can cover more analysis strategies than a single conventional analysis, they are limited by the number of teams.
+They, therefore, cover a specific selection but not all possible strategies.
 
- @silberzahn2018 and @botvinik-nezer2020, therefore, suggest using multiverse analysis (also known as specification-curve analysis) as an alternative to crowdsourcing.
-This approach requires identifying and running all plausible analytical strategies.
-Plausible strategies are defined as statistically valid, non-redundant, tests appropriate to the research question.
-Their aggregated results are then used to make inferences about the research question.
-Despite being statistically more complex and computationally more intense, this approach has the advantages of only needing one, or ideally a couple, researchers.
-Moreover, a researcher's inherent strategy bias is neutralised and noise is made transparent [@simonsohn2020].
-Multiverse approaches are therefore well suited for assessing the scope of variation induced by different analytical strategies.
-However, one aspect has been, to my knowledge, neglected thus far.
-To properly assess the scope of variation it needs to be contextualised.
-That is, the space in which it can vary needs to be defined in order to understand to what degree the scope covers the space.
-The concept of statistical association provides a suitable analogy.
-Covariation and correlation both indicate the direction of relationship (positive, null, negative), however, only correlation allows for assessing the association's strength.
-Correlation is the standardised version of covariation, and only knowing the its limits gives the correlation coefficient meaning.
-The same principle applies to analyses on a broader level.
+ Another approach to assess the effects of analysis strategies on research findings is multiverse analysis (also known as specification-curve analysis, [@simonsohn2020].
+This approach has been proposed by @botvinik-nezer2020 and @silberzahn2018 as an alternative to crowdsourcing.
+In this approach, all reasonable analysis strategies are identified and performed.
+*Reasonable strategies* are defined as all statistically valid, non-redundant tests that are appropriate for the research question.
+The aggregated results of all reasonable analysis strategies are used to make inferences about the research question.
+Although the approach is statistically complex and computationally intense, it can be performed by one or a few researchers.
+Moreover, a researcher’s inherent strategy bias is neutralised, and noise is made transparent [@simonsohn2020].
+Multiverse approaches are therefore well suited to better understand how different analysis strategies influence research results.
 
 ## Current project
-     Taken together, researchers make many analytical decisions throughout conducting a study introducing so-called *researcher degrees of freedom*.
-Crowdsourcing analysis presents a strategy to understand their induced variation, but is impractical due to its time and personnel requirements, and is still limited by the number of participating teams.
-Multiverse analysis present a promising, low-personnel, but computationally intense alternative to assessing the scope this variation.
-Despite observing substantial variation in crowdsourced data analysis projects, little is known about the extend to which these approaches cover the full range of possible results.
-Determining the full results space and its underlying mechanisms is an important exercise to understand the impact of analytical strategies.
-To my knowledge no previous study has attempted to define and understand a study's entire results space.
-In this project, running all possible analytical strategies, I, therefore, aim to define and understand the results space of @silberzahn2018.
-By doing so, I hope further the understanding of how an analytic strategy affects its result.
+     Researchers make analytical decisions that introduce *researchers degrees of freedom*.
+Crowdsourcing analysis and multiverse analysis are promising tools to assess the variation of results induced by analysis strategies.
+However, these approaches do not assess the full range of results obtained by all possible analysis strategies but usually focus on a subset of reasonable analysis strategies.
+Therefore, it is unclear how the spaces of all reasonable and all possible analysis strategies relate to each other.
+In particular, it is unknown whether the space of all reasonable results covers the full range of possible results or whether it covers only a narrow sub-space of possible results.
+In this project, all reasonable and all possible analysis strategies were performed to define and understand the results space of @silberzahn2018.
+Furthermore, ways to visualise the space and its structure intuitively and succinctly were proposed.
+The project is, thus, intended to further the basic understanding of how analysis strategies influence research results and to provide tools for further studies on this topic.
 
@@
 ## Analysis plan
-     The project's objective is to define the results space of the @silberzahn2018 to draw conclusions about the mechanisms of multiverse analysis.
+     The project's objective was to define the space and the structure of all possible results of @silberzahn2018.
 The results space refers to the numerical interval between the lowest and highest possible outcome.
-It is created by running every possible analytical strategy.
-Hence, it is closely related to assessing an effect's robustness.
-Robustness refers to an effect's consistency under different model specifications.
-@patel2015 developed a standardised approach to assessing an effect's robustness which they termed assessing the *vibration of effect*.
-I will therefore use this standardised approach to define the results space.
+It is created by running every possible analysis strategy.
+Hence, it is closely related to assessing the robustness of an effect.
+Robustness refers to the consistency of an effect under different model specifications.
+@patel2015 developed a standardised approach to assess an effect's robustness.
+I used this standardised approach to define the results space.
 The approach essentially comprises two parameters: the statistical models and the covariates (or control variables).
-In @silberzahn2018 the analysts used numerous different statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression.
-In total there were 29 different modelling approaches.
+In @silberzahn2018, the analysts used numerous statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression.
+In total, there were 29 different modelling approaches.
 Additionally, each team used a different set of covariates.
-Across all teams there 15 covariates used.
-This gives $2^n$ i.e., $2^{15} = 32,768$ possible combinations of covariates.
-Adding all modelling possibilities to the equation gives a total of $29 * 2^{15} = 950,272$ combinations to run.
-This number exceeds the project's scope, hence, it needs to be reduce to a more manageable count.
-I am, therefore, focusing on one modelling approach.
-In particular, on the approach that produced the median outcome of all analyses.
-The median, being the middle number of any set of values, is a reasonable starting point in order to estimate the results space.
-Nevertheless, even focusing on one analytical approach still leaves $1 * 2^{15} = 32,768$ possibilities.
-Due to computational limitations I will first run a sample of 200, followed by a sample of 1,000 combinations.
-If there's a substantial difference between the median and the spread I will run more models, if not, I will make the assumption that 1,000 reasonably estimates all ~33k combinations.
+Across all teams, 15 covariates were used, resulting in $2^n$, i.e., $2^{15} = 32,768$ possible combinations of covariates.
+Adding all modelling possibilities to the equation yields a total of $29 * 2^{15} = 950,272$ possible analysis strategies.
+As this number of strategies exceeds the current project's resources, it was reduced to a more manageable number.
+Therefore, the project focused on the modelling approach that produced the median outcome of all analyses.
+The median, being the middle number of a given set of values, is a reasonable starting point to estimate the results space.
+However, even focusing on one analysis approach leaves $1 * 2^{15} = 32,768$ possibilities.
+Due to computational limitations, a random sample of 1,000 analysis strategies were conducted.
 
- In @silberzahn2018 the median outcome was produced by @team23 which will hereinafter be referred to as *team 23* due to it's designated team number in the original study.
+ In @silberzahn2018, the median outcome was produced by @team23, which will hereinafter be referred to as *team 23* due to its team number in the original study.
 *Team 23* first transformed the data and then conducted a mixed-model logistic regression.
-The first step of this project will be replicating the transformation and the analysis.
-Replicating other researchers analyses has the benefit of checking their work and, if results are indeed replicated, increasing confidence in them.
-It also ensures that this project has the same starting point as *team 23* did.
-Given the team made all their scripts publicly available I expect this step to be straightforward.
-Next is drawing a random sample of covariates, without replacement.
-"Without replacement" ensures all covariate combinations are unique in the sample.
-The covariates will be appended to the base (or core) variables.
-Base variables are those variables that are primarily assessed to answer the research question.
-In this case whether a football player's skin colour affects the odds of being sent off the field.
-*Team 23* defined two interaction terms as the base: "skin tone *X* implicit bias" and "skin tone *X* explicit bias".
-(The variables are described in the "data" section.) Hence, all models will have the following structure:
+The first step of the present project was to replicate their transformation and analysis.
+Such replication can increase the confidence in the previous and the current approach.
+Moreover, it ensures that this project has the same starting point as *team 23*.
+As *team 23* made all scripts publicly available, this step was straightforward.
+The next step of the project was to define the results space and its structure.
+To this end, a random sample of covariates without replacement was drawn.
+“Without replacement” ensured that all covariate combinations were unique in the sample.
+The covariates were appended to the base (or core) variables.
+Base variables were those variables that were primarily assessed to answer the research question.
+In this case, the research question was whether a football player’s skin colour affects the odds of being sent off the field.
+*Team 23* defined two interaction terms as the base: “skin tone X implicit bias” and “skin tone X explicit bias.” (The variables are described in the “data” section.) Hence, all models had the following structure:
 $$
@@
 $$ 
-While running each model, the relevant statistics will be extracted.
-Those are the coefficient (or effect) of skin tone, its standard error, test-statistic and p-value.
-Just as *team 23* did I will then calculate the 95% confidence intervals (CI).
-The estimates and their CIs are then transformed to odds ratios (OR) through exponentiating them to the power of two.
-Odds ratios quantify the strength of association between two variables.
-If greater than one the dependent variable is more likely to occur given the independent variable, if lower than one it less likely.
+The relevant outcome parameters of each model were extracted.
+These parameters were the coefficient (i.e., effect) of skin tone, its standard error, test statistic and p-value.
+Similar to *Team 23* I calculated the 95% confidence intervals (CI).
+The estimates and their CIs were transformed to odds ratios (OR) through exponentiating them to the power of two.
+OR quantify the strength of association between two variables.
+An OR greater than one indicates that the dependent variable is more likely to occur given the independent variable, if it is lower than one it is less likely to occur.
+Eight OR outliers were excluded from visualisation.
+Four approached infinity, one was in the lower million range, one was 150, and a final one was 4.6.
 
- These odds ratios will be visualised to assessed through a conventional specification curve plot, a rain-cloud plot and scatter plot. The following describes the three graphs in more detail:
+ To provide insights into the results space and its structure, the OR were visualised and assessed by three complementary plots, (i) a specification curve plot, (ii) a raincloud plot, (iii) and a scatter plot.
+The following describes the three plots in more detail:
 
-(i) The specification curve plot was developed to be a descriptive plot i.e., raw data is being plotted without any aggregation done.
-Its objective is to describe the results space while allowing the reader to identify the specified covariates for each outcome [@simonsohn2020].
-It therefore has two vertically stacked components.
-The top part shows a sorted scatter plot.
-Its horizontal axis shows the models specifications and its vertical axis displays the outcome measure.
-The scatter points are sorted from lowest to highest outcome measure.
+(i) The specification curve is a descriptive plot that shows raw outcome data without any aggregation.
+Its objective is to describe the results space while allowing the reader to identify the model specifications for each outcome [@simonsohn2020].
+It has two vertically stacked components.
+The upper component is a sorted scatter plot.
+Its horizontal axis lists the specifications, and its vertical axis displays the outcome measure.
+The data points are sorted from lowest to highest outcome measure.
 This way the lowest outcome measure is on the bottom left corner and the highest in the top right one.
-The bottom part of the plot represents a table.
-As in the top plot the horizontal axis holds the model specifications, the vertical axis lists the covariates.
-This arrangement allows each cell to specify the presence or absence for each covariate in a given specified model.
-For the top and bottom plot to work together it is imperative that the horizontal axes are identically arranged.
-If so, for every point in the top plot (i.e., for every outcome) the specified covariates can be seen in the bottom plot.
-For an example check Figure \ref{fig:sca_plot}.
-As the current project has 1,000+ specifications it becomes very hard to identify specific covariates given the space limitations.
-Hence, I to had come up with different more intuitive ways of visualisation.
+The lower component of the plot is a table.
+Similar to the upper component, the horizontal axis lists the specifications, while the vertical axis lists the covariates.
+This arrangement allows each cell to specify the presence or absence of each covariate in a given model specification.
+For the integration of information from the upper and lower components, the horizontal axes are identically arranged.
+Thus, for each point in the upper component (i.e., for every outcome), the specified covariates can be seen in the lower component (see Figure \ref{fig:sca_plot} for an example).
+However, with the more than 1,000+ specifications of the current project, it was hard to identify specific covariates.
+Hence, the specification curve was complemented with other plots.
 
-(ii) The rain-cloud plot was developed by @allen2021.
-Its objective is to give an unbiased, transparent view on the raw data.
-It therefore combines a density, box and scatter plot.
-If stacked vertically from top to bottom, respectively, they look like a cloud with rain drops, hence, its name.
-The advantage of this plot is to be able to assess the raw data including key statistics (median, 25th and 75th quartiles and CIs) as well as the probability density all in one succinct plot.
-The current project seeks to define and describe a results space the rain-cloud plot is, therefore, the ideal choice of graphs.
-(For an example check the results section.) Unfortunately, the plot does not allow to identify the covariates giving rise to the outcomes or their specified effects.
-As the project also seeks to understand the mechanisms of multiverse analysis it important to get insights about the specified covariates effects.
-Therefore a third plot is developed which is to be assessed in tandem with this plot.
+(ii) The raincloud plot @allen2021 combines a probability density plot, a box plot, and a scatter plot to give an unbiased, transparent view of the raw data.
+If the three plots are stacked vertically from top to bottom, they look like an eponymous cloud with rain drops.
+The strength of the raincloud plot is that it visualizes the raw data, summary statistics (median, 25th and 75th quartiles and CIs) and the probability density in a single plot.
+As the current project seeks to define and describe a results space, the raincloud plot is an ideal tool (for an example check Figure \ref{fig:rain_cloud_plot}).
+However, the plot does not allow to identify which covariates cause which outcomes.
+As the project also sought to understand the structure of the results space, it was essential to gain insights into how specific covariates influenced the outcomes.
+Therefore a third plot which complemented and extended the raincloud plot was developed.
 
-(iii) The third and final plot is a sorted scatter plot.
-While there is nothing special about the plot itself its content is important.
-Given the very high number of models that are run, it does not make sense to assess each model and its effects in isolation.
-Instead the effects for each covariates need to be aggregated.
-Hence, the goal of this plot is to visualise the specified effect for each covariate.
-Prior to doing so, the covariates first need to be aggregated and their specified effects need to be calculated.
-To this end, each covariate is recoded into a binary factor: included in the model yes/no.
-This newly defined factors are than used as an independent variables in an ANOVA.
-The previously calculated odds rations are used as dependent variables.
-The estimates of the fitted model are the specified effect for each covariates.
-These are visualised similarly to the top part of the specification curve plot.
-The covariates are on the horizontal axis, will the odds ratios are on the vertical axis.
-Here too, the lowest outcome is in the bottom left corner and the highest outcome in the top right one.
-Additionally the point are coloured indicating a statistically significant effect (or non-significant).
+(iii) The third plot was a sorted scatter plot.
+The goal of this plot was to visualise the effects of each covariate.
+Given the large number of models, assessing all models and their effects individually did not make sense.
+Hence, the effects of each covariate were aggregated.
+To this end, each covariate was recoded into a binary factor: included in the model yes/no.
+These newly defined factors were then used as independent variables in an ANOVA.
+The previously calculated OR were used as dependent variables.
+The estimates of the fitted model were the specified effect for each covariate.
+These were visualised similar to the top part of the specification curve plot.
+The covariates were on the horizontal axis, while the OR were on the vertical axis.
+The lowest outcome was in the bottom left corner, and the highest was in the top right corner.
+Additionally, the colours of the points indicated the statistical significance of the effects.
 
- Taken together, the current project's objective is to define the results space of @silberzahn2018 in order to make inferences about the underlying mechanisms of multiverse analysis.
-The results space is created by running all possible model and covariate combinations.
-Applied to @silberzahn2018 this means running roughly 1M models.
-Due to time and computational limitations, I am focusing on the model that produced the median outcome of all analytical strategies.
-Hence, I will run a mixed-model logistic regression with 1,000 randomly sampled covariate combinations.
-For each model relevant statistics will extracted and visualised.
-The conventional specification curve plot does not deal well with a large number of model specifications.
-Therefore, I will produce a rain-cloud and scatter plot, which visualise the results space (and its distribution) as well as the specified covariate effects, respectively.
-For the latter, I will first have to run an additional statistical model to calculate the specified effects.
-These graphs will give insight into the results space, its building blocks and ultimately into the mechanisms of multiverse analysis.
+ In summary, the current project aimed to define the space and structure of the results of @silberzahn2018.
+The space of possible results is defined by running all models with all possible covariate combinations.
+Applied to @silberzahn2018 this would result in about 1M models.
+Due to time and computational limitations, the project focused on the modelling approach that produced the median outcome of all analyses.
+Hence, a mixed-model logistic regression with 1,000 randomly sampled covariate combinations was run.
+For each model, relevant parameters were extracted and visualised.
+The specification curve is not ideal for large numbers of model specifications.
+It was, therefore, complemented and extended by raincloud and scatter plots, which visualise the results space and the specified covariate effects, respectively.
 
@@
      This section provides a brief overview of the code and explains the reasoning behind it.
-As mentioned in the analysis section the first step was replicating the data transformation and analysis of *team 23* [@team23].
+First, the data transformation and analysis of *Team 23* [@team23] were replicated.
 The team made their project folder publicly available.
-It contains three scripts relevant to this project: the data exploration, transformation and analysis.
-After duplicating their project folder on my local hard drive all scripts run without issues (only the working directories needed adjustment).
+It contains three scripts relevant to the current project: data exploration, transformation, and analysis.
+After duplicating their project folder on the local hard drive, all scripts ran without issues (only the working directories needed adjustment).
 The data exploration included the reasoning behind transforming the data and some cleaning.
 The data transformation restructured the data into a more intuitive format (more details on the topics of exploration and transformation are in the data section).
-The analysis script prepared the data by assigning variables classes (factors, numerical or boolean) and standardised a few variables i.e., they were centred around the mean.
+The analysis script prepared the data by assigning variable classes (factors, numerical or boolean) and standardised a few variables, i.e.
+they were centred around the mean.
 Finally, the models were specified.
-The team ran both frequentist and Bayesian models, however, this project covers the frequentist approach, only.
+The team ran both frequentist and Bayesian models.
+However, the current project focused on the frequentist approach.
 
- My code is based on @haessler2020 and @patel2015.
-The rough structure of the code is shown in the pseudocode table (Figure X).
-Here, I am briefly motivating two analytical choices I made.
-For a closer look please check the commented code itself:
+ The code was based on @haessler2020 and @patel2015.
+The basic structure of the code is shown in the pseudocode table (see Algorithm \ref{alg:pseudocode}).
+Here, two important analytical decisions are briefly motivated.
+For more details, please refer to the commented code:
 
@@
 *Team 23* ran a mixed-model logistic regression.
-This model has two relevant components to this explanation: fixed effects and random effects.
-Fixed effects are those that are consistently observed in different situations because the construct is of direct interest to the research question.
+This model has two relevant components: fixed effects and random effects.
+Fixed effects are consistently observed in different situations because the construct is of direct interest to the research question.
 In this case, it is, for example, skin tone.
-Independent of the player, the game or the league skin tone is always observed.
-Their counter parts are random effects which change between situations.
-A player, for example, is just one "unit" a measurement, and is himself not directly relevant to answering the research question.
-The function used by *team 23* requires to specify a random effect.
-However, given the project seeks to sample from all possible covariate combinations, a random effect is not always included.
-Hence, the "non-random" counter part to this function needs to be utilised.
-For this project, it was important to check that both functions use the same estimation method (maximum likelihood or restricted maximum likelihood estimation) to calculate the statistics.
-It was important to check to ensure the outcomes are comparable.
+Independent of the player, the game or the league, skin tone is always observed.
+Their counterparts are random effects which change between situations.
+A player, for example, is just one “unit” of measurement and is herself/himself not directly relevant to answering the research question.
+The function used by *Team 23* requires specifying a random effect.
+However, given that the current project sought to sample from all possible covariate combinations, a random effect was not always included.
+Hence, the “non-random” counterpart to this function was used.
+Therefore, this project ensured that both functions used the same estimation method (maximum likelihood or restricted maximum likelihood estimation) to ensure that the outcomes are comparable.
 
-(ii) No multiple comparison.
-This is based on the fact that in statistics null hypotheses are one rejected if the probability of finding a false positive is below 5% (known as the significance level, $\alpha$).
-The more statistical tests a run, the more likely it is that one of those tests is indeed a false positive.
-Hence, the are techniques that account of this issue.
-It is usually good practice to account for those "multiple comparisons".
-Here, I refrained from doing so though.
-This project seeks to simulate multiple researchers running different models.
-Therefore, for the the purpose of this project all models are run independently by individual researchers.
-Those would not have knowledge of the other analyses and, hence, would not account for them.
-To main the highest possible ecological validity, I therefore did not account of multiple comparisons.
+(ii) No correction for multiple comparisons.
+In statistics, it is common practice to reject a null hypothesis if the probability of finding a false positive is below 5% (known as the significance level, $\alpha$).
+The more statistical tests are run, the more likely it is that a result of at least one of the tests is a false positive.
+Hence, it is good practice to account for those “multiple comparisons.” The current project, nevertheless did not correct for multiple comparisons.
+This project sought to simulate multiple researchers running different models.
+Those researchers would not know the other analyses and, hence, would not account for them.
+To maintain the highest possible ecological validity no correction for multiple comparisons was made.
+\newline
 
 \begin{algorithm}
-\caption{Pseudocode overview--in progress}
-\DontPrintSemicolon
-\SetAlgoLined
-\KwData{Prepared data based on team 23 script}
-\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} 
-$variables$ $\leftarrow$ Define dependent, base variables and covariates \\
-$specifications$ $\leftarrow$ Use $variables$ to create matrix containing all possible covariate combinations \\
-$formula$ $\leftarrow$ Paste $specifications$ by row and append as column to $specifications$ \\
-$formula.ranef; formula.ef$ $\leftarrow$ Separate formulas based on including a random-structure \\
-$samlple.ranef; sample.ef$ $\leftarrow$ Random, without replacement, sample from $formula.ranef; formula.ef$
+\caption{Outlining high-level structure of the code. "Ran" in e.g. "FormulaRanEf" refers to "random-structure". "Ef" refers to "effect".}
+\label{alg:pseudocode}
+\SetKwInOut{Output}{Output}
+\textbf{data} $\leftarrow$ prepared data based on team 23 \\
   \BlankLine
- initialization\;
-  \While{not at end of this document}{
-    read current\;
-    \eIf{understand}{
-      go to next section\;
-      current section becomes this one\;
-      }{
-      go back to the beginning of current section\;
-      }
-    }
-\KwResult{List containing model statistics}
+  $Variables$ $\leftarrow$ Define dependent ($dv$), base variables ($basevar$) and covariates ($covar$) \\
+  $Specifications$ $\leftarrow$ Use $Variables$ to create matrix containing all possible covariate combinations \\
+  \BlankLine
+  $Formula$ $\leftarrow$ Paste $Specifications$ by row and append as column to $Specifications$ \\
+  $FormulaRanEf; FormulaEf$ $\leftarrow$ Separate formulas based on including a random-structure \\
+  \BlankLine
+  $SampleRanEf; SampleEf$ $\leftarrow$ Unique, random sample from $FormulaRanEf; FormulaEf$ \\
+  \BlankLine
+  $ExtractorFunction$ $\leftarrow$ Extracting model statistics and store them in object \\
+  \BlankLine
+  $OutputsRanEf$ $\leftarrow$ \textbf{lapply(data, $ExtractorFunction$(glmer($dv$, x = $basevar + SampleRanEf$)))} \\
+  $OutputsRanEfEF$ $\leftarrow$ \textbf{lapply(data, $ExtractorFunction$(glm($dv$, x = $basevar + SampleEf$)))} \\
+  \BlankLine
+  $ModelOutputs$ $\leftarrow$ Merge $OutputsRanEf$ and $OutputsEF$ \\
+  $ModelOutputs$ $\leftarrow$ Compute odds ratios and confidence intervals \\
+  \BlankLine
+  $JoinedData$ $\leftarrow$ \textbf{LeftJoin($Specifications$, $ModelOutputs$, by = $Formula$)} \\
+  \BlankLine
+  \Output{Rain cloud plot}
+  \BlankLine
+  $CovariateEffects$ $\leftarrow$ \textbf{LinearModel($OddsRatios$, x = $Specifications$, data = $JoinedData$)} \\
+  \BlankLine
+  \Output{Covariate effects plot}
+  \Output{SCA plot, sorted by $CovariateEffects$}
+  \BlankLine
+  \Output{Volcano plot}
 \end{algorithm}
@@
  The code is written in R (Version 1.4.1103) on macOS Big Sur (Version 11.4) and can be retrieved from my [GitHub repository](https://github.com/sebastianplnr/msc_dissertation_project).
-The code from *team 23* [@team23] can be retrieved from their [OSF repository](https://osf.io/akqt4/).
+The code from *Team 23* [@team23] can be retrieved from their [OSF repository](https://osf.io/akqt4/).
 Following R packages were used *here 1.0.1* [@here], *data.table 1.14.0* [@datatable], *tidyverse 1.3.1* [@tidyverse], *lme4 1.1-27.1* [@lme4], *pbmcapply 1.5.0* [@pbmcapply], *PupillometryR 0.0.3* [@PupillometryR] and *cowplot 1.1.1* [@cowplot].
@@
      The data was retrieved from @silberzahn2018.
-It contains information about football players, their encounters with referees and the received cards (yellow, yellow/red and red).
-Moreover, a player's position, age, club, league country, victories, ties, defeats and goals.
-In addition, a skin tone rating based on two independent raters was included.
-The referees are numerically coded to protect their identify.
-The referees origin countries are also included as well as an implicit and explicit racism bias scores for their respective countries.
-The exploratory data analysis (EDA) of *team 23* showed that each row of the dataset represents a unique player-referee combination listing all their encounters as well as the couple's total number of received/assigned cards.
-The team states that it preferred a different data format.
-More specifically, a format where each player-referee encounter is reflected by one row.
-This way each encounter had a maximum of one red card.
-To achieve this format the data had to be "disaggregated" i.e., transformed.
-Further EDA showed that receiving a red card is highly unlikely, 0.008%, i.e., data is highly skewed.
-This property informed the teams decision to use a type of logistic regression (a statistical modelling technique equipped to deal with skewed/binary distributions).
-Finally, the team excluded all referee that did have at least 22 encounters.
-Their reasoning was a game includes (at least) 22 players, 11 players per team, if this were not the case referee most likely did not officiate in one for the four leagues of interest (England, Germany, Italy and Spain).
-This step excluded roughly 66% of the referees but retained 97.4% of the cases.
+It contained information about football players, their encounters with referees and the received cards (yellow, yellow/red and red).
+Moreover, it included a player’s position, age, club, league country, victories, ties, defeats, and goals.
+In addition, a skin tone rating based on two independent judges was included.
+The referees were numerically coded to protect their identity.
+The referees’ countries of origin were also included as well as implicit and explicit racism bias scores for their respective countries.
+The exploratory data analysis (EDA) of *Team 23* showed that each row of the dataset represents a unique player-referee combination listing all their encounters as well as the couple’s total number of received/assigned cards.
+*Team 23* stated that it preferred a different data format where each player-referee encounter is reflected by one row.
+This way, each encounter had a maximum of one red card.
+To achieve this format, the data had to be transformed, i.e.
+disaggregated.
+Further EDA showed that receiving a red card was highly unlikely ($p = 0.008%$), i.e.
+the data were highly skewed.
+The team, therefore, used a logistic regression which is a statistical modelling technique equipped to deal with skewed/binary distributions.
+Figure \ref{fig:var_plots} shows an overview of the properties of the core variables used in their analysis.
+
+ Finally, the team excluded all referees who did not have at least 22 encounters with players.
+Every football game includes (at least) 22 players, 11 players per team.
+If a referee has less than 22 player-encounters, there are missing cases.
+According to team 23, this was mostly the case in referees officiating games of minor leagues.
+As they focused on the major European leagues (England, Germany, Italy, and Spain), referees with less than 22 player-encounters were excluded.
+This step excluded about 66% of the referees but retained 97.4% of all player-referee combinations.
 The final dataset used for the current project contained 335,537 observations and 19 variables.
-Figure \ref{fig:var_plots} shows the properties of the variables relevant to current project's analysis \newline
+\newline
 
-```{r var_plots, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "A) Showing the strong skew of receiving no red to receiving a red card.
-B) Showing most players had a lower skin tone rating i.e., most players were on the whiter side.
-C) Showing there is not much variation in implict racial scores.
-D) Showing more variation in explitic reacial bias scores.
-The latter two were centred around the mean, hence, zero does not mean there is no bias."}
+```{r var_plots, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Team 23 base variables properties.
+A) There was a strong skew towards receiving no red card compared to receiving a red card.
+B) Most players had a lower skin tone rating, i.e., a brighter skin colour.
+C) The implicit racial bias scores hardly varied.
+D) The explicit racial bias scores varied slightly more than the implicit scores.
+(The latter two were centred around the mean; hence, zero does not mean there is no bias.)"}
 
@@
 ## Replicating *Team 23* analysis
-     As mentioned in the above section the first steps were replicating the *team 23* analysis using their provided scripts.
-The results are reproducible without any adjustment.
-Skin tone has a significant effect on the odds of being send off the field ($OR = 1.311$, 95%CI [1.099, 1.563], $p = 0.003$).
-This means while keeping all other variables constant for every unit increase in skin tone rating (darker skin tone), the odds of being send off the field increase by roughly 131%.
-The interaction terms of skin tone and implicit racial bias ($OR = 0.004$, 95%CI [0.000, 23.259], $p = 0.211$) as well as skin tone and explicit bias ($OR = 1.837$, 95%CI [0.493, 6.848], $p = 0.365$) are both non-significant.
-In accordance with the original analysis, there are also significant differences between leagues and positions as well as implicit racial bias scores.
-Explicit bias scores are on verge of being non-significant.
-This analysis only aimed at testing reproducibility, which proved to be true, further results details are forfeited.
+     First, the analysis of *Team 23* was replicated using their scripts.
+Skin tone significantly affected the odds of being sent off the field ($OR = 1.311$, 95%CI [1.099, 1.563], $p = 0.003$).
+This means that when keeping all other variables constant, for every unit increase in skin tone rating (darker skin tone), the odds of being sent off the field increased by about 131%.
+The interaction terms of skin tone and implicit racial bias ($OR = 0.004$, 95%CI [0.000, 23.259], $p = 0.211$) as well as skin tone and explicit racial bias ($OR = 1.837$, 95%CI [0.493, 6.848], $p = 0.365$) were both non-significant.
+In accordance with the original analysis concerning additional variables, there were significant differences of OR between leagues and positions as well as implicit racial bias scores.
+Explicit bias scores were on the verge of being non-significant.
+Thus, the results of team 23 were reproducible without any adjustment.
 
-## Specification curve
-     Shown in Figure \ref{fig:sca_plot}.
-The horizontal axes of the top and bottom chart show the specifications sorted based on their odds rations from lowest to highest.
-The odds ratios themselves are shown on the vertical axis of the top plot.
-The points in the top plot, which look like a line, each represent the outcome of one statistical model.
-Black points refers to statistical significant, red to non-significant outcomes.
-Overall are 89.7% of the outcomes significant.
-Except for the lowest outcome ($OR_{lowest} = 1.080$) are all points close to the their respective neighbours.
-This indicates that none of the specified models are responsible for a sudden increase in the outcome measure.
-The relatively smooth line suggest that sampling 1,000 specifications is, in fact, representative for all 33k models.
-The law of large numbers  The points can be divided into roughly three sections all non-significant ($\leq 57^{th}$), mixture of (non-)significance ($58^{th} - 247^{th}$) and all significance ($\geq 248^{th}$).
+## Exploring the results space - specification curve
+     Second, the specification curve was calculated to visualise the results space (Figure \ref{fig:sca_plot}).
+The horizontal axes of the top and bottom chart show the specifications sorted by their OR from lowest to highest.
+The OR are shown on the vertical axis of the top plot.
+The points in the top plot (which together look like a line) each represent the outcome of one statistical model.
+Black and red points refer to statistically significant and non-significant outcomes, respectively.
+Overall, 89.7% of the outcomes were significant.
+Except for the lowest outcome ($OR_{lowest} = 1.080$) all points were close to their neighbours.
+This indicates that none of the specified models was responsible for a sudden increase in the outcome measure.
 The bottom table lists all covariates on its vertical axis, sorted from most positive to most negative impact.
-For example, player has the strongest positive impact, while club has the strongest negative impact.
+For instance, the covariate player had the most substantial positive impact, while the covariate club had the most substantial negative impact.
 Each column of the table represents one mode.
 A coloured cell indicates the presence of the variable, an uncoloured cell its absence.
-Just like a above, black indicates the outcome measure was significant, red non-significance.
+Again, black and red colours indicate that the outcome measure was significant or non-significant, respectively.
 
- The top and bottom plot work in tandem.
+ The top and bottom plots work in tandem.
 Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table.
-Regular specification curve analyses focus on all *reasonable* specifications, hence, their number of specifications is much lower and one is, in fact, able to tell the different model specifications apart.
-However, the current project's goal was to define the results space, which requires all *possible* specifications.
-The specifications are therefore huddled extremely closely which does not allow the plot to functions in its supposed way.
-Thus, I create two alternative graphs.
+Regular specification curve analyses focus on all *reasonable* specifications; hence, the number of specifications is much lower than in the present study.
+With these lower numbers of specifications, it is possible to tease the different model specifications apart.
+However, the current project’s goal was to define the whole results space, i.e.
+all *possible* specifications.
+With such a high number of specifications, it is no longer possible to tease the different specifications apart.
+It is, however, interesting to look at the top and bottom rows of the table.
+The top row shows that the covariate player was particularly often included in models with a higher outcome.
+Conversely, the covariate club was particularly often included in models with a lower outcome.
 \newline
 
-```{r sca_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Convential specification curve.
-Each outcome (i.e., point) in the top corresponds to the indicated covariates in the bottom table.
-Due to the large number of specifications, this plot is not as informative as it is supposed to be.
-It is, therefore, merely making an argument for needing and alternative.", warning = FALSE}
+```{r sca_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Specification curve.
+Results space ranges from about 1.05 to 1.4.
+However, due to the large number of specifications, the specification curve does not allow for identifying the relevance of the different covariates.", warning = FALSE}
 
@@
 
-## Results space
-     Shown in Figure \ref{fig:rain_cloud_plot}.
-The horizontal axis shows the odds rations.
-From top to bottom the three component show the probability density distribution, the box plot and the raw data, respectively.
-The dashed line indicates the original results of *team 23*, which has also been the median outcome of all analytical strategies in @silberzahn2018.
+## Exploring the results space – raincloud plot
+     Third, the results space was further explored by calculating raincloud plots (Figure \ref{fig:rain_cloud_plot}).
+The horizontal axis shows the OR.
+From top to bottom, the three components show the probability density distribution, the box plot and the raw data.
 Based on the raw data points, the results space can be defined as the interval between 1.081 and 1.383.
-The 1st quatile equals to 1.206, the median equals to 1.248 and the 3rd quartile equals to 1.293.
-Therefore, as also evidenced by the probability density distribution, 50% of the lay withing a very narrow interval (between 1st and 3rd quartile).
-The outcome of *team 23* (1.31) is still included in the interval, but it is outside the middle 50% of the data.
+The 1st quartile was 1.206, the median 1.248 and the 3rd quartile 1.293.
+Thus, the interquartile range where 50% of the data were included was relatively small (1.206-1.293), indicating a low statistical dispersion.
+The dashed line indicates the original results of *Team 23*, which has been the median outcome of all analysis strategies in @silberzahn2018.
+The outcome of *Team 23* (1.31) was included in the results space but outside the middle 50% of the data.
+The CI calculated by team 23 ranged from 1.10 to 1.56.
+The lower bound of their CI was close to the lower margin of the results space of the present study.
+In contrast, the upper bound of their CI was notably higher than the present results space.
+This suggests a potential overestimating of the variability of the results in the original analysis.
 
- The probability density distribution also has an interesting shape, having two peaks.
-This suggests a particular covariate combination structure causing two focal points.
-Based on this graph alone though it is not possible to decipher the underlying structure though.
-What also seem to be more apparent in this graph than in the specification curve is that the non-significant values also seem to revolve round to focal points.
-One being around 1.13 and the other around 1.19.
-This also suggest some influential covariate or covariate structure that is cause non-significance.
-
- Finally, for transparency purpose, it also needs to be mentioned that eight points were excluded from the visualisation.
-Four of them were "infinite", one effectively approaching infinity, one being in the lowest million range, one being 150, and a final being 4.6.
-I specifically chose not to include the latter one in the graph as it would despite is relatively close numerical proximity skewed the distribution being at least a visual outlier.
+ The probability density distribution reveals another exciting feature of the data.
+The distribution has two peaks which suggest two latent covariate structures whose outcomes centre around two points.
+Those two focal points were located at about 1.13 and 1.19.
+Based on this graph alone, it is impossible to decipher the latent structures, i.e.
+determine the responsible covariates or combinations of covariates.
+To test the hypothesis of two latent structures, it would be interesting to observe how the distribution evolves when more specifications are run.
+In the presence of latent structures, the peaks should get more pronounced.
+In the absence of latent structures, the peaks should smoothen out, and the distribution should approach a normal distribution.
 \newline
 
-```{r rain_cloud_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "The rain cloud plot is comprised of three parts.
-The probability density distribution, the box plot and the raw data points.
-This allows for an transparent and bias assessment of the results space and its distribution."}
+```{r rain_clod_plot, echo = FALSE, fig.width = 8, fig.cap = "Raincloud plot.
+The results space ranges from about 1.05 to 1.4.
+The distribution showed two peaks which suggests an influential latent covariate structure.
+The original median effect was within the results space but outside the middle 50% of the data."}
 
@@
 
-## Specified covariate effects
-     The final graph shows the specified covariate effects.
-On the horizontal axis are the covariates, on the vertical axis the estimates.
+## Specifying covariate effects – scatter plot 
+     Fourth, to better understand the relevance of the covariates, another scatter plot was calculated (Figure \ref{fig:covar_plot}).
+This graph shows the specified covariate effects.
+On the horizontal axis are the covariates; on the vertical axis are their estimated effects.
 The covariates are sorted from negative to positive impact.
@@
 The error bars represent the 95% CI.
-Four out of 15 covariates are significant, roughly 27%.
-Thought the significant covariates' effects seem relatively weak, ranging from -0.1 to 0.05, only.
-Interesting is also which covariates are significant.
-The *team 23* analysis included league, position, referee and player as covariates.
-Of those only player is a significant covariate.
-Using crude mathematics, one might make the case that the added effect of players is the reason for moving the overall effect from the median, roughly 1.25 to 1.30.
-Though that's more speculation than causal reasoning.
-As mentioned above, the roughly 90% of all outcome were statistically significant, given that only 27% of the covariates have an overall significant effects suggest that the effect is of skin tone itself is fairly robust.
+Four out of 15 covariates were significant.
+All effect sizes, including the significant covariates’, were relatively weak.
+The CIs were relatively small, which was likely due to the large sample size.
+The OR of the model without any covariates (i.e., core variables only) was 1.262.
+Adding the sum of all positive covariate effects (games + goals + referee country + player, respectively) to the base model resulted in a maximum possible effect of 1.357.
+The maximum observed OR was 1.382, however.
+This difference suggests that there were more impactful covariate combinations or interactions that have not yet been identified.
+Vice versa, subtracting the sum of all negative covariate effects from the base model resulted in an OR of 1.070, whereas the minimum observed effect was an OR of 1.081.
+This suggests that there were also covariate combinations or interactions with a higher negative impact.
 \newline
 
-```{r covar_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "The specified covariate effects indicate the extend to which a covariate influences the outcome measures."}
+```{r covar_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Scatter plot.
+Four out of 15 covariates were significant, although their effects were relatively weak.
+This suggests that covariates interactions were responsible for the higher outcome values."}
 
@@
 
-```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center"}
+## Exploring the results structure – volcano plot
+     Lastly, another graph was produced despite not being planned.
+Therefore, it is only used to strengthen previous points and probe future research briefly.
+The results above suggest one or more meaningful latent covariate interactions; these were further explored.
+Although @patel2015 had a different objective, their methodology was similar to the current project.
+They visualised the outcomes as a so-called volcano plot (see Figure \ref{fig:rain_cloud_plot}), showing the relationship between the p-values and effect sizes.
+A similar graph was calculated for the current project.
+The two distinct “streams” provided further support for two distinct underlying constructs.
+Future research might specify which covariate combination underlies this structure.
+
+```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Volcano plot.
+The exploratory data visualisation supports the hypothesis of a systematic latent covariate structure with at least two underlying latent constructs."}
 
@@
 
-\newpage
+# Discussion
+     In the present multiverse project, I performed not only all *reasonable* but for the first time all *possible* analysis strategies to define and understand the results space of @silberzahn2018.
+The objective was to investigate whether the space of all *reasonable* results covers the full range of *possible* results or whether it covers only a narrow sub-space of *possible* results.
+Due to time and computational limitations, the current project focused on one analysis strategy and a random sample of covariate combinations.
+The chosen analysis strategy (i.e., the statistical model) was the one producing the median outcome in the original study.
 
-# Discussion
-The current project used multiverse analysis to define the results space of the @silberzahn2018.
-Among others multiverse analysis combines different statistical modelling approaches along with covariate combinations.
-The goals is assess the robustness of effects.
-Due to time and resources limitations, I chose to focus on statistical modelling approach, but explore every possible covariate combination.
-Usually, multiverse analysis would focus on all plausible (or reasonable) analytic strategies.
-However, this project sought to define the results space of all *possible* analytical outcomes to investigate in which space the outcomes of the *plausible* analytical strategies vary in the first place.
-Only if the results space is defined, is possible assess its content.
-The concept of statistical association, provided a neat analogy.
-There is covariation and correlation, both quantify the variability of two variables.
-Both indicate the direction of relationship (positive, none, negative).
-However, to assess the strength of the relationship only correlation is informative.
-Correlation is the standardised version of covariation, hence, it has a scale to it.
-And only knowing the limits to space of variation gives the correlation coefficient meaning.
-The same applies to analyses on a broader level.
-If researchers want to be able to tell if their analysis reflects a true effect or simply a luck guess, they need to know the entire results space.
-Only then they can identify the robustness of effects.
+ First, this “median” analysis was successfully replicated, which ensured an identical starting point for all further analyses.
+Next, 1,000 unique covariate combinations were sampled, respective models were, and results were visualised.
+After removing outliers, the data were distributed in a relatively narrow results space ranging from roughly 1.1 to 1.4.
+The median outcome of @silberzahn2018 was not included in the middle 50% of the data, though it was part of the 95% confidence interval.
+Therefore, it is concluded that the covariates of that model had an above-average effect, i.e.
+this covariate combination was more influential than a random combination of covariates.
 
- Due to time and resource limitation this project focused on pne
+ Moreover, the distribution of the results space showed two peaks.
+To understand the distribution’s underlying structure, the covariate effects were analysed.
+The results showed that 4 out of 15 covariates were statistically significant, though their effects were small.
+Moreover, when adding the individual covariate effects to the outcome of the base model, i.e.
+the model without any covariates, the observed outcome values are always well within the results space.
+Thus, interactions of covariates likely contribute to the results space.
+The final volcano plot supports this hypothesis by showing at least two streams indicating at least two underlying constructs.
+Future research might investigate these latent structures.
+In short, the current project found that the full results space is narrower than the CIs of the original “median” analysis of @silberzahn2018 indicated.
+This suggests that the original study overestimated the variability.
+This was surprising as the full results space would have been expected to be larger than the estimated CIs.
+If the variability were overestimated for all analyses, this would suggest that there is more agreement between the analyses than previously assumed.
 
-In order to assess the strength of the association covariation is hardly useful as covariate it does not have a standarised scale to it.
+ A common problem in research is p-hacking, i.e.
+tweaking the analysis to find significant effects, which are considered more “publishable”.
+Recent preventive efforts include pre-registering the theoretical framework and analysis plan.
+The idea is to be transparent about the analysis strategy and holding oneself accountable to the pre-defined standards.
+Additionally, the data and analysis code are made publicly available.
+These effects also enabled the current project.
+@silberzahn2018 found that 2/3 of all outcomes were statistically significant.
+This poses the question of whether researchers fished for significant results or whether these outcomes reflect the true relationship.
+To scrutinise this question, @patel2015 suggested assessing the robustness of an effect, i.e.
+its consistency in different conditions.
+The current project built on their methodology and also assessed the robustness of an effect.
+Considering that about 90% of all observed outcomes were significant, the effect of skin tone is considered to be robust.
+Hence, researchers have likely detected a true effect.
+
+ All [materials](https://github.com/sebastianplnr/msc_dissertation_project) used for this analysis are publicly available, and all results should be reproducible and expendable.
+The biggest constraint of this project were time and computational resources.
+However, reasonable decisions were made to achieve the best possible compromise between rigour and feasibility.
+The constraints and the findings of the current project yield great potential for future research.
+A first step could be to run not only a random sample but all possible covariate combinations.
+It would be interesting to know whether this procedure would confirm or correct the results space observed in the present study.
+Moreover, it would be highly interesting to decipher the underlying covariate structure responsible for the observed “streams.” Moreover, it would be interesting to run all other statistical models, including all covariate combinations.
+Based on the present findings, one might hypothesise that the full result spaces are narrower than the estimated 95% CIs.
+If this hypothesis would be confirmed, further research might investigate how multiverse analysis can account for such overestimation.
 

### 2021-08-22 scripts/4_multiverse_analysis.R

@@ -143,7 +143,6 @@ n_sample = 1000
 prop_ranef = length(all_models[all_models_index]) / (length(all_models[!all_models_index]) + length(all_models[all_models_index]))
 prop_ef = length(all_models[!all_models_index]) / (length(all_models[!all_models_index]) + length(all_models[all_models_index]))
 
-
 ranef_sample =  sample(formula_ranef, n_sample*prop_ranef, replace = FALSE)
 ef_sample = sample(formula_ef, n_sample*prop_ef, replace = FALSE)
 
@@ -239,6 +238,7 @@ write.csv(df, here::here("data", "4_model_outcomes_data.csv"), row.names = FALSE
 # In case models, were not run, load their outcomes
 # df = read.csv(here::here("data", "4_model_outcomes_data.csv"))
 
+
 # Build data frame that includes the specifications and the analysis results
 analysed_specifications = inner_join(specifications, df, by = "formula")
 
@@ -262,6 +262,7 @@ analysed_specifications = analysed_specifications %>% filter(estimate_oddsratio 
 alpha_level = 0.05
 z = 1.96 # 95% CI
 analysed_specifications$below_alpha = with(analysed_specifications, ifelse(p_value < alpha_level, "Significant", "Non-significant"))
+
 
 # Selecting relevant columns for a better overview (covariates and "estimate_oddsratio")
 analysed_specifications = analysed_specifications[ , c(base_var, covar, "p_value", "below_alpha", "estimate_oddsratio", "ci_lower_oddsratio", "ci_upper_oddsratio")]
@@ -345,7 +346,7 @@ impact_df$below_alpha = with(impact_df, ifelse(as.numeric(impact_pvalue) < 0.05,
 write.csv(impact_df, here::here("data", "4_covariate_effect_data.csv"), row.names = FALSE)
 
 
-#.........................................# Covariate Effects plot #.......................................#
+#.........................................# Covariate effects plot #.......................................#
 
 # Order covariate effects low to high
 impact_df = impact_df[order(impact_df$estimate_oddsratio), ]

### 2021-08-23 dissertation/msc_dissertation.Rmd
Further editing the dissertation

@@
 
-knitr::opts_chunk$set(echo = FALSE, fig.pos = 'H')
+knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.align = "center", fig.pos = 'H')
 
@@
 
- \newline
+\newpage
 
-__ABSTRACT__ Numerous studies have show that analytical choices researchers make throughout the process of conducting a study affect the outcome.
-Recent crowdsourcing approaches to data analysis show that even well-intentioned analytical choices can make the difference between find a significant and strong or a non-significant effect.
-IN PROGRESS
-
-\newpage
+# Abstract
+Crowdsourcing and multiverse analysis usually only assess reasonable analysis strategies.
+Little is known about how the results space of these *reasonable* analyses relates to the results space of all *possible* analyses.
+The current project, therefore, aimed at defining and understanding the results space and structure, respectively, of all possible analyses for @silberzahn2018.
+Due to time and resources limitations, the project focused on the analysis producing the median outcome and randomly sampled 1,000 covariate combinations out of all possible ones.
+The findings indicate that the full results space is narrower than the original confidence intervals suggest.
+If this were to be true for all analyses conducted in @silberzahn2018, the variability between the analysis strategies were likely smaller than previously though.
+Future studies have to confirm this hypothesis by first running the "median model" with all covariate combinations and for each statistical model.
 
@@
 There were 29 different analyses with 21 different combinations of covariates.
-Twenty teams found a statistically significant effect of skin colour on the risk of being sent off the field.
+Twenty teams found a statistically significant effect of skin colour on the odds of being sent off the field.
 The authors also controlled for researchers’ prior believes and experience as well as peer-rated analysis quality, but these factors did not account for the variation of results.
@@
 
- Another approach to assess the effects of analysis strategies on research findings is multiverse analysis (also known as specification-curve analysis, [@simonsohn2020].
+ Another approach to assess the effects of analysis strategies on research findings is multiverse analysis [also known as specification-curve analysis, @simonsohn2020].
 This approach has been proposed by @botvinik-nezer2020 and @silberzahn2018 as an alternative to crowdsourcing.
@@
 @patel2015 developed a standardised approach to assess an effect's robustness.
-I used this standardised approach to define the results space.
+This standardised approach to define the results space was used.
 The approach essentially comprises two parameters: the statistical models and the covariates (or control variables).
@@
 As this number of strategies exceeds the current project's resources, it was reduced to a more manageable number.
-Therefore, the project focused on the modelling approach that produced the median outcome of all analyses.
+Therefore, the project focused on the modelling approach producing the median outcome of all analyses.
 The median, being the middle number of a given set of values, is a reasonable starting point to estimate the results space.
@@
 
- In @silberzahn2018, the median outcome was produced by @team23, which will hereinafter be referred to as *team 23* due to its team number in the original study.
+ In @silberzahn2018, the median outcome was produced by @team23, which will hereinafter be referred to as *Team 23* due to its team number in the original study.
 *Team 23* first transformed the data and then conducted a mixed-model logistic regression.
-The first step of the present project was to replicate their transformation and analysis.
+The first step of the current project was to replicate their transformation and analysis.
 Such replication can increase the confidence in the previous and the current approach.
-Moreover, it ensures that this project has the same starting point as *team 23*.
-As *team 23* made all scripts publicly available, this step was straightforward.
+Moreover, it ensures that this project has the same starting point as *Team 23*.
+As team made all scripts publicly available, this step was straightforward.
 The next step of the project was to define the results space and its structure.
@@
 These parameters were the coefficient (i.e., effect) of skin tone, its standard error, test statistic and p-value.
-Similar to *Team 23* I calculated the 95% confidence intervals (CI).
+Similar to *Team 23* the 95% confidence intervals (CI) were calculated.
 The estimates and their CIs were transformed to odds ratios (OR) through exponentiating them to the power of two.
@@
 
-(ii) The raincloud plot @allen2021 combines a probability density plot, a box plot, and a scatter plot to give an unbiased, transparent view of the raw data.
+(ii) The raincloud plot [@allen2021] combines a probability density plot, a box plot, and a scatter plot to give an unbiased, transparent view of the raw data.
 If the three plots are stacked vertically from top to bottom, they look like an eponymous cloud with rain drops.
@@
 The data exploration included the reasoning behind transforming the data and some cleaning.
-The data transformation restructured the data into a more intuitive format (more details on the topics of exploration and transformation are in the data section).
+The data transformation restructured the data into a more intuitive format (more details on the topics of exploration and transformation are in the "data" section).
 The analysis script prepared the data by assigning variable classes (factors, numerical or boolean) and standardised a few variables, i.e.
@@
  The code was based on @haessler2020 and @patel2015.
-The basic structure of the code is shown in the pseudocode table (see Algorithm \ref{alg:pseudocode}).
+The basic structure of the code is shown in the pseudocode table (see Pseudocode \ref{alg:pseudocode}).
 Here, two important analytical decisions are briefly motivated.
@@
 \begin{algorithm}
+\SetAlgorithmName{Pseudocode}{pseudocode}{List of pseudocodes}
 \caption{Outlining high-level structure of the code. "Ran" in e.g. "FormulaRanEf" refers to "random-structure". "Ef" refers to "effect".}
@@
   $OutputsRanEf$ $\leftarrow$ \textbf{lapply(data, $ExtractorFunction$(glmer($dv$, x = $basevar + SampleRanEf$)))} \\
-  $OutputsRanEfEF$ $\leftarrow$ \textbf{lapply(data, $ExtractorFunction$(glm($dv$, x = $basevar + SampleEf$)))} \\
+  $OutputsEf$ $\leftarrow$ \textbf{lapply(data, $ExtractorFunction$(glm($dv$, x = $basevar + SampleEf$)))} \\
   \BlankLine
-  $ModelOutputs$ $\leftarrow$ Merge $OutputsRanEf$ and $OutputsEF$ \\
+  $ModelOutputs$ $\leftarrow$ Merge $OutputsRanEf$ and $OutputsEf$ \\
   $ModelOutputs$ $\leftarrow$ Compute odds ratios and confidence intervals \\
@@
 The code from *Team 23* [@team23] can be retrieved from their [OSF repository](https://osf.io/akqt4/).
-Following R packages were used *here 1.0.1* [@here], *data.table 1.14.0* [@datatable], *tidyverse 1.3.1* [@tidyverse], *lme4 1.1-27.1* [@lme4], *pbmcapply 1.5.0* [@pbmcapply], *PupillometryR 0.0.3* [@PupillometryR] and *cowplot 1.1.1* [@cowplot].
+Following R packages were used: *here 1.0.1* [@here], *data.table 1.14.0* [@datatable], *tidyverse 1.3.1* [@tidyverse], *lme4 1.1-27.1* [@lme4], *pbmcapply 1.5.0* [@pbmcapply], *PupillometryR 0.0.3* [@PupillometryR] and *cowplot 1.1.1* [@cowplot].
 
@@
 disaggregated.
-Further EDA showed that receiving a red card was highly unlikely ($p = 0.008%$), i.e.
+Further EDA showed that receiving a red card was highly unlikely (0.8%), i.e.
 the data were highly skewed.
@@
 If a referee has less than 22 player-encounters, there are missing cases.
-According to team 23, this was mostly the case in referees officiating games of minor leagues.
-As they focused on the major European leagues (England, Germany, Italy, and Spain), referees with less than 22 player-encounters were excluded.
+According to *Team 23*, this was mostly the case in referees officiating games of minor leagues.
+As they focused on the major European leagues (England, Germany, France, and Spain), referees with less than 22 player-encounters were excluded.
 This step excluded about 66% of the referees but retained 97.4% of all player-referee combinations.
@@
 
-```{r var_plots, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Team 23 base variables properties.
+```{r var_plots, fig.cap = "Team 23 base variables properties.
 A) There was a strong skew towards receiving no red card compared to receiving a red card.
@@
 Explicit bias scores were on the verge of being non-significant.
-Thus, the results of team 23 were reproducible without any adjustment.
+Thus, the results of *Team 23* were reproducible without any adjustment.
 
@@
 Overall, 89.7% of the outcomes were significant.
-Except for the lowest outcome ($OR_{lowest} = 1.080$) all points were close to their neighbours.
+Except for the lowest outcome ($OR_{lowest} = 1.081$) all points were close to their neighbours.
 This indicates that none of the specified models was responsible for a sudden increase in the outcome measure.
-The bottom table lists all covariates on its vertical axis, sorted from most positive to most negative impact.
-For instance, the covariate player had the most substantial positive impact, while the covariate club had the most substantial negative impact.
-Each column of the table represents one mode.
+The bottom table lists all covariates on its vertical axis, sorted from most positive (top) to most negative impact (bottom).
+For instance, the covariate "player" had the most substantial positive impact, while the covariate "club" had the most substantial negative impact.
+Each column of the table represents one model.
 A coloured cell indicates the presence of the variable, an uncoloured cell its absence.
@@
 It is, however, interesting to look at the top and bottom rows of the table.
-The top row shows that the covariate player was particularly often included in models with a higher outcome.
-Conversely, the covariate club was particularly often included in models with a lower outcome.
+The top row shows that the covariate "player" was particularly often included in models with a higher outcome.
+Conversely, the covariate "club" was particularly often included in models with a lower outcome.
 \newline
 
-```{r sca_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Specification curve.
-Results space ranges from about 1.05 to 1.4.
+```{r sca_plot, fig.cap = "Specification curve.
+Results space ranges from about 1.1 to 1.4.
 However, due to the large number of specifications, the specification curve does not allow for identifying the relevance of the different covariates.", warning = FALSE}
@@
 The 1st quartile was 1.206, the median 1.248 and the 3rd quartile 1.293.
-Thus, the interquartile range where 50% of the data were included was relatively small (1.206-1.293), indicating a low statistical dispersion.
-The dashed line indicates the original results of *Team 23*, which has been the median outcome of all analysis strategies in @silberzahn2018.
+Thus, the interquartile range where 50% of the data were included was relatively narrow (1.206 to 1.293), indicating a low statistical dispersion.
+The dashed line indicates the original result of *Team 23*, which has been the median outcome of all analysis strategies in @silberzahn2018.
 The outcome of *Team 23* (1.31) was included in the results space but outside the middle 50% of the data.
@@
 
- The probability density distribution reveals another exciting feature of the data.
+ The probability density distribution reveals another interesting feature of the data.
 The distribution has two peaks which suggest two latent covariate structures whose outcomes centre around two points.
@@
 
-```{r rain_clod_plot, echo = FALSE, fig.width = 8, fig.cap = "Raincloud plot.
-The results space ranges from about 1.05 to 1.4.
+```{r rain_cloud_plot, fig.cap = "Raincloud plot.
+The results space ranges from about 1.08 to 1.38.
 The distribution showed two peaks which suggests an influential latent covariate structure.
-The original median effect was within the results space but outside the middle 50% of the data."}
+The original median effect was within the results space but outside the middle 50 percent of the data."}
 
@@
 Vice versa, subtracting the sum of all negative covariate effects from the base model resulted in an OR of 1.070, whereas the minimum observed effect was an OR of 1.081.
-This suggests that there were also covariate combinations or interactions with a higher negative impact.
+This suggests that there were also covariate combinations with a higher negative impact.
 \newline
 
-```{r covar_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Scatter plot.
+```{r covar_plot, fig.cap = "Scatter plot.
 Four out of 15 covariates were significant, although their effects were relatively weak.
@@
      Lastly, another graph was produced despite not being planned.
-Therefore, it is only used to strengthen previous points and probe future research briefly.
-The results above suggest one or more meaningful latent covariate interactions; these were further explored.
+Therefore, it is only used to strengthen previous points and probe future research.
+The results above suggest one or more meaningful latent covariate combinations; these were further explored.
 Although @patel2015 had a different objective, their methodology was similar to the current project.
@@
 A similar graph was calculated for the current project.
-The two distinct “streams” provided further support for two distinct underlying constructs.
-Future research might specify which covariate combination underlies this structure.
+The two (potentially three) distinct “streams” provided further support for two (or three) distinct underlying constructs.
+Future research might specify which covariate combinations underly these structures.
 
-```{r volcano_plot, echo = FALSE, fig.width = 8, fig.align = "center", fig.cap = "Volcano plot.
-The exploratory data visualisation supports the hypothesis of a systematic latent covariate structure with at least two underlying latent constructs."}
+```{r volcano_plot, fig.cap = "Volcano plot. The exploratory data visualisation supports the hypothesis of a systematic latent covariate structure with at least two underlying latent constructs."}
 
@@
   scale_color_manual(values = c("red", "black")) +
-  labs(title = "Systemtic latent structure",
+  labs(title = "Systemtic latent structures",
        subtitle = paste0("N = ", nrow(analysed_specifications), collapse = "")) +
@@
 # Discussion
-     In the present multiverse project, I performed not only all *reasonable* but for the first time all *possible* analysis strategies to define and understand the results space of @silberzahn2018.
+     In the present multiverse project, not only all *reasonable* but for the first time all *possible* analysis strategies were performed to define and understand the results space of @silberzahn2018.
 The objective was to investigate whether the space of all *reasonable* results covers the full range of *possible* results or whether it covers only a narrow sub-space of *possible* results.
@@
 
- First, this “median” analysis was successfully replicated, which ensured an identical starting point for all further analyses.
-Next, 1,000 unique covariate combinations were sampled, respective models were, and results were visualised.
-After removing outliers, the data were distributed in a relatively narrow results space ranging from roughly 1.1 to 1.4.
+ First, the median analysis was successfully replicated, which ensured an identical starting point for all further analyses.
+Next, 1,000 unique covariate combinations were sampled, respective models were run, and results were visualised.
+After removing outliers, the data were distributed in a relatively narrow results space ranging from about 1.1 to 1.4.
 The median outcome of @silberzahn2018 was not included in the middle 50% of the data, though it was part of the 95% confidence interval.
@@
 These effects also enabled the current project.
-@silberzahn2018 found that 2/3 of all outcomes were statistically significant.
+@silberzahn2018 found that $2/3$ of all outcomes were statistically significant.
 This poses the question of whether researchers fished for significant results or whether these outcomes reflect the true relationship.
@@
 its consistency in different conditions.
-The current project built on their methodology and also assessed the robustness of an effect.
+The current project built on their methodology and, therefore, also assessed the robustness of an effect.
 Considering that about 90% of all observed outcomes were significant, the effect of skin tone is considered to be robust.

### 2021-08-23 scripts/4_multiverse_analysis.R
Cleaning up viz

@@ -278,7 +278,7 @@ write.csv(analysed_specifications, here::here("data", "4_model_outcomes_specific
 
 #............................................# Rain cloud plot #...........................................#
 
-vibration_of_effect = analysed_specifications %>%
+results_space_plot = analysed_specifications %>%
   filter(estimate_oddsratio < 4) %>% 
   ggplot(aes(x = "", y = estimate_oddsratio)) +
   geom_flat_violin(aes(fill = ""), trim = FALSE, colour = "dark grey", fill = "dark grey") +
@@ -289,7 +289,7 @@ vibration_of_effect = analysed_specifications %>%
   annotate("text", x = 1.58, y = 1.375, label = "Silberzahn et al. (2018) - Median OR", color = "black") +
   scale_color_manual(values = c("red", "black")) +
   scale_y_continuous(name = "Odds ratio", breaks = c(seq(1.0, 1.5, 0.05)), limits = c(0.99, 1.51), expand = c(0, 0)) +
-  labs(title = "Vibration of effect due to covariate specification",
+  labs(title = "Results space defined through covariate specification",
        subtitle = paste0("N = ", nrow(analysed_specifications), ", Odds ratio (OR)", collapse = "")) +
   coord_flip() +
   theme_classic() +
@@ -310,9 +310,9 @@ vibration_of_effect = analysed_specifications %>%
         plot.subtitle = element_text(face = "italic")) +
   guides(colour = guide_legend(override.aes = list(alpha = 1), reverse = TRUE))
 
-vibration_of_effect
+results_space_plot
 
-ggsave(here::here("figures", "vibration_of_effect.png"), vibration_of_effect, width = 11, height = 5)
+ggsave(here::here("figures", "results_space_plot.png"), results_space_plot, width = 11, height = 5)
 
 
 #............................................# Covariates effects #.........................................#
@@ -351,7 +351,7 @@ write.csv(impact_df, here::here("data", "4_covariate_effect_data.csv"), row.name
 # Order covariate effects low to high
 impact_df = impact_df[order(impact_df$estimate_oddsratio), ]
 
-covariate_effects = impact_df %>%
+covariate_effects_plot = impact_df %>%
   ggplot(aes(x = reorder(x = impact_names, X = as.numeric(impact_coef)), y = as.numeric(impact_coef))) +
   geom_point(aes(color = factor(below_alpha))) +
   geom_hline(yintercept = 0) +
[…]

### 2021-08-23 dissertation/msc_dissertation.Rmd
Update dissertation

@@
 The strength of the raincloud plot is that it visualizes the raw data, summary statistics (median, 25th and 75th quartiles and CIs) and the probability density in a single plot.
-As the current project seeks to define and describe a results space, the raincloud plot is an ideal tool (for an example check Figure \ref{fig:rain_cloud_plot}).
+As the current project seeks to define and describe a results space, the raincloud plot is an ideal tool (see Figure \ref{fig:results_space_plot} for an example).
 However, the plot does not allow to identify which covariates cause which outcomes.
@@
 ## Exploring the results space – raincloud plot
-     Third, the results space was further explored by calculating raincloud plots (Figure \ref{fig:rain_cloud_plot}).
+     Third, the results space was further explored by calculating raincloud plots (Figure \ref{fig:results_space_plot}).
 The horizontal axis shows the OR.
@@
 
-```{r rain_cloud_plot, fig.cap = "Raincloud plot.
+```{r results_space_plot, fig.cap = "Raincloud plot.
 The results space ranges from about 1.08 to 1.38.
@@
 
-vibration_of_effect = dat_models_specifications %>%
+results_space_plot = dat_models_specifications %>%
   filter(estimate_oddsratio < 4) %>% 
@@
 
-vibration_of_effect
+results_space_plot
 
@@
 ## Specifying covariate effects – scatter plot 
-     Fourth, to better understand the relevance of the covariates, another scatter plot was calculated (Figure \ref{fig:covar_plot}).
+     Fourth, to better understand the relevance of the covariates, another scatter plot was calculated (Figure \ref{fig:covariate_effects_plot}).
 This graph shows the specified covariate effects.
@@
 
-```{r covar_plot, fig.cap = "Scatter plot.
+```{r covariate_effects_plot, fig.cap = "Scatter plot.
 Four out of 15 covariates were significant, although their effects were relatively weak.
@@
 
-covariate_effects = dat_covariates %>%
+covariate_effects_plot = dat_covariates %>%
   ggplot(aes(x = reorder(x = impact_names, X = impact_coef), impact_coef)) +
@@
 
-covariate_effects
+covariate_effects_plot
 
@@
 Although @patel2015 had a different objective, their methodology was similar to the current project.
-They visualised the outcomes as a so-called volcano plot (see Figure \ref{fig:rain_cloud_plot}), showing the relationship between the p-values and effect sizes.
+They visualised the outcomes as a so-called volcano plot (see Figure \ref{fig:latent_structure_plot}), showing the relationship between the p-values and effect sizes.
 A similar graph was calculated for the current project.
@@
 
-```{r volcano_plot, fig.cap = "Volcano plot. The exploratory data visualisation supports the hypothesis of a systematic latent covariate structure with at least two underlying latent constructs."}
+```{r latent_structure_plot, fig.cap = "Volcano plot.
+The exploratory data visualisation supports the hypothesis of a systematic latent covariate structure with at least two underlying latent constructs."}
 
-volcano_plot = dat_models_specifications %>% 
+latent_structure_plot = dat_models_specifications %>% 
   filter(estimate_oddsratio < 4) %>% 
@@
 
-volcano_plot
+latent_structure_plot
 

### 2021-08-23 scripts/4_multiverse_analysis.R
Another update

@@ -366,13 +366,13 @@ covariate_effects_plot = impact_df %>%
         axis.text.x = element_text(angle = 45, hjust = 1),
         plot.margin = unit(c(0.5, 1, 0.5, 1), "cm"),
         legend.position = "top",
-        legend.justification = "left",
+        legend.justification = "right",
         legend.title.align = 1,
-        legend.text = element_text(size = 11),
-        legend.direction = "vertical",
+        legend.text = element_text(size = 10),
+        legend.direction = "horizontal",
         legend.title = element_blank(),
-        legend.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
-        legend.box.margin = margin(l = 0.1, b = -0.719, unit = "cm"),
+        legend.margin = margin(t = -0.33, b = -0.2, unit = "cm"),
+        legend.box.margin = margin(t = -0.33, b = -0.2, unit = "cm"),
         plot.title = element_text(face = "bold"),
         plot.subtitle = element_text(face = "italic")) +
   guides(colour = guide_legend(reverse = TRUE))

### 2021-08-23 dissertation/msc_dissertation.Rmd
Final commit

@@
   | Dissertation
-author: "Sebastian Ploner, sploner1@sheffield.ac.uk, University of Sheffield"
+author: "Registration no.: 200256786"
 output:
@@
 # Abstract
-Crowdsourcing and multiverse analysis usually only assess reasonable analysis strategies.
-Little is known about how the results space of these *reasonable* analyses relates to the results space of all *possible* analyses.
-The current project, therefore, aimed at defining and understanding the results space and structure, respectively, of all possible analyses for @silberzahn2018.
-Due to time and resources limitations, the project focused on the analysis producing the median outcome and randomly sampled 1,000 covariate combinations out of all possible ones.
+
+Most research studies can be analysed by using different strategies which can cause variability in the results.
+Crowdsourcing and multiverse analysis are recent approaches to better understand the influence of analysis strategies on research findings.
+These approaches usually assess different *reasonable* but not all *possible* analysis strategies.
+Thus, little is known about how the results space of these *reasonable* analyses relates to the results space of all *possible* analyses.
+The current project therefore aimed at defining and understanding the results space and structure of all possible analyses for an exemplary study by @silberzahn2018.
+To address this topic, the project focused on one statistical model producing the median outcome and calculated the model for 1,000 randomly sampled covariate combinations.
 The findings indicate that the full results space is narrower than the original confidence intervals suggest.
-If this were to be true for all analyses conducted in @silberzahn2018, the variability between the analysis strategies were likely smaller than previously though.
-Future studies have to confirm this hypothesis by first running the "median model" with all covariate combinations and for each statistical model.
+Thus, the variability between the analysis strategies in @silberzahn2018 was likely smaller than previously thought.
+Future investigations might extend the present study by running even more statistical models with all possible covariate combinations.
 
@@
 Although the approach is statistically complex and computationally intense, it can be performed by one or a few researchers.
-Moreover, a researcher’s inherent strategy bias is neutralised, and noise is made transparent [@simonsohn2020].
+Moreover, it neutralises a researcher’s inherent strategy bias and makes noise transparent [@simonsohn2020].
 Multiverse approaches are therefore well suited to better understand how different analysis strategies influence research results.
@@
 
+\newpage
+
 # Analysis
@@
 In @silberzahn2018, the analysts used numerous statistical models like multiple linear regression, mixed-model logistic regression or Bayesian logistic regression.
-In total, there were 29 different modelling approaches.
+In total, they used 29 different modelling approaches.
 Additionally, each team used a different set of covariates.
@@
 Therefore, the project focused on the modelling approach producing the median outcome of all analyses.
-The median, being the middle number of a given set of values, is a reasonable starting point to estimate the results space.
+The median is the middle number of a given set of values and therefore a reasonable starting point to estimate the results space.
 However, even focusing on one analysis approach leaves $1 * 2^{15} = 32,768$ possibilities.
@@
 Moreover, it ensures that this project has the same starting point as *Team 23*.
-As team made all scripts publicly available, this step was straightforward.
+As *Team 23* made all scripts publicly available, this step was straightforward.
 The next step of the project was to define the results space and its structure.
@@
 Except for the lowest outcome ($OR_{lowest} = 1.081$) all points were close to their neighbours.
-This indicates that none of the specified models was responsible for a sudden increase in the outcome measure.
+This indicates that none of the specified models caused a sudden increase in the outcome measure.
 The bottom table lists all covariates on its vertical axis, sorted from most positive (top) to most negative impact (bottom).
@@
 The dashed line indicates the original result of *Team 23*, which has been the median outcome of all analysis strategies in @silberzahn2018.
-The outcome of *Team 23* (1.31) was included in the results space but outside the middle 50% of the data.
+The outcome of *Team 23* (1.31) was included in the results space but outside the interquartile range.
 The CI calculated by team 23 ranged from 1.10 to 1.56.
@@
 In contrast, the upper bound of their CI was notably higher than the present results space.
-This suggests a potential overestimating of the variability of the results in the original analysis.
+This suggests a potential overestimation of the variability of the results in the original analysis.
 
@@
 The distribution showed two peaks which suggests an influential latent covariate structure.
-The original median effect was within the results space but outside the middle 50 percent of the data."}
+The original median effect was within the results space but outside the interquartile range."}
 
@@
 Four out of 15 covariates were significant.
-All effect sizes, including the significant covariates’, were relatively weak.
+All effect sizes, including the significant covariates' ones, were relatively weak.
 The CIs were relatively small, which was likely due to the large sample size.
@@
 ## Exploring the results structure – volcano plot
-     Lastly, another graph was produced despite not being planned.
+     Lastly, another graph, which had not been planned, was produced.
 Therefore, it is only used to strengthen previous points and probe future research.
@@
 Future research might investigate these latent structures.
-In short, the current project found that the full results space is narrower than the CIs of the original “median” analysis of @silberzahn2018 indicated.
+Moreover, the current project found that the full results space is narrower than the CIs of the original “median” analysis of @silberzahn2018 indicated.
 This suggests that the original study overestimated the variability.
